### 2010-12-02 trunk/cs870/project/Report/Paper.tex

\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran}
\usepackage{cite}
\usepackage{url}
\usepackage{times}
\usepackage{graphicx}
\usepackage{url}
\usepackage{clrscode}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{array}
\usepackage{color}
\usepackage{balance}

\begin{document}

\title{An Implementation of the Active Contours without Edged model and the Logic Framework for Active contours on multi-channel images}

\author{Karim Ali and Sarah Nadi\\
\{karim, snadi\}@cs.uwaterloo.ca \\
David R. Cheriton School of Computer Science\\
University of Waterloo\\
}

% The paper headers
\markboth{CS870 Project Report - Active Contour without Edges}%
{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}

\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
%\boldmath
The abstract goes here.
\end{abstract}
}

\maketitle

\section{Introduction}
Image segmentation or boundary detection is a very important problem in the area of Image Processing, and has received a lot of attention in the past. The
classical Active Contour model (or Snakes) proposed by Kass et al.~\cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it very highly adopted later on
(E.g.~\cite{caselles1997geodesic}. Most of these models used the level set formulation for propagating fronts to evolve the curve. However, these models highly
depended on curvature motion (motion defined by the gradient of the curve) which led to poor performance in smoothed edges. To overcome these limitations, Chan
and Vese~\cite{chan2001active} propose an active contour model that does not depend on the edges (i.e the gradient) for propagating the curve to detect the
boundary of the object. Instead, they use a region-based approach based on the Mumford-Shah model~\cite{mumford1989optimal} to divide the image
into two regions: one inside the propagating curve, and one outside. The curve is at the boundary of the object if there is no difference in intensities inside
the curve as well as outside the curve.

The Active Contours without Edges model proposed by Chan and Vese (referred to as Chan-Vese model throughout the rest of this paper) was very successful in
detecting objects even in noisy or blurry images. It could also detect holes in objects which was usually a limitation in previous models. As an extension to
this work, Sandberg and Chan~\cite{sandberg2005logic} propose a logic framework (referred to as the Sandberg-Chan model throughout the rest of this
paper) that  performs logical operations on multiple images according to the curve propagation proposed in the Chan-Vese model. To achieve that, previous models
usually had two steps. They would either first segment the object in each channel separately then combine the segmented objects according to the logic operation
through bitwise operations (CITE) or they would apply logic operations to the different images then segment the resulting image (CITE). The first approach is
very costly, and the second approach requires a lot of prior knowledge about the intensities of each image. To overcome these drawbacks, the Sandberg-Chan
model is based on the idea of fitting a single contour to the object on all channels according to the logic operator, and based on regions.

In this paper, we report on our findings after implementing each of these models. We implemented both models in Matlab, and experimented with several images.
In this paper, we explain the details of our implementation as well as our results. The rest of this paper is organized as follows: Section~\ref{sec:bg} first
provides brief background about each of the two models implemented in this paper. Section~\ref{sec:implementation} then explains our implementation. In this
section, we explain how we implemented the models and any variations from the original papers. Section~\ref{sec:results} explains our evaluation criteria, and
presents the results we obtained. Section~\ref{sec:difficulties} discusses some of the difficulties we faced, and points out some BLA. Section~\ref{sec:concl}
concludes this paper by summarizing our findings.

\begin{figure*}[t]
\centering
\includegraphics[width=8cm]{explaining.png}
\caption{Region Based Model}
\label{fig:region}
\end{figure*}

\section{Background}
\label{sec:bg}

\subsection{Chan-Vese Model}
\label{sec:chan-vese}

The Chan-Vese model is a region based model for detecting objects in an image. It is based on a restriction Mumford-Shah model which divides an image into
regions and represents each region by a piecewise constant (the minimal partition problem). Figure~\ref{fig:region} shows what is meant by a region based
model. The figure shows an image $u_{0}$ which has a gray triangular object in it. The red curve, C, is the initial contour used to detect this object. The
main idea behind the model is that the curve divides the image into two regions: that inside the curve and that outside. Each region is represented by a
constant, c, which is the average intensity of the image values in each region. In order for the curve to fit the object, there must be no variation of the
intensities inside the curve as well as outside. In other words, this turns into a minimization problem of the difference of intensities inside plus those
outside. For example, in the figure, the point $(x_{1},y_{1})$ will have to be outside the curve in order to minimize the difference between the points inside
the curve. Similarly, the point $(x_{4}, y_{4})$ will have to be inside the curve.

More formally, the Euler-Lagrange equation (as derived in the original paper) representing the time motion of the curve C is shown in Equation~\ref{eqn:cvpde}.

\begin{equation}
\label{eqn:cvpde}
\frac{\partial{\phi}}{\partial{t}} = \delta_{\varepsilon}(\phi)[\mu\; \mathrm{div}(\frac{\bigtriangledown \phi}{|\bigtriangledown \phi|}) - \nu -
\lambda_{1}(u_{0} - c_{1})^2 +\lambda_{2}(u_{0} - c_{2})^2] 
\end{equation}

(REST OF EQUATION HERE)

The parameter $\mu$ is a scaling parameter for the length of the curve represented in terms of curvature. The smalled $\mu$ is, the more the length of the curve
can increase without penalizing the minimization. This allows the model to detect smalled objects and holes. The larger $\mu$ is, the less freedom there is for
the curve to increase in length, and thus, it will only be able to detect larger objects. The parameter $\nu$ is also a scaling term for the area of the curve.
However, the authors do not use the area term in the Euler-Lagrange derivation, and always set $\nu$ to 0. It seems that $\mu$ is sufficient to scale the curve
according to the objects that need to be detected. Finally, $\lambda_{1}$ and $\lambda_{2}$ are weighting parameters for the forces inside the curve and
outside the curve respectively. Since we want to give both forces equal weight, the authors set $\lambda_{1} = \lambda_{2} = 1$ in all their experiments.

\subsection{Sandberg-Chan Model}
\label{sec:sandberg-chan}

\section{Implementation}
\label{sec:implementation}

\section{Experimental Results}
\label{sec:results}

\subsection{Evaluation Criteria}
\label{sec:criteria}

In order to make sure we correctly implemented the models, we had several evaluation criteria. Table~\ref{tab:criteria} summarizes these criteria, and explains
the goals of each. For the rest of this section, we proceed by presenting our experimental results in the order of these criteria starting with the simplest
cases, and incrementally challenging the model.

\begin{table*}[h!t!]
\centering
\begin{tabular}{ c | c}
\textbf{Criteria} & \textbf{Goal} \\
\hline\hline
Detecting Boundaries & Ability to correctly detect object boundaries of simple objects\\
\hline
Curve Position & Ability to correctly detect object boundaries irrespective of the initial curve position\\
\hline
Detecting Holes & Ability to detect holes in objects, and not simply stop on outside boundary\\
\hline
Blurred Images & Ability to correctly (as much as possible) detect object boundaries in blurred images\\
\hline
Noisy Images & Ability to correctly (as much as possible) detect object boundaries in noisy images\\
\hline
Parameter Settings & Ability to respond correctly to the different parameter settings
\end{tabular}
\centering
\caption{Evaluation Criteria}
\label{tab:criteria}
\end{table*}

\subsection{Criteria 1: Successfully Detecting Boundaries}

\section{Difficulties and Discussion}
\label{sec:difficulties}

\section{Conclusion}
\label{sec:concl}
The conclusion goes here.

\bibliographystyle{IEEEtran}

\bibliography{references}

% insert where needed to balance the two columns on the last page with
% biographies
%\newpage

\end{document}

### 2010-12-02 trunk/cs870/project/CurveEvolution/evolveCurve.m

@@ -30,7 +30,7 @@ title('Initial Contour');
 % Initialize phi with phi0
 phi = phi0;
 
-% deltaT should satisfy the stability condition: deltaT = 1/h^2
+% deltaT should satisfy the stability condition: deltaT <= 1/h^2
 deltaT = (1/(imageSize))^2;
 
 % Plot the segmentation title

### 2010-12-02 trunk/cs870/project/CurveEvolution/evolveCurve.m
cleaned up evolveChannels

@@ -18,43 +18,40 @@ imageSize = size(image,1);
 
 % Plot the input image
 figure();
-subplot(2,2,1);
+subplot(2, 2, 1);
 imshow(image);
 title('Input Image');
 
 % Plot the initial contour
-subplot(2,2,2);
-contour(flipud(phi0), [0 0], 'r','LineWidth',1);
+subplot(2, 2, 2);
+contour(flipud(phi0) ,[0 0], 'r', 'LineWidth', 1);
 title('Initial Contour');
 
 % Initialize phi with phi0
 phi = phi0;
 
 % deltaT should satisfy the stability condition: deltaT <= 1/h^2
-deltaT = (1/(imageSize))^2;
-
-% Plot the segmentation title
-subplot(2,2,3);
-title('Segmentation');
-
+deltaT = (1 / (imageSize))^2;    
+    
 %---------------------------------------------------------------------------
 % Evolve the curve now!
 for n=1:iterations
     % Draw the original image
-    subplot(2,2,3);
-    imshow(image, 'initialmagnification','fit','displayrange',[0 255]);
+    subplot(2, 2, 3);
+    imshow(image ,'initialmagnification', 'fit', 'displayrange', [0 255]);
+    title('Segmentation');
     hold on;
-    
+
     % Plot the contour on the image
-    contour(phi, [0 0], 'r','LineWidth',1);
-    title(strcat('Iteration: ',num2str(n)));
+    contour(phi, [0 0], 'r','LineWidth', 1);
+    title(strcat('Iteration: ', num2str(n)));
     drawnow;
     
     % Plot the segmentation
-    seg = phi>0;
-    subplot(2,2,4);
+    seg = phi >= 0;
+    subplot(2, 2, 4);
     imshow(seg);
-    title(strcat('Segmentation at iteration: ' ,num2str(n)));
+    title(strcat('Segmentation at iteration: ', num2str(n)));
     
     % Update phi
     phi_new = updatePhi(phi, image, deltaT, mu, nu, lambda);
@@ -70,10 +67,10 @@ for n=1:iterations
 end
 
 % Display the final segmentation
-seg = phi>0;
-subplot(2,2,4);
+seg = phi >= 0;
+subplot(2, 2, 4);
 imshow(seg);
-title(strcat('Final segmentation at iteration: ' ,num2str(n)));
+title(strcat('Final segmentation at iteration: ', num2str(n)));
 
 % Release the hold on the plot
 hold off;

### 2010-12-02 trunk/cs870/project/CurveEvolution/updatePhiChannels.m

@@ -11,7 +11,7 @@
 %   logicop = the required logical operator
 %   varargin = the number of channels to operate on
 
-function resultingPhi = updatePhiChannels(phi,deltaT, mu, lambda, ...
+function resultingPhi = updatePhiChannels(phi, deltaT, mu, lambda, ...
                                             logicop, varargin)
 
 % |gradient(phi)| is used instead of the dirac delta function. This was an

### 2010-12-02 cs870/project/CurveEvolution/evolveChannels.m
cleaned up evolveChannels

@@ -1,69 +1,88 @@
-function evolveChannels(iterations)
+% evolveChannels: evlove the curve along multiple channels
+%
+% Input parameters:
+%   iterations = number of iterations
+%   phi0 = the intial contour
+%   mu = weight for the length of the curve
+%   lambda = weight for the inside and outside energies
+%   logicop = the logical operator to be applied
+%   image = the original image
+%   doReinit = reinitializing phi is optional
+%   varargin = multiple input channels, each channel is a 2D matrix
 
-plotStep = 1;               
-t0 = 0;                      % Start at time t = 0
+function evolveChannels(iterations, phi0, mu, lambda, logicop, ...
+                        image, doReinit, varargin)
 
-image1 = rgb2gray(imread('data/donuttopright.jpg'));
-image2 = rgb2gray(imread('data/donutbottomleft.jpg'));
-image3 = rgb2gray(imread('data/donutbottomright.jpg'));
+%---------------------------------------------------------------------------
+% Initializing some parameters
 
-objective = rgb2gray(imread('data/donut.jpg'));
+% Get the image size, any channel should work since they have the same size
+imageSize = size(varargin{1}, 1);
 
-display('in evolve channels');
-imageSize = size(image1,1); % either image should work since they are the same size
+% Set some parameters to be used to organize the subplots
+subPlotColumns = size(varargin, 2) + 1;
+subPlotRows = 2;
 
-
-phi0 =   cone(80, [100 100], imageSize);
-
+% Plot the input channels
 figure();
-numOfSubplots = 3;
-subplot(2,numOfSubplots,1); imshow(image1); title('Input Image 1');
-subplot(2,numOfSubplots,2); imshow(image2); title('Input Image 2');
-subplot(2,numOfSubplots,3); contour(flipud(phi0), [0 0], 'r','LineWidth',1); title('initial contour');
+for n=1:size(varargin, 2)
+    subplot(subPlotRows, subPlotColumns, n);
+    imshow(varargin{n});
+    title(strcat('Input Channel ', num2str(n)));
+end
 
-image1 = double(image1);
-image2 = double(image2);
-image3 = double(image3);
-
+% Plot the initial contour
+subplot(subPlotRows, subPlotColumns, subPlotColumns);
+contour(flipud(phi0), [0 0], 'r', 'LineWidth', 1);
+title('Initial Contour');
 
 % Initialize phi with phi0
[…]

### 2010-12-02 cs870/project/Report/Paper.tex

--- 2010-12-02 trunk/cs870/project/Report/Paper.tex
@@
 
-\begin{figure*}[t]
-\centering
-\includegraphics[width=8cm]{explaining.png}
-\caption{Region Based Model}
-\label{fig:region}
-\end{figure*}
-
 \section{Background}
@@
 constant, c, which is the average intensity of the image values in each region. In order for the curve to fit the object, there must be no variation of the
-intensities inside the curve as well as outside. In other words, this turns into a minimization problem of the difference of intensities inside plus those
-outside. For example, in the figure, the point $(x_{1},y_{1})$ will have to be outside the curve in order to minimize the difference between the points inside
-the curve. Similarly, the point $(x_{4}, y_{4})$ will have to be inside the curve.
+intensities inside the curve as well as outside. In other words, this turns into a minimization problem of the difference of intensities inside (fitting term
+one, $F_{1}$) plus those
+outside (fitting term two, $F_{2}$). For example, in the figure, the point $(x_{1},y_{1})$ will have to be outside the curve in order to minimize the difference
+between the points inside
+the curve. Similarly, the point $(x_{4}, y_{4})$ will have to be inside the curve. Figure~\ref{fig:fitting} shows all the possibilities of the curve's fitting
+according to its position with respect to the object.
 
@@
 \end{equation}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=6cm]{explaining.png}
+\caption{Region Based Model}
+\label{fig:region}
+\end{figure}
 
@@
 
+The main goal of the Sandberg-Chan model~\cite{sandberg2005logic} is to perform logic operations on a combination of different images accurately and
+efficiently. For example, finding the union of two images where different parts of the object are occluded in each image so that a complete object can be
+obtained. In order to do that, they use the Chan-Vese model to detect objects, and simultaneously include the logic operations in the minimization problem.
+Since they are using the Chan-Vese model, the problem must be viewed in terms of regions as well. Accordingly, they define the two main logic operations
+(union and intersection) in terms of regions where the union of two images is the union of the insides of the curve with respect to the images plus the
+intersection of the outsides of the curve with respect to the images. For example, Figure~\ref{fig:logic-op} (taken from the original paper) shows that the
+union of $A_{1}$ and $A_{2}$ (in terms of its shape) can be obtained by taking the union of the insides of the object or the intersection of the outsides.
+Similarly, the intersection of two images would be the intersection of the insides plus the union of the outsides. Thus, in order to obtain accurate results
+irrespective of the object's intensity outside versus inside, we should consider \textit{both} the logical operation required for the region inside the object
+as well as that outside.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=6cm]{fitting.png}
+\caption{Chan-Vese Curve Fitting}
+\label{fig:fitting}
+\end{figure}
+
+More formally, the authors define two logic variables $z_{i}^{in}$ and $z_{i}^{out}$ to denote whether a point $(x,y)$ should be in the moving curve C or not
+with respect to image i. Since we are trying to minimize the fitting of the curve, they use 0 to denote true and 1 to denote false (the reverse of the usual
+convention. Following the Chan-Vese model, they represent each of the regions inside and outside the curve by a constant c. In this paper, they represent
+$c_{in}$ as $c_{+}$ and $c_{out}$ as $c_{-}$. Equations~\ref{eqn:zin} and~\ref{eqn:zout} show how to calculate $z_{in}$ and $z_{out}$ respectively in terms of
+the Chan-Vese model. We note here that in the original paper there was a typo in these equations where they divide by the maximum intensity of each image
+instead of the maximum intensity squared. However, in order to have $z_{in}$ and $z_{out}$ have values from 0 to 1 that represent logic values, we need to
+divide by the maximum intensity squared. The equations shown below have been corrected for that.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{logicop.png}
+\caption{Region Based View Union}
+\label{fig:logic-op}
+\end{figure*}
+
+\begin{equation}
+\label{eqn:zin}
+z_{i}^{in}(u_{0}^{i},x,y,C) = \frac{|u_{0}^{i} - c_{+}^i|^2}{(max^{2}_{(x,y) \in u_{0}^{i}} u_{0}^i)^2}
+\end{equation}
+
+\begin{equation}
+\label{eqn:zout}
+z_{i}^{out}(u_{0}^{i},x,y,C) = \frac{|u_{0}^{i} - c_{-}^i|^2}{(max_{(x,y) \in u_{0}^{i}} u_{0}^i)^2}
+\end{equation}
+
+In order to perform logic operations on $z_{in}$ and $z_{out}$, the authors introduce interpolation functions that mimic the behavior of the regular truth
+table, but for continuous values between 0 and 1. The union and intersection functions for two variables are shown in Equations~\ref{eqn:funion}
+and~\ref{eqn:finters} respectively. These equations can be simply extended to any number of variables.
+
+\begin{equation}
+\label{eqn:funion}
+f_{\cup} = (z_{1} . z_{2})^{1/2}
+\end{equation}
+
+\begin{equation}
+\label{eqn:finters}
+f_{\cup} = 1 = ((1 - z_{1}) .(1-  z_{2}))^{1/2}
+\end{equation}
+
+Based on these functions, the Euler-Lagrange equation for any logic operation is shown in Equation~\ref{eqn:svpde}. According to the desired logic operation,
+$f_{in}$ and $f_{out}$ will be specified accordingly. For example, if we are doing the union of the images, then we will need to perform a union operation on
+the insides. Thus, $f_{in}$ will be replaced by Equation~\ref{eqn:funion}. Similarly, we will need to do the intersection of the outsides so $f_{out}$ will be
+replaced by Equation~\ref{eqn:finters}.
+
+\begin{equation}
+\label{eqn:svpde}
+\frac{\partial{\phi}}{\partial{t}} = \delta(\phi)[\mu\bigtriangledown(\frac{\bigtriangledown \phi}{|\bigtriangledown \phi|})  - \lambda(f_{in}(z_{1}^{in}, ...,
+z_{n}^{in}) - f_{out}(z_{1}^{out}, ..., z_{n}^{out})
+\end{equation}
+
+with the boundary condition
+
+$\frac{\delta(\phi)}{|\bigtriangledown\phi|}\frac{\partial\phi}{\partial n^{\rightarrow}} = 0$
+
 \section{Implementation}
@@
 
-\subsection{Evaluation Criteria}
-\label{sec:criteria}
-
 In order to make sure we correctly implemented the models, we had several evaluation criteria. Table~\ref{tab:criteria} summarizes these criteria, and explains
-the goals of each. For the rest of this section, we proceed by presenting our experimental results in the order of these criteria starting with the simplest
+the goals of each. Some of these criteria apply for both models, while other apply to one or the other. For the rest of this section, we proceed by presenting
+our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
 cases, and incrementally challenging the model.
@@
 \centering
-\begin{tabular}{ c | c}
-\textbf{Criteria} & \textbf{Goal} \\
+\begin{tabular}{ c | c |c}
+\textbf{Criteria} & \textbf{Goal} & \textbf{Applies to Model}\\
 \hline\hline
-Detecting Boundaries & Ability to correctly detect object boundaries of simple objects\\
+Detecting Boundaries & Ability to correctly detect object boundaries of simple objects& Both Models\\
 \hline
-Curve Position & Ability to correctly detect object boundaries irrespective of the initial curve position\\
+Curve Position & Ability to correctly detect object boundaries irrespective of the initial curve position & Both Models\\
 \hline
-Detecting Holes & Ability to detect holes in objects, and not simply stop on outside boundary\\
+Detecting Holes & Ability to detect holes in objects, and not simply stop on outside boundary & Both Models\\
 \hline
-Blurred Images & Ability to correctly (as much as possible) detect object boundaries in blurred images\\
+Blurred Images & Ability to correctly (as much as possible) detect object boundaries in blurred images & Both Models\\
 \hline
-Noisy Images & Ability to correctly (as much as possible) detect object boundaries in noisy images\\
+Noisy Images & Ability to correctly (as much as possible) detect object boundaries in noisy images & Both Models\\
 \hline
-Parameter Settings & Ability to respond correctly to the different parameter settings
+Union Operator & Ability to correctly obtain the union of two or more images & Sandberg-Chan\\
+\hline
+Intersection Operator & Ability to correctly obtain the intersection of two or more images & Sandberg-Chan\\
+\hline
+Complement & Ability to correctly obtain the union or intersection of two or more images containing complements & Sandberg-Chan\\
+\hline
+Parameter Settings & Ability to respond correctly to the different parameter settings & Both Models\\
 \end{tabular}
@@
 
-\subsection{Criteria 1: Successfully Detecting Boundaries}
+\subsection{Chan-Vese Model Results}
+
+\subsubsection*{Successful Detection of Boundaries}
+
+\subsubsection*{Independence of Initial Curve Position}
+
+\subsubsection*{Successful Detection of Holes}
+
+\subsubsection*{Reasonable Performance with Blurred Images}
+
+\subsubsection*{Reasonable Performance with Noisy Images}
+
+\subsubsection*{Successful Response to Parameter Settings}
+
+\subsubsection*{Other Tests}
+
+\subsection{Sandberg-Chan Model Results}
+
+\subsubsection*{Successful Detection of Boundaries}
+
+\subsubsection*{Independence of Initial Curve Position}
+
+\subsubsection*{Successful Detection of Holes}
+
+\subsubsection*{Reasonable Performance with Blurred Images}
+
+\subsubsection*{Reasonable Performance with Noisy Images}
+
+\subsubsection*{Union}
+
+\subsubsection*{Intersection}
+
+\subsubsection*{Complement}
+
+\subsubsection*{Successful Response to Parameter Settings}
+
+\subsubsection*{Other Tests}
 

### 2010-12-02 trunk/cs870/project/InitialPhi/cone.m
cleaned up cone

@@ -1,17 +1,24 @@
-%Sarah Nadi 20290927
+% cone: create a level set function phi for a cone. Therefore, when we 
+% visualize the zero level of phi it will give us a circle.
+% 
+% Output parameters:
+%   phi = the level set function for a cirlce given in matrix form
+% 
+% Input parameters:
+%   radius = the radius of the circle (i.e. the zero level set of phi).
+%   center = [x y] vector for the center of the circle.
+%   imagedDims = the grid that will be used for approximation.
 
-%this constructs a cone with the equation given in the paper  to make sure
-%that inside the cirlce is positive and outside is negative
-function phi = cone(radius, center, m)
-h = 1 / (m - 1);
+function phi = cone(radius, center, imageDims)
 
-phi = zeros(m,m);
-for i = 1 : m 
-    for j = 1 : m
-        phi(i,j) = -1*(sqrt( ((j-1) - center(1))^2 + ((i-1) - center(2))^2)) + radius;
-    end
-end
-end
+% Since we're allowed to analytically initialize phi, we can use the equation 
+% given in class to fill out the matrix.
+% General formula => phi = sqrt((x - centerX)^2 + (y - centerY)^2) - radius
+xy = cell(2, 1);
+xy{1} = 1 : 1 : imageDims(1);
+xy{2} = 1 : 1 : imageDims(2);
 
+axes = cell(2, 1);
+[ axes{:} ] = ndgrid(xy);
 
-
+phi = sqrt( (axes{1} - center(1)).^2 + (axes{2} - center(2)).^2 ) - radius;

### 2010-12-03 trunk/cs870/project/Logic/fin.m
Logic cleanup

@@ -1,14 +1,20 @@
-function output = fin(phi,logicop, varargin)
+% fin: calculate the energy inside the contour
+% 
+% Output parameters:
+%   output = the energy inside the contour
+% 
+% Input parameters:
+%   phi = current phi
+%   logicop = logic operator
+%   varargin = input channels
 
-%this function determines which logic operation will be applied to the
-%inside region
+function output = fin(phi, logicop, varargin)
 
+% If the logic operator is 'union', then the energy inside the contour is given
+% by the intersection. The opposite is true.
 switch logicop
     case 'union'
-        output = scunion(phi,'inside', varargin{:});               
+        output = scunion(phi, 'inside', varargin{:});               
     case 'intersection'
         output = scintersection(phi, 'inside', varargin{:});
 end
-
-end
-

### 2010-12-03 trunk/cs870/project/Logic/fout.m

@@ -1,14 +1,20 @@
+% fout: calculate the energy outside the contour
+% 
+% Output parameters:
+%   output = the energy outside the contour
+% 
+% Input parameters:
+%   phi = current phi
+%   logicop = logic operator
+%   varargin = input channels
+
 function output = fout(phi, logicop, varargin)
 
-%this function determines which logic operation will be applied to the
-%outside region
-
+% If the logic operator is 'union', then the energy outside the contour is given
+% by the intersection. The opposite is true.
 switch logicop
     case 'union'
-        output = scintersection(phi,'outside', varargin{:});
+        output = scintersection(phi, 'outside', varargin{:});
     case 'intersection'
-        output = scunion(phi,'outside', varargin{:});        
+        output = scunion(phi, 'outside', varargin{:});        
 end
-
-end
-

### 2010-12-03 trunk/cs870/project/Logic/zin.m

@@ -1,16 +1,18 @@
+% zin: compute value of the logic variable for the inside of the curve
+% 
+% Output parameters:
+%   output = the logic value for the inside of the curve
+% 
+% Input parameters:
+%   image = the original image
+%   phi = current phi
+
 function output = zin(image, phi)
 
-
 cin = mean(image(phi > 0));
-
 
 numerator = (image - cin).^2;
 
 denomenator = max(max(image))^2;
 
-output = double(numerator)/double(denomenator);
-
-
-
-end
-
+output = double(numerator) / double(denomenator);

### 2010-12-03 cs870/project/Logic/complement.m
Logic cleanup

@@ -1,6 +1,12 @@
+% complement: calculate the complement of a logic variable
+% 
+% Output parameters:
+%   output = complement of the input
+% 
+% Input parameters:
+%   input = logic variable
+
 function output = complement(input)
 
 output = 1 - input;
-
-end
 

### 2010-12-03 cs870/project/Logic/scintersection.m

@@ -1,6 +1,14 @@
+% scintersection: the objective function for the intersection logic operator
+% 
+% Output parameters:
+%   output = the logic intersection of multiple channels
+% 
+% Input parameters:
+%   phi = current phi
+%   region = 'inside' or 'outside'
+%   varargin = input channels
+
 function output = scintersection(phi, region, varargin)
-
-
 
 output = ones(size(phi,1), size(phi,2));
 switch region
@@ -15,8 +23,4 @@ switch region
 end
 
 output = real(output);
-
 output = 1 - output.^(1/size(varargin,2));
-
-end
-

### 2010-12-03 cs870/project/Logic/scunion.m

@@ -1,5 +1,14 @@
-function output = scunion(phi, region, varargin)
+% scunion: the objective function for the union logic operator
+% 
+% Output parameters:
+%   output = the logic union of multiple channels
+% 
+% Input parameters:
+%   phi = current phi
+%   region = 'inside' or 'outside'
+%   varargin = input channels
 
+function output = scunion(phi, region, varargin)
 
 output = ones(size(phi,1), size(phi,2));
 switch(region)
@@ -16,7 +25,3 @@ end
 output = real(output);
 
 output = output.^(1/size(varargin,2));
-
-
-end
-

### 2010-12-03 cs870/project/Logic/zout.m

@@ -1,13 +1,18 @@
+% zout: compute value of the logic variable for the outside of the curve
+% 
+% Output parameters:
+%   output = the logic value for the outside of the curve
+% 
+% Input parameters:
+%   image = the original image
+%   phi = current phi
+
 function output = zout(image, phi)
 
 cout = mean(image(phi < 0));
 
 numerator = (image - cout).^2;
 
-denomenator = max(max((image)))^2;
+denomenator = max(max(image))^2;
 
-output = double(numerator)/double(denomenator);
-
-
-end
-
+output = double(numerator) / double(denomenator);

### 2010-12-03 trunk/cs870/project/CurveEvolution/test1.m
test1 : Chan Vese

@@ -1,3 +1,12 @@
+% test: testing
+% 
+% Output parameters:
+%   output = the logic value for the inside of the curve
+% 
+% Input parameters:
+%   image = the original image
+%   phi = current phi
+
 function test(iterations)
 
 plotStep = 1;               

### 2010-12-03 cs870/project/CurveEvolution/evolveChannels.m
test1: Chane-Vese model

@@ -27,7 +27,7 @@ subPlotRows = 2;
 figure();
 for n=1:size(varargin, 2)
     subplot(subPlotRows, subPlotColumns, n);
-    imshow(varargin{n});
+    imshow(varargin{n}, 'initialmagnification','fit','displayrange',[0 255]);
     title(strcat('Input Channel ', num2str(n)));
 end
 
@@ -66,7 +66,9 @@ for n=1:iterations
     phi_new = updatePhiChannels(phi, deltaT, mu, lambda, logicop, varargin{:});
     
     % Reinitialization is optional
-    phi_new = vif(doReinit, reinit(phi_new), phi_new);
+    if(doReinit)
+        phi_new = reinit(phi_new);
+    end
     
     % Is the solution stationary ?
     if(stop(image, phi, phi_new))

### 2010-12-03 cs870/project/Helper/readGSImage.m

@@ -1,0 +1,13 @@
+% readGSImage: read any image as a grayscale double image
+%
+% Output parameters:
+%   image = the image as a grayscale double matrix
+%
+% Input parameters:
+%   fileName = the file name of the image, it's assumed all images are in
+%   the data folder
+
+function image = readGSImage(fileName)
+
+image = rgb2gray(imread(strcat('data/', fileName)));
+image = double(image);

### 2010-12-03 cs870/project/Helper/readRGBImage.m

@@ -1,0 +1,18 @@
+% readRGBImage: read any image as 3 components (R, G, B)
+%
+% Output parameters:
+%   [r g b] = the r, g, and b components of the image
+%
+% Input parameters:
+%   fileName = the file name of the image, it's assumed all images are in
+%   the data folder
+
+function [r g b] = readRGBImage(fileName)
+
+[image] = imread(strcat('data/', fileName));
+image = double(image);
+
+% Separate the components
+r = image(:, :, 1);
+g = image(:, :, 2);
+b = image(:, :, 3);

### 2010-12-03 cs870/project/InitialPhi/cone.m

@@ -7,7 +7,7 @@
 % Input parameters:
 %   radius = the radius of the circle (i.e. the zero level set of phi).
 %   center = [x y] vector for the center of the circle.
-%   imagedDims = the grid that will be used for approximation.
+%   imagedDims = a vector of the dimensions of the image.
 
 function phi = cone(radius, center, imageDims)
 
@@ -19,6 +19,6 @@ xy{1} = 1 : 1 : imageDims(1);
 xy{2} = 1 : 1 : imageDims(2);
 
 axes = cell(2, 1);
-[ axes{:} ] = ndgrid(xy);
+[ axes{:} ] = ndgrid(xy{:});
 
 phi = sqrt( (axes{1} - center(1)).^2 + (axes{2} - center(2)).^2 ) - radius;

### 2010-12-03 trunk/cs870/project/CurveEvolution/evolveCurve.m
test1: Chane-Vese model

@@ -7,8 +7,9 @@
 %   mu = weight for the length of the curve
 %   nu = weight for the area of the curve
 %   lambda = weight for the inside and outside energies
+%   doReinit = reinitializing phi is optional
 
-function evolveCurve(iterations, image, phi0, mu, nu, lambda)
+function evolveCurve(iterations, image, phi0, mu, nu, lambda, doReinit)
 
 %---------------------------------------------------------------------------
 % Initializing some parameters
@@ -19,7 +20,7 @@ imageSize = size(image,1);
 % Plot the input image
 figure();
 subplot(2, 2, 1);
-imshow(image);
+imshow(image, 'initialmagnification','fit','displayrange',[0 255]);
 title('Input Image');
 
 % Plot the initial contour
@@ -55,6 +56,11 @@ for n=1:iterations
     
     % Update phi
     phi_new = updatePhi(phi, image, deltaT, mu, nu, lambda);
+    
+    % Reinitialization is optional
+    if(doReinit)
+        phi_new = reinit(phi_new);
+    end
     
     % Is the solution stationary ?
     if(stop(image, phi, phi_new))

### 2010-12-03 trunk/cs870/project/CurveEvolution/stop.m

@@ -14,4 +14,8 @@ function indicator = stop(image, old, new)
 % introduces new points inside the curve or not.
 error = size(image(old >= 0), 1) - size(image(new >=0), 1);
 
-indicator = vif(error == 0, 1, 0);
+if(error == 0)
+    indicator = true;
+else
+    indicator = false;
+end

### 2010-12-03 trunk/cs870/project/CurveEvolution/test1.m

@@ -1,78 +1,20 @@
-% test: testing
-% 
-% Output parameters:
-%   output = the logic value for the inside of the curve
-% 
-% Input parameters:
-%   image = the original image
-%   phi = current phi
+% test1: testing the implementation of the Chan-Vese model
 
-function test(iterations)
+function test1()
 
-plotStep = 1;               
-t0 = 0;                      % Start at time t = 0
+iterations = 100;
 
-image1 = rgb2gray(imread('data/donuttopright.jpg'));
-image2 = rgb2gray(imread('data/donutbottomleft.jpg'));
-image3 = rgb2gray(imread('data/donutbottomright.jpg'));
+image = readGSImage('brain.jpg');
+imageSize = size(image, 1);
 
-objective = rgb2gray(imread('data/donut.jpg'));
+phi0 = cone(imageSize/4, [imageSize/2 imageSize/2], size(image));
 
-display('in evolve channels');
-imageSize = size(image1,1); % either image should work since they are the same size
+mu = 0.01;
 
+nu = 0;
 
-phi0 =   cone(80, [100 100], imageSize);
+lambda = 1;
 
-figure();
-numOfSubplots = 3;
-subplot(2,numOfSubplots,1); imshow(image1); title('Input Image 1');
-subplot(2,numOfSubplots,2); imshow(image2); title('Input Image 2');
-subplot(2,numOfSubplots,3); contour(flipud(phi0), [0 0], 'r','LineWidth',1); title('initial contour');
+doReinit = false;
 
-image1 = double(image1);
-image2 = double(image2);
-image3 = double(image3);
-
-
-% Initialize phi with phi0
-phi = phi0;
-
-%---------------------------------------------------------------------------
-% Evolve the curve
-t = t0;
-deltaT = (1/(imageSize))^2;
-
-        display('starting iterations');
-for n=1:iterations
-    if(mod(n, plotStep) == 0) 
-        subplot(2,numOfSubplots,4);
-        imshow(objective, 'initialmagnification','fit','displayrange',[0 255]);     
-        hold on;
-        contour(phi, [0 0], 'r','LineWidth',1);      title(strcat('iteration: ',num2str(n)));
-        drawnow;
-        seg = phi>=0;
-        subplot(2,numOfSubplots,5); imshow(seg); title('Segmentation');            
-    end
-    lambda = 255*255;
[…]

### 2010-12-03 cs870/project/CurveEvolution/evolveChannels.m
Tests

@@ -58,7 +58,7 @@ for n=1:iterations
     
     % Plot the segmentation
     seg = phi >= 0;
-    subplot(subPlotRows, subPlotColumns, subPlotColumns + 2);
+    subplot(subPlotRows, subPlotColumns, subPlotRows * subPlotColumns);
     imshow(seg);
     title(strcat('Segmentation at iteration: ', num2str(n)));
     
@@ -82,7 +82,7 @@ end
 
 % Display the final segmentation
 seg = phi >= 0;
-subplot(subPlotRows, subPlotColumns, subPlotColumns + 2);
+subplot(subPlotRows, subPlotColumns, subPlotRows * subPlotColumns);
 imshow(seg);
 title(strcat('Final segmentation at iteration: ', num2str(n)));
 

### 2010-12-03 cs870/project/Tests/test2.m
test2

@@ -1,0 +1,24 @@
+% test2: testing the implementation of the Sandberg-Chan model
+
+function test2()
+
+iterations = 100;
+
+image = readGSImage('image.jpg');
+imageSize = size(image, 1);
+
+image1 = readGSImage('image1.jpg');
+image2 = readGSImage('image2.jpg');
+image3 = readGSImage('image3.jpg');
+
+phi0 = cone(imageSize/4, [imageSize/2 imageSize/2], size(image));
+
+mu = 0.01;
+
+lambda = 255*255;
+
+logicop = 'intersection';
+
+doReinit = false;
+
+evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, image1, image2, image3);

### 2010-12-03 cs870/project/Logic/scintersection.m
implementation is done el-7amdolelah :D !

@@ -11,6 +11,7 @@
 function output = scintersection(phi, region, varargin)
 
 output = ones(size(phi,1), size(phi,2));
+
 switch region
     case 'inside'
         for i = 1 : size(varargin, 2)

### 2010-12-03 cs870/project/Logic/scunion.m

@@ -11,6 +11,7 @@
 function output = scunion(phi, region, varargin)
 
 output = ones(size(phi,1), size(phi,2));
+
 switch(region)
     case 'inside'
         for i = 1 : size(varargin, 2)

### 2010-12-03 cs870/project/Tests/test1.m

@@ -4,7 +4,7 @@ function test1()
 
 iterations = 100;
 
-image = readGSImage('brain.jpg');
+image = readGSImage('flowers.jpg');
 imageSize = size(image, 1);
 
 phi0 = cone(imageSize/4, [imageSize/2 imageSize/2], size(image));

### 2010-12-03 trunk/cs870/project/InitialPhi/cone.m
implementation is done el-7amdolelah :D !

@@ -21,4 +21,6 @@ xy{2} = 1 : 1 : imageDims(2);
 axes = cell(2, 1);
 [ axes{:} ] = ndgrid(xy{:});
 
-phi = sqrt( (axes{1} - center(1)).^2 + (axes{2} - center(2)).^2 ) - radius;
+% The equation is flipped because phi should be positive inside, and negative
+% outside.
+phi = radius - sqrt( (axes{1} - center(1)).^2 + (axes{2} - center(2)).^2 );

### 2010-12-03 trunk/cs870/project/Logic/fin.m

@@ -12,9 +12,10 @@ function output = fin(phi, logicop, varargin)
 
 % If the logic operator is 'union', then the energy inside the contour is given
 % by the intersection. The opposite is true.
+
 switch logicop
-    case 'union'
+    case 'union'        
         output = scunion(phi, 'inside', varargin{:});               
-    case 'intersection'
+    case 'intersection'        
         output = scintersection(phi, 'inside', varargin{:});
 end

### 2010-12-03 trunk/cs870/project/Logic/fout.m

@@ -12,9 +12,10 @@ function output = fout(phi, logicop, varargin)
 
 % If the logic operator is 'union', then the energy outside the contour is given
 % by the intersection. The opposite is true.
+
 switch logicop
     case 'union'
         output = scintersection(phi, 'outside', varargin{:});
-    case 'intersection'
+    case 'intersection'        
         output = scunion(phi, 'outside', varargin{:});        
 end

### 2010-12-03 trunk/cs870/project/Tests/test2.m

@@ -4,12 +4,14 @@ function test2()
 
 iterations = 100;
 
-image = readGSImage('image.jpg');
+image = imread('flowers.jpg');
 imageSize = size(image, 1);
 
-image1 = readGSImage('image1.jpg');
-image2 = readGSImage('image2.jpg');
-image3 = readGSImage('image3.jpg');
+[r g b] = readRGBImage('flowers.jpg');
+
+%image1 = readGSImage('image1.jpg');
+%image2 = readGSImage('image2.jpg');
+%image3 = readGSImage('image3.jpg');
 
 phi0 = cone(imageSize/4, [imageSize/2 imageSize/2], size(image));
 
@@ -17,8 +19,8 @@ mu = 0.01;
 
 lambda = 255*255;
 
-logicop = 'intersection';
+logicop = 'union';
 
 doReinit = false;
 
-evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, image1, image2, image3);
+evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, r, g, b);

### 2010-12-03 trunk/cs870/project/Report/Paper.tex
first 4 examples in chan vese

@@
 
-\subsection{Evaluation Criteria}
-\label{sec:criteria}
-
 In order to make sure we correctly implemented the models, we had several evaluation criteria. Table~\ref{tab:criteria} summarizes these criteria, and explains
-the goals of each. For the rest of this section, we proceed by presenting our experimental results in the order of these criteria starting with the simplest
-cases, and incrementally challenging the model.
+the goals of each. Some of these criteria apply for both models, while other apply to one or the other. For the rest of this section, we proceed by presenting
+our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
+cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image.
 
@@
 
-\subsection{Criteria 1: Successfully Detecting Boundaries}
+\subsection{Chan-Vese Model Results}
+
+\subsubsection*{Successful Detection of Boundaries}
+
+Figure~\ref{fig:cv_eg1} shows that the implemented model can successfully detect object boundaries in a simple image. In this case, the evolving curve was
+successfully able to detect the boundaries of both objects. Figure~\ref{fig:cv_eg2} shows a slightly more complicated example where the evolving contour
+successfully detects the contour of the brain. Note that there is a tiny area to the bottom left of the brain that was also successfully detected.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg1.png}
+\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
+0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
+\label{fig:cv_eg1}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg2.png}
+\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2)} + 32.8$, $\mu =
+0.01$, no reinitialization, cpu = 1.95 s, took 7 iterations.}
+\label{fig:cv_eg2}
+\end{figure*}
+
+\subsubsection*{Independence of Initial Curve Position}
+
+To ensure that the position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the same brain image
+used in Figure~\ref{fig:cv_eg2}. In Figure~\ref{fig:cv_eg2}, the initial contour was completely overlapping the object. Accordingly, we tried two other
+positions. The first one is shown in Figure~\ref{fig:cv_eg3} where the initial contour only partially overlaps with the object. The second one is shown in
+Figure~\ref{fig:cv_eg4} where the initial contour does not overlap the object in any part. Visually, the obtained segmentation was the same with a slightly
+higher number of iterations and CPU time. In all three cases, the brain was correctly segmented. We compared the area segmented as the brain in each case to
+make sure the same segmentation was obtained, and in each case, we got the same area of 0.5268. Note, however, that when part of the curve lies outside the
+object, the definition of inside and outside changes since the curve becomes an open curve at one point. This should not make a difference as long as the
+contour lies correctly at the boundaries of the object.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg3.png}
+\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2)} + 20$, $\mu =
+0.01$, no reinitialization, cpu = 2.26 s, took 9 iterations.}
+\label{fig:cv_eg3}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg4.png}
+\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
+$\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
+\label{fig:cv_eg4}
+\end{figure*}
+
+\subsubsection*{Successful Detection of Holes}
+
+In order to make sure the contour does not simply stop at the outside boundary of an object and ignore any details inside such as holes, we first tried a
+simple donut example, and then tried two more complicated donut images.
+
+\subsubsection*{Reasonable Performance with Blurred Images}
+
+\subsubsection*{Reasonable Performance with Noisy Images}
+
+\subsubsection*{Successful Response to Parameter Settings}
+
+\subsubsection*{Other Tests}
+
+flowers here
+
+\subsection{Sandberg-Chan Model Results}
+
+\subsubsection*{Successful Detection of Boundaries}
+
+\subsubsection*{Independence of Initial Curve Position}
+
+\subsubsection*{Successful Detection of Holes}
+
+\subsubsection*{Reasonable Performance with Blurred Images}
+
+\subsubsection*{Reasonable Performance with Noisy Images}
+
+\subsubsection*{Union}
+
+\subsubsection*{Intersection}
+
+\subsubsection*{Complement}
+
+\subsubsection*{Successful Response to Parameter Settings}
+
+\subsubsection*{Other Tests}
 

### 2010-12-03 trunk/cs870/project/Helper/readGSImage.m
added parameter to add noise or not

@@ -7,7 +7,14 @@
 %   fileName = the file name of the image, it's assumed all images are in
 %   the data folder
 
-function image = readGSImage(fileName)
+function image = readGSImage(fileName, addNoise)
 
-image = rgb2gray(imread(strcat('data/', fileName)));
-image = double(image);
+if(addNoise)
+    image = imread(strcat('data/', fileName));
+    image = imnoise(image,'salt & pepper');    
+    image = rgb2gray(image);   
+else      
+    image = rgb2gray(imread(strcat('data/', fileName)));   
+end
+
+ image = double(image);  

### 2010-12-03 trunk/cs870/project/Report/Paper.tex
added more explanation to fgures

@@
 our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
-cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image.
+cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
+the images in the Chan-Vese model were run on a BLA COMPUTER and all the images in the Sandberg-Chan model were run on a BLA computer.
 
@@
 \caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
 $\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
@@
 
-In order to make sure the contour does not simply stop at the outside boundary of an object and ignore any details inside such as holes, we first tried a
-simple donut example, and then tried two more complicated donut images.
+In order to make sure the contour does not simply stop at the outside boundary of an object and ignore any details inside such as holes, we tried two donut
+images. Figure~\ref{fig:cv_eg5} shows how the curve is able to detect the hole in the donut. Additionally, it is able to detect the various sprinkles on it.
+We note that the bottom right boundary is not very exact, and this is because the lighting effect in the image makes the intensity of this area very close to
+that of the background This is a known drawback of the Chan-Vese model since it only divides the image into two regions, and requires a big variation in
+intensities for the curve to change. Another example showing this limitation is shown in Figure~\ref{fig:cv_eg7} where the flower parts with very low
+intensities are not detected. Figure~\ref{fig:cv_eg6} shows that multiple holes in an image can also be successfully detected.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg5.png}
+\caption{Successful detection of holes.  Top: the evolving curve (in red) over time where the first image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.5)^2 + (y - 65.5)^2)} + 32.8$,
+$\mu =0.01$, no reinitialization, cpu = 8.12 s, took 12 iterations.}
+\label{fig:cv_eg5}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg6.png}
+\caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 15.07 s, took 23 iterations.}
+\label{fig:cv_eg6}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg7.png}
+\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
+image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
+\label{fig:cv_eg7}
+\end{figure*}
 
@@
 \subsubsection*{Successful Response to Parameter Settings}
-
-\subsubsection*{Other Tests}
-
-flowers here
 

### 2010-12-09 cs870/project/Logic/fin.m
complement

@@ -6,16 +6,17 @@
 % Input parameters:
 %   phi = current phi
 %   logicop = logic operator
+%   complement = vector of all channels that we will deal with their complement
 %   varargin = input channels
 
-function output = fin(phi, logicop, varargin)
+function output = fin(phi, logicop, complement, varargin)
 
 % If the logic operator is 'union', then the energy inside the contour is given
 % by the intersection. The opposite is true.
 
 switch logicop
     case 'union'        
-        output = scunion(phi, 'inside', varargin{:});               
+        output = scunion(phi, 'inside', complement, varargin{:});               
     case 'intersection'        
-        output = scintersection(phi, 'inside', varargin{:});
+        output = scintersection(phi, 'inside', complement, varargin{:});
 end

### 2010-12-09 cs870/project/Logic/sccomplement.m

@@ -1,4 +1,4 @@
-% complement: calculate the complement of a logic variable
+% sccomplement: calculate the complement of a logic variable
 % 
 % Output parameters:
 %   output = complement of the input
@@ -6,7 +6,7 @@
 % Input parameters:
 %   input = logic variable
 
-function output = complement(input)
+function output = sccomplement(input)
 
 output = 1 - input;
 

### 2010-12-09 trunk/cs870/project/CurveEvolution/evolveChannels.m
complement

@@ -8,10 +8,11 @@
 %   logicop = the logical operator to be applied
 %   image = the original image
 %   doReinit = reinitializing phi is optional
+%   complement = vector of all channels that we will deal with their complement
 %   varargin = multiple input channels, each channel is a 2D matrix
 
 function evolveChannels(iterations, phi0, mu, lambda, logicop, ...
-                        image, doReinit, varargin)
+                        image, doReinit, complement, varargin)
 
 %---------------------------------------------------------------------------
 % Initializing some parameters
@@ -47,7 +48,7 @@ deltaT = (1 / (imageSize))^2;
 for n=1:iterations
     % Draw the original image
     subplot(subPlotRows, subPlotColumns, subPlotColumns + 1);
-    imshow(image, 'initialmagnification', 'fit', 'displayrange', [0 255]);
+    imshow(varargin{1}, 'initialmagnification', 'fit', 'displayrange', [0 255]);
     title('Segmentation');
     hold on;
 
@@ -63,7 +64,8 @@ for n=1:iterations
     title(strcat('Segmentation at iteration: ', num2str(n)));
     
     % Update phi
-    phi_new = updatePhiChannels(phi, deltaT, mu, lambda, logicop, varargin{:});
+    phi_new = updatePhiChannels(phi, deltaT, mu, lambda, logicop, ...
+                                complement, varargin{:});
     
     % Reinitialization is optional
     if(doReinit)
@@ -71,13 +73,15 @@ for n=1:iterations
     end
     
     % Is the solution stationary ?
-    if(stop(image, phi, phi_new))
+    if(stop(phi, phi_new))
         display(strcat('Stopped at iteration: ', num2str(n)));
         break;
     end
     
     % The new phi will be used as phi for the next iteration
     phi = phi_new;
+    
+    % pause;
 end
 
 % Display the final segmentation

### 2010-12-09 trunk/cs870/project/CurveEvolution/evolveCurve.m

@@ -63,7 +63,7 @@ for n=1:iterations
     end
     
     % Is the solution stationary ?
-    if(stop(image, phi, phi_new))
+    if(stop(phi, phi_new))
         display(strcat('Stopped at iteration: ', num2str(n)));
         break;
     end

### 2010-12-09 trunk/cs870/project/CurveEvolution/stop.m

@@ -8,14 +8,8 @@
 %   old = old phi
 %   new = new phi
 
-function indicator = stop(image, old, new)
+function indicator = stop(old, new)
 
 % Basically, what we're trying to do here is to detect if the new phi
 % introduces new points inside the curve or not.
-error = size(image(old >= 0), 1) - size(image(new >=0), 1);
-
-if(error == 0)
-    indicator = true;
-else
-    indicator = false;
-end
+indicator = isequal(old > 0, new > 0);

### 2010-12-09 trunk/cs870/project/CurveEvolution/updatePhiChannels.m

@@ -9,10 +9,11 @@
 %   mu = scaling term for the length of the contour
 %   lambda = scaling term for the forces
 %   logicop = the required logical operator
+%   complement = vector of all channels that we will deal with their complement
 %   varargin = the number of channels to operate on
 
 function resultingPhi = updatePhiChannels(phi, deltaT, mu, lambda, ...
-                                            logicop, varargin)
+                                            logicop, complement, varargin)
 
 % |gradient(phi)| is used instead of the dirac delta function. This was an
 % option stated in the paper.
@@ -22,8 +23,8 @@ normOfGradient = norm(gradient(phi));
 curvature = kappa(phi);
 
 % Calculate the forces
-forceInside = fin(phi, logicop, varargin{:});
-forceOutside = fout(phi, logicop, varargin{:});
+forceInside = fin(phi, logicop, complement, varargin{:});
+forceOutside = fout(phi, logicop, complement, varargin{:});
 
 % Calculate phi_t according to the PDE in the paper
 phi_t =  normOfGradient * ...

### 2010-12-09 trunk/cs870/project/Logic/fout.m

@@ -6,16 +6,17 @@
 % Input parameters:
 %   phi = current phi
 %   logicop = logic operator
+%   complement = vector of all channels that we will deal with their complement
 %   varargin = input channels
 
-function output = fout(phi, logicop, varargin)
+function output = fout(phi, logicop, complement, varargin)
 
 % If the logic operator is 'union', then the energy outside the contour is given
 % by the intersection. The opposite is true.
 
 switch logicop
     case 'union'
-        output = scintersection(phi, 'outside', varargin{:});
+        output = scintersection(phi, 'outside', complement, varargin{:});
     case 'intersection'        
-        output = scunion(phi, 'outside', varargin{:});        
+        output = scunion(phi, 'outside', complement, varargin{:});
 end

### 2010-12-09 trunk/cs870/project/Logic/scintersection.m

@@ -6,20 +6,29 @@
 % Input parameters:
 %   phi = current phi
 %   region = 'inside' or 'outside'
+%   complement = vector of all channels that we will deal with their complement
 %   varargin = input channels
 
-function output = scintersection(phi, region, varargin)
+function output = scintersection(phi, region, complement, varargin)
 
 output = ones(size(phi,1), size(phi,2));
 
 switch region
     case 'inside'
         for i = 1 : size(varargin, 2)
-            output = (1 - zin(varargin{i}, phi)).*output;
+            if(size(find(complement == i) > 0))
+                output = (1 - sccomplement(zin(varargin{i}, phi))).*output;
+            else
+                output = (1 - zin(varargin{i}, phi)).*output;
+            end
         end
     case 'outside'
         for i = 1 : size(varargin, 2)
-            output = (1 - zout(varargin{i}, phi)).*output;
+            if(size(find(complement == i) > 0))
+                output = (1 - sccomplement(zout(varargin{i}, phi))).*output;
+            else
+                output = (1 - zout(varargin{i}, phi)).*output;
+            end
         end
 end
 

### 2010-12-09 trunk/cs870/project/Logic/scunion.m

@@ -6,20 +6,29 @@
 % Input parameters:
 %   phi = current phi
 %   region = 'inside' or 'outside'
+%   complement = vector of all channels that we will deal with their complement
 %   varargin = input channels
 
-function output = scunion(phi, region, varargin)
+function output = scunion(phi, region, complement, varargin)
 
 output = ones(size(phi,1), size(phi,2));
 
 switch(region)
     case 'inside'
         for i = 1 : size(varargin, 2)
-            output = zin(varargin{i}, phi).*output;
+            if(size(find(complement == i) > 0))
+                output = sccomplement(zin(varargin{i}, phi)).*output;
+            else
+                output = zin(varargin{i}, phi).*output;
+            end
         end       
     case 'outside'
         for i = 1 : size(varargin, 2)
-            output = zout(varargin{i}, phi).*output;
+            if(size(find(complement == i) > 0))
+                output = sccomplement(zout(varargin{i}, phi)).*output;
+            else
+                output = zout(varargin{i}, phi).*output;
+            end
         end        
 end
 

### 2010-12-09 trunk/cs870/project/Tests/test2.m
tests

@@ -2,16 +2,19 @@
 
 function test2()
 
-iterations = 100;
+startTime = cputime;
 
-image = imread('flowers.jpg');
+iterations = 10;
+
+% image = imread('flowers.jpg');
+image = readGSImage('image.jpg');
 imageSize = size(image, 1);
+% 
+% [r g b] = readRGBImage('flowers.jpg');
 
-[r g b] = readRGBImage('flowers.jpg');
-
-%image1 = readGSImage('image1.jpg');
-%image2 = readGSImage('image2.jpg');
-%image3 = readGSImage('image3.jpg');
+image1 = readGSImage('image1.jpg');
+image2 = readGSImage('image2.jpg');
+% image3 = readGSImage('image3.jpg');
 
 phi0 = cone(imageSize/4, [imageSize/2 imageSize/2], size(image));
 
@@ -19,8 +22,15 @@ mu = 0.01;
 
 lambda = 255*255;
 
-logicop = 'union';
+logicop = 'intersection';
 
 doReinit = false;
 
-evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, r, g, b);
+complement = [2];
+
+evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, complement, image1, image2);
+% evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, r, g, b);
+
+timeTaken = cputime - startTime;
+
+display(strcat('Time taken = ', num2str(timeTaken), 'seconds'));

### 2010-12-09 cs870/project/Tests/test1.m
tests

@@ -2,6 +2,8 @@
 
 function test1()
 
+startTime = cputime;
+
 iterations = 100;
 
 image = readGSImage('flowers.jpg');
@@ -18,3 +20,7 @@ lambda = 1;
 doReinit = false;
 
 evolveCurve(iterations, image, phi0, mu, nu, lambda, doReinit);
+
+timeTaken = cputime - startTime;
+
+display(strcat('Time taken = ', num2str(timeTaken), 'seconds'));

### 2010-12-12 cs870/project/CurveEvolution/evolveCurve.m
clean up - final revision probably isA !

@@ -32,7 +32,7 @@ title('Initial Contour');
 phi = phi0;
 
 % deltaT should satisfy the stability condition: deltaT <= 1/h^2
-deltaT = (1 / (imageSize))^2;    
+deltaT = (1 / (imageSize))^2;
     
 %---------------------------------------------------------------------------
 % Evolve the curve now!

### 2010-12-12 cs870/project/CurveEvolution/updatePhi.m

@@ -23,15 +23,13 @@ cout = mean(image(phi<0));
 normOfGradient = norm(gradient(phi));
 
 % Difference of intensities outside - inside
-force = lambda * (double((image - cout).^2) - double((image - cin).^2));
+force = lambda * ((image - cout).^2 - (image - cin).^2);
 
 % Calculate the curvature term in the PDE
-curvature = kappa(phi);
+curvature = mu * kappa(phi);
 
 % Calculate phi_t according to the PDE in the paper
-phi_t =  normOfGradient * ( mu * curvature - nu + force);
+phi_t =  normOfGradient * (curvature - nu + force);
 
 % Update phi
 resultingPhi = phi + deltaT * phi_t;
-
-

### 2010-12-12 trunk/cs870/project/CurveEvolution/evolveChannels.m
clean up - final revision probably isA !

@@ -48,7 +48,7 @@ deltaT = (1 / (imageSize))^2;
 for n=1:iterations
     % Draw the original image
     subplot(subPlotRows, subPlotColumns, subPlotColumns + 1);
-    imshow(varargin{1}, 'initialmagnification', 'fit', 'displayrange', [0 255]);
+    imshow(image, 'initialmagnification', 'fit', 'displayrange', [0 255]);
     title('Segmentation');
     hold on;
 
@@ -80,8 +80,6 @@ for n=1:iterations
     
     % The new phi will be used as phi for the next iteration
     phi = phi_new;
-    
-    % pause;
 end
 
 % Display the final segmentation

### 2010-12-12 trunk/cs870/project/CurveEvolution/reinit.m

@@ -8,5 +8,8 @@
 
 function signedPhi = reinit(phi)
 
+% dx
+dx = 1 / size(phi, 1);
+
 % Per our class notes.
-signedPhi = phi ./ sqrt(phi.^2 + 1);
+signedPhi = phi ./ sqrt(phi.^2 + dx^2);

### 2010-12-12 trunk/cs870/project/Tests/test1.m

@@ -6,12 +6,12 @@ startTime = cputime;
 
 iterations = 100;
 
-image = readGSImage('flowers.jpg');
+image = readGSImage('grouping3.jpg');
 imageSize = size(image, 1);
 
-phi0 = cone(imageSize/4, [imageSize/2 imageSize/2], size(image));
+phi0 = cone(floor(imageSize/4), [floor(imageSize/2) floor(imageSize/2)], size(image));
 
-mu = 0.01;
+mu = 0.1;
 
 nu = 0;
 

### 2010-12-12 trunk/cs870/project/Tests/test2.m

@@ -9,7 +9,7 @@ iterations = 10;
 % image = imread('flowers.jpg');
 image = readGSImage('image.jpg');
 imageSize = size(image, 1);
-% 
+
 % [r g b] = readRGBImage('flowers.jpg');
 
 image1 = readGSImage('image1.jpg');
@@ -26,7 +26,7 @@ logicop = 'intersection';
 
 doReinit = false;
 
-complement = [2];
+complement = 2;
 
 evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, complement, image1, image2);
 % evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, r, g, b);

### 2010-12-12 trunk/cs870/project/Report/Paper.tex
added more examples in part1.. last sec in part 1 left isA

@@
 
+During implementation, we tried to stick to the authors' formulas and guidelines as much as possible. In this section we explain the choices we made regardless
+implementation. For the chan-vese model, we used the PDE given in Equation 9 in the original paper~\cite{chan2001active} which is shown in
+Equation~\ref{eqn:cvpde}.  We, first, tried to implement the delta dirac function, $\delta_{\varepsilon}\phi$, as defined in the paper. However, we did not
+get non-zero values everywhere as indicated by the authors. Accordingly, we chose to use $|\bigtriangledown\phi|$ instead of the delta dirac function as this
+was indicated as a valid alternative by the authors. To calculate $c_{1}$ and $c_{2}$, we simply calculated the mean of the values inside $\phi$ (specifically
+where $\phi > 0$) and the mean of the values outside $\phi$ (specifically where $\phi < 0$) respectively. To actually solve the PDE, we chose to use an explicit
+time stepping scheme for the finite difference discretization. That is, in each time step $\phi_{t} = \phi + \triangle t*force$. This was simpler to implement,
+and although $\triangle t$ must be chosen carefully to satisfy the stability condition, we did not suffer from performance problems due to this restriction.
+
+For the sandberg-chan model, we used the PDE given in original paper as well, and shown in Equation~\ref{eqn:svpde}. Again, we used $|\bigtriangledown\phi|$
+instead of the dirac function. We calculated $z_{in}$ and $z_{out}$ according to Equations~\ref{eqn:zin, eqn:zout} shown above which have been corrected for
+the typo in the original paper. For all other equations and calculations necessary in the model, we closely followed the original paper in our implementation.
+We also used an explicit time stepping scheme in this model.
+
 \section{Experimental Results}
@@
 
+We present the results from the Chan-Vese model in this section, and show which criteria were met and which were not. Unless otherwise specified, we set
+$\lambda_{1} = \lambda_{2} = 1$.
+
 \subsubsection*{Successful Detection of Boundaries}
@@
 successfully able to detect the boundaries of both objects. Figure~\ref{fig:cv_eg2} shows a slightly more complicated example where the evolving contour
-successfully detects the contour of the brain. Note that there is a tiny area to the bottom left of the brain that was also successfully detected.
+successfully detects the contour of the brain. Note that there is a tiny area to the bottom left of the brain that was also successfully detected. To ensure
+that open boundaries (e.g. lines) can also be detected, we used an image with several shapes shown in Figure~\ref{fig:cv_eg8};
 
@@
 \label{fig:cv_eg2}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg8.png}
+\caption{Successful detection of object with open boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
+0.01$, no reinitialization, cpu = 5.19 s, took 11 iterations.}
+\label{fig:cv_eg8}
 \end{figure*}
@@
 intensities for the curve to change. Another example showing this limitation is shown in Figure~\ref{fig:cv_eg7} where the flower parts with very low
-intensities are not detected. Figure~\ref{fig:cv_eg6} shows that multiple holes in an image can also be successfully detected.
+intensities are not detected. We will show that we can overcome this limitation with the Sandberg-Chan model for this particular example.
+Figure~\ref{fig:cv_eg6} shows that multiple holes in an image can also be successfully detected.
 
@@
 
+Figure~\ref{fig:cv_eg9} shows how the contour behaves in a blurry image. The photographer was successfully detected. Additionally, other objects in the image
+such as the tripod were found. However, objects with a very light intensity were again not detected.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg9.png}
+\caption{Ability to detect contours in blurry images.  Top: the evolving curve (in red) over time where the first
+image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 255 x 255, $\phi_{0}(x,y) = - \sqrt{(x - 127.5)^2 + (y - 127.5)^2)} +
+63.75$, $\mu =0.01$, no reinitialization, cpu = 5.41 s, took 15 iterations.}
+\label{fig:cv_eg9}
+\end{figure*}
+
 \subsubsection*{Reasonable Performance with Noisy Images}
+
+Figure~\ref{fig:cv_eg10} shows how two objects in a noisy image were successfully detected. In this image, we added Gaussian noise with mean zero and 0.01
+variance using Matlab's built-in noise function. The image shows that although some of the noise was detected in intermediate iterations, the contour continued
+to evolve until it was only surrounding the desired objects. Unfortunately, the effect of noise was not completely ignored in all cases. For example,
+Figure~\ref{fig:cv_eg11} shows that for the same image, but with 'salt \& pepper' noise, the noise was detected as objects. This should not have been the case
+since we used a large value $\mu = 5$ for the length scaling parameter. However, for some reason, varying $\mu$ did not have the intended effect as explained
+in the next section.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg10.png}
+\caption{Ability to detect contours in an image with Gaussian noise.  Top: the evolving curve (in red) over time where the first
+image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
+75$, $\mu =0.01$, no reinitialization, cpu = 4.4 s, took 7 iterations.}
+\label{fig:cv_eg10}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg11.png}
+\caption{Same image as Figure~\ref{fig:cv_eg10}, but with 'salt \& pepper' noise. Noise was still detected despite increasing $\mu$.  Top: the evolving curve
+(in red) over time where the first
+image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
+75$, $\mu = 5$, no reinitialization, cpu = 12.27 s, took 7 iterations.}
+\label{fig:cv_eg11}
+\end{figure*}
 
@@
 
+There are mainly three parameters in this model that can be varied: $\mu, \lambda_{1}, \lambda_{2}$. We would usually want both $\lambda_{1}$ and $\lambda_{2}$
+to be equal to 1 to indicate that we care both about the difference of intensities inside and outside. A smaller weighting for one of them would mean that we
+would ignore some of the variances in intensities in this area. A larger weighting means that we are magnifying the difference in intensities in that area
+(i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we keep $\lambda_{2}$ constant, and vary
+$lambda_1$. When we set $\lambda_1 = 0.2$, it is shown that less details within the interior of the brain is detected. When we set $\lambda_2 = 2$, we can see
+that more details within the brain are detected.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg12.png}
+\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour.}
+\label{fig:cv_eg12}
+\end{figure*}
+
+$\lambda_1 = \lambda_2 = 1$
+$\lambda_1 = 2, \lambda_2 = 1$
+$\lambda_1 = 0.2, \lambda_2 = 1$
+
 \subsection{Sandberg-Chan Model Results}
+
+For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly
 

### 2010-12-12 cs870/project/Report/Paper.tex
part 1 experiments done elhamdolelah

@@
 our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
-cases, and incrementally challenging the model.
+cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
+the images in the Chan-Vese model were run on a BLA COMPUTER and all the images in the Sandberg-Chan model were run on a BLA computer.
 
@@
 
+We present the results from the Chan-Vese model in this section, and show which criteria were met and which were not. Unless otherwise specified, we set
+$\lambda_{1} = \lambda_{2} = 1$.
+
 \subsubsection*{Successful Detection of Boundaries}
+
+Figure~\ref{fig:cv_eg1} shows that the implemented model can successfully detect object boundaries in a simple image. In this case, the evolving curve was
+successfully able to detect the boundaries of both objects. Figure~\ref{fig:cv_eg2} shows a slightly more complicated example where the evolving contour
+successfully detects the contour of the brain. Note that there is a tiny area to the bottom left of the brain that was also successfully detected. To ensure
+that open boundaries (e.g. lines) can also be detected, we used an image with several shapes shown in Figure~\ref{fig:cv_eg8};
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg1.png}
+\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
+0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
+\label{fig:cv_eg1}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg2.png}
+\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2)} + 32.8$, $\mu =
+0.01$, no reinitialization, cpu = 1.95 s, took 7 iterations.}
+\label{fig:cv_eg2}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg8.png}
+\caption{Successful detection of object with open boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
+0.01$, no reinitialization, cpu = 5.19 s, took 11 iterations.}
+\label{fig:cv_eg8}
+\end{figure*}
 
@@
 
+To ensure that the position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the same brain image
+used in Figure~\ref{fig:cv_eg2}. In Figure~\ref{fig:cv_eg2}, the initial contour was completely overlapping the object. Accordingly, we tried two other
+positions. The first one is shown in Figure~\ref{fig:cv_eg3} where the initial contour only partially overlaps with the object. The second one is shown in
+Figure~\ref{fig:cv_eg4} where the initial contour does not overlap the object in any part. Visually, the obtained segmentation was the same with a slightly
+higher number of iterations and CPU time. In all three cases, the brain was correctly segmented. We compared the area segmented as the brain in each case to
+make sure the same segmentation was obtained, and in each case, we got the same area of 0.5268. Note, however, that when part of the curve lies outside the
+object, the definition of inside and outside changes since the curve becomes an open curve at one point. This should not make a difference as long as the
+contour lies correctly at the boundaries of the object.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg3.png}
+\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2)} + 20$, $\mu =
+0.01$, no reinitialization, cpu = 2.26 s, took 9 iterations.}
+\label{fig:cv_eg3}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg4.png}
+\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
+$\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
+\label{fig:cv_eg4}
+\end{figure*}
+
 \subsubsection*{Successful Detection of Holes}
+
+In order to make sure the contour does not simply stop at the outside boundary of an object and ignore any details inside such as holes, we tried two donut
+images. Figure~\ref{fig:cv_eg5} shows how the curve is able to detect the hole in the donut. Additionally, it is able to detect the various sprinkles on it.
+We note that the bottom right boundary is not very exact, and this is because the lighting effect in the image makes the intensity of this area very close to
+that of the background This is a known drawback of the Chan-Vese model since it only divides the image into two regions, and requires a big variation in
+intensities for the curve to change. Another example showing this limitation is shown in Figure~\ref{fig:cv_eg7} where the flower parts with very low
+intensities are not detected. We will show that we can overcome this limitation with the Sandberg-Chan model for this particular example.
+Figure~\ref{fig:cv_eg6} shows that multiple holes in an image can also be successfully detected.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg5.png}
+\caption{Successful detection of holes.  Top: the evolving curve (in red) over time where the first image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.5)^2 + (y - 65.5)^2)} + 32.8$,
+$\mu =0.01$, no reinitialization, cpu = 8.12 s, took 12 iterations.}
+\label{fig:cv_eg5}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg6.png}
+\caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 15.07 s, took 23 iterations.}
+\label{fig:cv_eg6}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg7.png}
+\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
+image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
+\label{fig:cv_eg7}
+\end{figure*}
 
@@
 
-\subsubsection*{Other Tests}
+There are mainly three parameters in this model that can be varied: $\mu, \lambda_{1}, \lambda_{2}$. We would usually want both $\lambda_{1}$ and $\lambda_{2}$
+to be equal to 1 to indicate that we care both about the difference of intensities inside and outside. A smaller weighting for one of them would mean that we
+would ignore some of the variances in intensities in this area. A larger weighting means that we are magnifying the difference in intensities in that area
+(i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we vary
+$lambda_1$ and $\lambda_2$. When we set $\lambda_1 = 0.2$ and $\lambda_2 = 1$, less details within the interior of the brain is detected. When
+we set $\lambda_1 = 2$ and $\lambda_2 = 1$, we can see that more details within the brain are detected. When we set $\lambda_1 = 5$, and $\lambda_2 = 0.01$ (a
+very extreme case), we can see that parts of outside boundaries are not well detected, while many details are detected within the brain.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg12.png}
+\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the segmentation in each case.}
+\label{fig:cv_eg12}
+\end{figure*}
+
+The second parameter, $\mu$ controls how much we allow the length of the curve to increase. If $\mu$ is small, it means that the curve length can increase to
+detect multiple smaller objects without penalizing our minimization problem. On the other hand if $\mu$ is large, it means that any change in the curve length
+will be scaled up, and thus will limit the curve expansion to keep the force to a minimum. Unfortunately, we were not able to see this effect in our
+experiments. We tried varying increasing $\mu$ to be able to detect groupings of objects instead of the individual objects, but the segmentation was invariant
+to $\mu$. We give more details about how we tried to handle this, and why we believe it is not working in Section~\ref{sec:difficulties}.
 
 \subsection{Sandberg-Chan Model Results}
+
+For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly
 

### 2010-12-12 trunk/cs870/project/Report/Paper.tex
updates style

@@
-\documentclass[10pt,journal,letterpaper,compsoc]{IEEEtran}
+\documentclass[10pt,twocolumn,letterpaper]{article}
+
+\usepackage{cvpr}
 \usepackage{cite}
 \usepackage{url}
-\usepackage{times}
 \usepackage{graphicx}
-\usepackage{url}
 \usepackage{clrscode}
@@
 \usepackage{balance}
+
+\cvprfinalcopy
+\ifcvprfinal\pagestyle{empty}\fi
 
@@
 
-% The paper headers
-\markboth{CS870 Project Report - Active Contour without Edges}%
-{Shell \MakeLowercase{\textit{et al.}}: Bare Demo of IEEEtran.cls for Computer Society Journals}
+\maketitle
 
-\IEEEcompsoctitleabstractindextext{%
 \begin{abstract}
@@
 \end{abstract}
-}
-
-\maketitle
 
@@
 \centering
+\resizebox{\textwidth }{!}{
 \begin{tabular}{ c | c |c}
@@
 \end{tabular}
+}
 \centering
@@
 would ignore some of the variances in intensities in this area. A larger weighting means that we are magnifying the difference in intensities in that area
-(i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we keep $\lambda_{2}$ constant, and vary
-$lambda_1$. When we set $\lambda_1 = 0.2$, it is shown that less details within the interior of the brain is detected. When we set $\lambda_2 = 2$, we can see
-that more details within the brain are detected.
+(i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we vary
+$lambda_1$ and $\lambda_2$. When we set $\lambda_1 = 0.2$ and $\lambda_2 = 1$, less details within the interior of the brain is detected. When
+we set $\lambda_1 = 2$ and $\lambda_2 = 1$, we can see that more details within the brain are detected. When we set $\lambda_1 = 5$, and $\lambda_2 = 0.01$ (a
+very extreme case), we can see that parts of outside boundaries are not well detected, while many details are detected within the brain.
 
@@
 \includegraphics[width=12cm]{cv_eg12.png}
-\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour.}
+\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the segmentation in each case.}
 \label{fig:cv_eg12}
@@
 
-$\lambda_1 = \lambda_2 = 1$
-$\lambda_1 = 2, \lambda_2 = 1$
-$\lambda_1 = 0.2, \lambda_2 = 1$
+The second parameter, $\mu$ controls how much we allow the length of the curve to increase. If $\mu$ is small, it means that the curve length can increase to
+detect multiple smaller objects without penalizing our minimization problem. On the other hand if $\mu$ is large, it means that any change in the curve length
+will be scaled up, and thus will limit the curve expansion to keep the force to a minimum. Unfortunately, we were not able to see this effect in our
+experiments. We tried varying increasing $\mu$ to be able to detect groupings of objects instead of the individual objects, but the segmentation was invariant
+to $\mu$. We give more details about how we tried to handle this, and why we believe it is not working in Section~\ref{sec:difficulties}.
 

### 2010-12-12 cs870/project/Tests/test1.m
tests

@@ -6,7 +6,7 @@ startTime = cputime;
 
 iterations = 100;
 
-image = readGSImage('grouping3.jpg');
+image = readGSImage('multiple.jpg', true);
 imageSize = size(image, 1);
 
 phi0 = cone(floor(imageSize/4), [floor(imageSize/2) floor(imageSize/2)], size(image));
@@ -15,7 +15,7 @@ mu = 0.1;
 
 nu = 0;
 
-lambda = 1;
+lambda = [1 1];
 
 doReinit = false;
 

### 2010-12-12 trunk/cs870/project/Tests/test2.m
tests

@@ -1,19 +1,18 @@
 % test2: testing the implementation of the Sandberg-Chan model
-
 function test2()
 
 startTime = cputime;
 
 iterations = 10;
 
-% image = imread('flowers.jpg');
-image = readGSImage('image.jpg');
+image = imread('flowers.jpg');
+% image = readGSImage('image.jpg');
 imageSize = size(image, 1);
 
-% [r g b] = readRGBImage('flowers.jpg');
+[r g b] = readRGBImage('flowers.jpg');
 
-image1 = readGSImage('image1.jpg');
-image2 = readGSImage('image2.jpg');
+% image1 = readGSImage('image1.jpg');
+% image2 = readGSImage('image2.jpg');
 % image3 = readGSImage('image3.jpg');
 
 phi0 = cone(imageSize/4, [imageSize/2 imageSize/2], size(image));
@@ -22,14 +21,14 @@ mu = 0.01;
 
 lambda = 255*255;
 
-logicop = 'intersection';
+logicop = 'union';
 
 doReinit = false;
 
-complement = 2;
+complement = 0;
 
-evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, complement, image1, image2);
-% evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, r, g, b);
+% evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, complement, image1, image2, image3);
+evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, complement, r, g, b);
 
 timeTaken = cputime - startTime;
 

### 2010-12-12 trunk/cs870/project/Helper/readGSImage.m
read image

@@ -6,12 +6,13 @@
 % Input parameters:
 %   fileName = the file name of the image, it's assumed all images are in
 %   the data folder
+%   addNoise = flag to add noise or not
 
 function image = readGSImage(fileName, addNoise)
 
 if(addNoise)
     image = imread(strcat('data/', fileName));
-    image = imnoise(image,'salt & pepper');    
+    image = imnoise(image,'gaussian');
     image = rgb2gray(image);   
 else      
     image = rgb2gray(imread(strcat('data/', fileName)));   

### 2010-12-12 trunk/cs870/project/CurveEvolution/evolveChannels.m
channels

@@ -80,6 +80,8 @@ for n=1:iterations
     
     % The new phi will be used as phi for the next iteration
     phi = phi_new;
+    
+    pause;
 end
 
 % Display the final segmentation

### 2010-12-12 trunk/cs870/project/CurveEvolution/updatePhi.m

@@ -23,7 +23,7 @@ cout = mean(image(phi<0));
 normOfGradient = norm(gradient(phi));
 
 % Difference of intensities outside - inside
-force = lambda * ((image - cout).^2 - (image - cin).^2);
+force =  lambda(2)*(image - cout).^2 - lambda(1)*(image - cin).^2;
 
 % Calculate the curvature term in the PDE
 curvature = mu * kappa(phi);

### 2010-12-12 trunk/cs870/project/CurveEvolution/evolveChannels.m
channels no pause

@@ -80,8 +80,6 @@ for n=1:iterations
     
     % The new phi will be used as phi for the next iteration
     phi = phi_new;
-    
-    pause;
 end
 
 % Display the final segmentation

### 2010-12-12 cs870/project/Report/Paper.tex
typo

@@
 \label{eqn:zin}
-z_{i}^{in}(u_{0}^{i},x,y,C) = \frac{|u_{0}^{i} - c_{+}^i|^2}{(max^{2}_{(x,y) \in u_{0}^{i}} u_{0}^i)^2}
+z_{i}^{in}(u_{0}^{i},x,y,C) = \frac{|u_{0}^{i} - c_{+}^i|^2}{(max_{(x,y) \in u_{0}^{i}} u_{0}^i)^2}
 \end{equation}
@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=7cm]{sc_unionsimple.png}
+\caption{}
+\label{fig:sc_unionsimple}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=7cm]{sc_intersectionsimple.png}
+\caption{}
+\label{fig:sc_intersectionsimple}
+\end{figure}
+
 Figure~\ref{fig:cv_eg10} shows how two objects in a noisy image were successfully detected. In this image, we added Gaussian noise with mean zero and 0.01
@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=7cm]{threech.png}
+\caption{}
+\label{fig:threech}
+\end{figure}
+
 \begin{figure*}[t]
@@
 
-For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly
+For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly included in the logic operations functionality. We therefore, show
+the other criteria.
 
-\subsubsection*{Successful Detection of Boundaries}
+\subsubsection*{Logic Operators: union and intersection}
+
+To ensure that the union operator is working properly, we tried it on a very simple example of two rectangles shown in Figure~\ref{fig:sc_unionsimple}. The
+union of the two rectangles was successfully detected. Similarly, Figure~\ref{fig:sc_intersectionsimple} shows the intersection for the same two simple
+rectangles.
+
+Figure~\ref{fig:donutlogic} shows a more complicated example for both the intersection and union with the donut image shown in the previous section. We used
+two versions of the donuts with each one having a different part occluded. The results of the union and intersection operators are shown in the figure.
+Additionally, this example shows that the Sandberg-Chan model can still successfully detect holes despite the addition of the logic operations.
+
+A practical usage of this logic operations framework is to recover missing parts from different images, and combine them to produce a more complete image. We
+show this in Figure~\ref{fig:family} which shows two different version of a family image with a child missing in each image. When the union is applied to these
+two channels, the complete family image is recovered.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{donutlogic.png}
+\caption{}
+\label{fig:donutlogic}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{flowers.png}
+\caption{}
+\label{fig:threech}
+\end{figure*}
 
@@
 
-\subsubsection*{Successful Detection of Holes}
+Similar to the previous model, we needed to ensure that the position of the initial curve does not alter our results. Figure~\ref{fig:sc_initialcurve} shows
+the segmentation results obtained for the same image used in Figure~\ref{fig:sc_unionsimple}, but with a different position for the initial curve.
 
@@
 \subsubsection*{Reasonable Performance with Noisy Images}
-
-\subsubsection*{Union}
-
-\subsubsection*{Intersection}
 

### 2010-12-13 cs870/project/Report/Paper.tex
all figure examples done but need revision

@@
 cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
-the images in the Chan-Vese model were run on a BLA COMPUTER and all the images in the Sandberg-Chan model were run on a BLA computer.
+the images in the Chan-Vese model were run on a BLA COMPUTER and all the images in the Sandberg-Chan model were run on a BLA computer. As shown in the results,
+we tried different image sizes to ensure that our solution converges.
 
@@
 \includegraphics[width=7cm]{sc_unionsimple.png}
-\caption{}
+\caption{Simple union example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.32 s, 
+iterations = 2.}
 \label{fig:sc_unionsimple}
@@
 \includegraphics[width=7cm]{sc_intersectionsimple.png}
-\caption{}
+\caption{Simple intersection example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.83 s, 
+iterations = 2.}
 \label{fig:sc_intersectionsimple}
@@
 \includegraphics[width=7cm]{threech.png}
-\caption{}
+\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 * 200. $\phi_{0}(x,y) = - \sqrt{(x -
+100)^2 + (y - 100)^2)} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
+5. For intersection then union: cpu = 1.24 and iterations = 4.}
 \label{fig:threech}
@@
 \includegraphics[width=12cm]{cv_eg12.png}
-\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the segmentation in each case.}
+\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
 \label{fig:cv_eg12}
@@
 For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly included in the logic operations functionality. We therefore, show
-the other criteria.
+the other criteria. Unless otherwise specified, $\mu = 0.1$ and $\lambda = 255*255$ in the experiments below.
 
@@
 \includegraphics[width=12cm]{donutlogic.png}
-\caption{}
+\caption{Performing union and intersection on images with holes. Size = 300 * 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
+75$, no reinitialization. For union: cpu = 8.63 sec and iterations = 10. For intersection: cpu = 7.68 sec and iterations = 10.}
 \label{fig:donutlogic}
@@
 \includegraphics[width=12cm]{flowers.png}
-\caption{}
-\label{fig:threech}
+\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
+Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
+size = cpu = 1.73 sec, iterations = 7.}
+\label{fig:flowers}
 \end{figure*}
@@
 
-Similar to the previous model, we needed to ensure that the position of the initial curve does not alter our results. Figure~\ref{fig:sc_initialcurve} shows
+Similar to the previous model, we needed to ensure that the position of the initial curve does not alter our results. Figure~\ref{fig:sc_position} shows
 the segmentation results obtained for the same image used in Figure~\ref{fig:sc_unionsimple}, but with a different position for the initial curve.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=7cm]{sc_position.png}
+\caption{Same example shown in Figure~\ref{fig:sc_unionsimple}, but with different initial curve position. Initial curve does not overlap with either objects.
+Same segmentation is obtained. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 4.43s, 4
+iterations.}
+\label{fig:sc_position}
+\end{figure}
 
@@
 
+Figure~\ref{fig:sc_blurry} shows the union performed on two blurry triangles. We blurred the triangles used in Figure~\ref{fig:threech}, and performed the
+union operation on them. The full triangle was segmented (the union) despite the blurred edges.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg11.png}
+\caption{Successful union with blurred images. Top: the evolving curve for $A_1 \cup A_2$
+(in red) over time where the first
+image shows the initial
+contour. Bottom: evolving segmentation over time until the union is detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
+50$, no reinitialization, cpu = 6 s, took 11 iterations.}
+\label{fig:sc_blurry}
+\end{figure*}
+
 \subsubsection*{Reasonable Performance with Noisy Images}
+
+To be able to perform well in noisy images, $\lambda$ must be decreased in order to ignore the noise. We added 'Salt \& pepper' noise with a variation of 0.1 to
+the images shown in Figure~\ref{fig:sc_noisy}, and tried to perform the union. In the first case shown, we set $\lambda = 255*255$ which is the normal setting
+used for clean images, and in the second case, we set $\lambda = 25$. We show the final segmentation for both cases. In the first case, the noise was detected,
+while in the second case, the noise was ignored since $\lambda$ was small. This shows that we also satisfy the response to parameter settings criteria since
+$\lambda$ was the only variation mentioned by the authors in the original paper~\cite{sandberg2005logic}. We note that the contour evolved much more slowly
+with a smalled $\lambda$, and thus took much more CPU time, and a larger number of iterations.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{sc_noisy.png}
+\caption{Same image as Figure~\ref{fig:threech}, but with 'salt \& pepper' noise with a variation of 0.1. Decreasing $\lambda$ from 255*255 to 125 allowed less
+noise to be detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
+50$, no reinitialization. Cpu = 3.17 sec and iterations = 5 when $\lambda = 255*255$. Cpu = 146.46 sec and iterations = 125 when $\lambda =
+125$.}
+\label{fig:sc_noisy}
+\end{figure*}
 
@@
 
-\subsubsection*{Successful Response to Parameter Settings}
+Unfortunately, the complement did not work properly in our implementation. We closely followed the definitions in the original paper where $z_i^{in'} = 1 -
+z_i^{in}$ and $z_i^{out'} = 1 - z_i^{out}$. The results was actually successfully detected, but continuous flickering occurred, and the curve did not stop in
+one position. We discuss this results in more details in Section~\ref{sec:disc}.
 
 \subsubsection*{Other Tests}
+
+Since the Sandberg-Chan model can successfully detect objects in one channel which are not in another when performing the union, we were curious to know
+whether this would allow us to perform successful segmentation on colored images by combining the information in the three channels R,G,B. Accordingly, we used
+the image in Figure~\ref{fig:cv_eg7} which the Chan-Vese model was not able to completely segment. We extracted three channels from the colored image: Red,
+Green, Black, and performed the union on the three channels. The results are shown in Figure~\ref{fig:flowers}. The complete contour of the flowers was
+successfully detected unlike that obtained in Figure~\ref{fig:cv_eg7}.
 

### 2010-12-13 trunk/cs870/project/Report/Paper.tex
trying to fix missing references

@@
 
-\title{An Implementation of the Active Contours without Edged model and the Logic Framework for Active contours on multi-channel images}
+\title{An Implementation of the Active Contours without Edges model and the Logic Framework for Active contours on multi-channel images}
 
@@
 \centering
-\includegraphics[width=12cm]{cv_eg1.png}
+\includegraphics[width=11cm]{cv_eg1.png}
 \caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
@@
 \centering
-\includegraphics[width=12cm]{cv_eg7.png}
+\includegraphics[width=10cm]{cv_eg7.png}
 \caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
@@
 \subsubsection*{Reasonable Performance with Noisy Images}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=6cm]{sc_unionsimple.png}
+\caption{Simple union example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.32 s, 
+iterations = 2.}
+\label{fig:sc_unionsimple}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=6cm]{sc_intersectionsimple.png}
+\caption{Simple intersection example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.83 s, 
+iterations = 2.}
+\label{fig:sc_intersectionsimple}
+\end{figure}
 
@@
 (i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we vary
-$lambda_1$ and $\lambda_2$. When we set $\lambda_1 = 0.2$ and $\lambda_2 = 1$, less details within the interior of the brain is detected. When
+$\lambda_1$ and $\lambda_2$. When we set $\lambda_1 = 0.2$ and $\lambda_2 = 1$, less details within the interior of the brain is detected. When
 we set $\lambda_1 = 2$ and $\lambda_2 = 1$, we can see that more details within the brain are detected. When we set $\lambda_1 = 5$, and $\lambda_2 = 0.01$ (a
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=12cm]{cv_eg12.png}
-\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the segmentation in each case.}
+\includegraphics[width=7cm]{threech.png}
+\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 * 200. $\phi_{0}(x,y) = - \sqrt{(x -
+100)^2 + (y - 100)^2)} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
+5. For intersection then union: cpu = 1.24 and iterations = 4.}
+\label{fig:threech}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=7cm]{cv_eg12.png}
+\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
 \label{fig:cv_eg12}
-\end{figure*}
+\end{figure}
 
@@
 
-For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly
+For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly included in the logic operations functionality. We therefore, show
+the other criteria. Unless otherwise specified, $\mu = 0.1$ and $\lambda = 255*255$ in the experiments below.
 
-\subsubsection*{Successful Detection of Boundaries}
+\subsubsection*{Logic Operators: union and intersection}
+
+To ensure that the union operator is working properly, we tried it on a very simple example of two rectangles shown in Figure~\ref{fig:sc_unionsimple}. The
+union of the two rectangles was successfully detected. Similarly, Figure~\ref{fig:sc_intersectionsimple} shows the intersection for the same two simple
+rectangles.
+
+Figure~\ref{fig:donutlogic} shows a more complicated example for both the intersection and union with the donut image shown in the previous section. We used
+two versions of the donuts with each one having a different part occluded. The results of the union and intersection operators are shown in the figure.
+Additionally, this example shows that the Sandberg-Chan model can still successfully detect holes despite the addition of the logic operations.
+
+A practical usage of this logic operations framework is to recover missing parts from different images, and combine them to produce a more complete image. We
+show this in Figure~\ref{fig:family} which shows two different version of a family image with a child missing in each image. When the union is applied to these
+two channels, the complete family image is recovered.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{donutlogic.png}
+\caption{Performing union and intersection on images with holes. Size = 300 * 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
+75$, no reinitialization. For union: cpu = 8.63 sec and iterations = 10. For intersection: cpu = 7.68 sec and iterations = 10.}
+\label{fig:donutlogic}
+\end{figure*}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{flowers.png}
+\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
+Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
+size = cpu = 1.73 sec, iterations = 7.}
+\label{fig:flowers}
+\end{figure*}
 
@@
 
-\subsubsection*{Successful Detection of Holes}
+Similar to the previous model, we needed to ensure that the position of the initial curve does not alter our results. Figure~\ref{fig:sc_position} shows
+the segmentation results obtained for the same image used in Figure~\ref{fig:sc_unionsimple}, but with a different position for the initial curve.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=7cm]{sc_position.png}
+\caption{Same example shown in Figure~\ref{fig:sc_unionsimple}, but with different initial curve position. Initial curve does not overlap with either objects.
+Same segmentation is obtained. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 4.43s, 4
+iterations.}
+\label{fig:sc_position}
+\end{figure}
 
@@
 
+Figure~\ref{fig:sc_blurry} shows the union performed on two blurry triangles. We blurred the triangles used in Figure~\ref{fig:threech}, and performed the
+union operation on them. The full triangle was segmented (the union) despite the blurred edges.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg11.png}
+\caption{Successful union with blurred images. Top: the evolving curve for $A_1 \cup A_2$
+(in red) over time where the first
+image shows the initial
+contour. Bottom: evolving segmentation over time until the union is detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
+50$, no reinitialization, cpu = 6 s, took 11 iterations.}
+\label{fig:sc_blurry}
+\end{figure*}
+
 \subsubsection*{Reasonable Performance with Noisy Images}
 
-\subsubsection*{Union}
+To be able to perform well in noisy images, $\lambda$ must be decreased in order to ignore the noise. We added 'Salt \& pepper' noise with a variation of 0.1 to
+the images shown in Figure~\ref{fig:sc_noisy}, and tried to perform the union. In the first case shown, we set $\lambda = 255*255$ which is the normal setting
+used for clean images, and in the second case, we set $\lambda = 25$. We show the final segmentation for both cases. In the first case, the noise was detected,
+while in the second case, the noise was ignored since $\lambda$ was small. This shows that we also satisfy the response to parameter settings criteria since
+$\lambda$ was the only variation mentioned by the authors in the original paper~\cite{sandberg2005logic}. We note that the contour evolved much more slowly
+with a smalled $\lambda$, and thus took much more CPU time, and a larger number of iterations.
 
-\subsubsection*{Intersection}
+\begin{figure}[t]
+\centering
+\includegraphics[width=7cm]{sc_noisy.png}
+\caption{Same image as Figure~\ref{fig:threech}, but with 'salt \& pepper' noise with a variation of 0.1. Decreasing $\lambda$ from 255*255 to 125 allowed less
+noise to be detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
+50$, no reinitialization. Cpu = 3.17 sec and iterations = 5 when $\lambda = 255*255$. Cpu = 146.46 sec and iterations = 125 when $\lambda =
+125$.}
+\label{fig:sc_noisy}
+\end{figure}
 
@@
 
-\subsubsection*{Successful Response to Parameter Settings}
+Unfortunately, the complement did not work properly in our implementation. We closely followed the definitions in the original paper where $z_i^{in'} = 1 -
+z_i^{in}$ and $z_i^{out'} = 1 - z_i^{out}$. The results was actually successfully detected, but continuous flickering occurred, and the curve did not stop in
+one position. We discuss this results in more details in Section~\ref{sec:difficulties}.
 
 \subsubsection*{Other Tests}
+
+Since the Sandberg-Chan model can successfully detect objects in one channel which are not in another when performing the union, we were curious to know
+whether this would allow us to perform successful segmentation on colored images by combining the information in the three channels R,G,B. Accordingly, we used
+the image in Figure~\ref{fig:cv_eg7} which the Chan-Vese model was not able to completely segment. We extracted three channels from the colored image: Red,
+Green, Black, and performed the union on the three channels. The results are shown in Figure~\ref{fig:flowers}. The complete contour of the flowers was
+successfully detected unlike that obtained in Figure~\ref{fig:cv_eg7}.
 

### 2010-12-13 trunk/cs870/project/Report/Paper.tex
conc only missing

@@
 \usepackage{graphicx}
-\usepackage{clrscode}
 \usepackage{tabularx}
@@
 \begin{abstract}
-%\boldmath
-The abstract goes here.
+In this project, we provide an implementation for the Chan-Vese model for active contours without edges, and the Sandberg-Chan logic framework for active
+contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It differs
+from other active contour modesl in that it does not depend on the gradient of the contour ($\phi$), therefore it is more capable of detecting objects whose
+boundaries may not be defined by a gradient (e.g. blurry images, noisy images). The Sandberg-Chan model builds on top of the Chan-Vese model by allowing the
+contour to evolve simultaneously on multiple channels. It also offers a set of logic operations that can be applied to those channels (union, intersection,
+complement, and their combinations). We evaluate our implementation based on several criteria, such as: independence of initial curve position, successful
+detection of holes, performance with blurred and noisy images, and response to scaling parameters.
 \end{abstract}
@@
 \label{eqn:finters}
-f_{\cup} = 1 = ((1 - z_{1}) .(1-  z_{2}))^{1/2}
+f_{\cap} = 1 = ((1 - z_{1}) .(1-  z_{2}))^{1/2}
 \end{equation}
@@
 \label{sec:difficulties}
+We have faced many challenges throughout the implementation of this project. Some of the challenges caused major delays in the implementation cycle and hindered
+us from improving our code and adding more features to it.
+
+\subsection{Varying $\mu$}
+$\mu$ is a scaling factor that is used in both the Chan-Vese model and the Sandberg-Chan model to control the weight given to the length of the contour.
+Therefore, varying the value of $\mu$ will affect the ability to segment objects in noisy images, especially when the noise is similar to Matlab's \textit{salt
+and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure [\ref{fig:cvmu}] shows the
+segmentation of an object in a noisy for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting that
+this challenge kept us busy for over two days, after which we decided we should move on to other parts of the project.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{cvmu.png}
+\caption{$\mu$ does not have any effect on the length of the contour. \textit{Salt and pepper} noise was added to the original image for testing purposes. The
+figure shows the segmentation for the same image for various values of $\mu$.}
+\label{fig:cvmu}
+\end{figure*}
+
+\subsection{Computing the complement}
+The Sandberg-Chan model supports the complement logic operator, so one can segment the logic object ($A \cap \neg B$) (i.e. segment what is in channel A and not
+in channel B). The model defines the complement of the object in channel \textit{i} as follows:
+
+\begin{equation}
+\label{eqn:complement}
+\begin{align*}
+{z_{i}}^{in^{'}} = 1 - {z_{i}}^{in}, \\
+{z_{i}}^{out^{'}} = 1 - {z_{i}}^{out}.
+\end{align*}
+\end{equation}
+
+Although the definition is very simple, it does not work in practice. We implemented the complement logic operator as defined in Equation
+[\ref{eqn:complement}], and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
+flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure [\ref{fig:sccomp}] shows this scenrio. We tried to
+debug the reasons for this problem for three days.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{sccomp.png}
+\caption{The complement operator causes the contour to go through an infinite loop between two segmentations: the correct segmentation, and the segmentation
+of $A \cap B$.}
+\label{fig:sccomp}
+\end{figure*}
+
+\subsection{Defining the initial contour ($\phi_0$)}
+Both the Chan-Vese model and the Sandberg-Chan model defines the initial contour $\phi_0$ as a signed distance function where $\phi_0$ is positive inside the
+contour $C$, zero on the boundary of $C$ (\partial C), and negative outside the contour $C$. Figure [\ref{fig:phinote}] illustrates those assumptions.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\textwidth]{phinote.png}
+\caption{Assumptions about the way $\phi_0$ is defined.}
+\label{fig:phinote}
+\end{figure}
+
+Since we model our initial contour as a circle, we have to modify the regular equation of the circle to match those assumption. Therefore, we initialize
+$\phi_0$ as follows:
+
+\begin{equation}
+\label{eqn:phinote}
+\phi_{0} (x, y) = radius - \sqrt{ (x - c_x)^2 + (y - c_y)^2 },
+\end{equation}
+where $c_x$ and $c_y$ are the x-coordinate and y-coordinate of the center of the circle respectively.
+
+Although this definition of $\phi_0$ is compatible with most of the images we used to test our implementation, it did fail to correctly segment the objects in
+some images. In fact, the results were the opposite of the used logic operator (i.e. the output is the intersection of the channels, while it should be the
+union). After spending quite sometime to resolve this problem and checking what might be the source of the error, we tried using the regular equation of a
+circle to initialize $\phi_0$ (Equation [\ref{eq:phinotenormal]}). We were very intrigued to notice that this modification resulted in the correct segmentation
+of the input channels. Figure [\ref{fig:phidef}] shows the final segmentation using both definitions for $\phi_0$.
+
+\begin{equation}
+\label{eqn:phinotenormal}
+\phi_{0} (x, y) = \sqrt{ (x - c_x)^2 + (y - c_y)^2 } - radius,
+\end{equation}
+where $c_x$ and $c_y$ are the x-coordinate and y-coordinate of the center of the circle respectively.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{phidef.png}
+\caption{}
+\label{fig:phidef}
+\end{figure*}
+
+\subsection{Mistyped Equations}
+As previously mentiond in Section [\ref{sec:sandberg-chan}], there is a typo in the equations that define $z_in$ and $z_out$ in the Sandberg-Chan model
+(equation 3 in thier paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
+typos consumed over a week of our time until we were able to find out what are the correct equations in both cases. We could have spent this time overcoming
+other challenges that we faced while implementing both models.
 
@@
 \label{sec:concl}
-The conclusion goes here.
 
 \bibliographystyle{IEEEtran}
-
 \bibliography{references}

### 2010-12-13 trunk/cs870/project/Report/Paper.tex
all there but figs not working

@@
 
+\begin{figure}[t]
+\centering
+\includegraphics[width=7cm]{family.png}
+\caption{Combining iterations = 7. union 4.01 s intersection 4.12 s}
+\label{fig:donutlogic}
+\end{figure}
+
 \begin{figure*}[t]
@@
 \includegraphics[width=\textwidth]{cvmu.png}
-\caption{$\mu$ does not have any effect on the length of the contour. \textit{Salt and pepper} noise was added to the original image for testing purposes. The
-figure shows the segmentation for the same image for various values of $\mu$.}
+\caption{$\mu$ does not have any effect on the length of the contour. \textit{Salt and pepper} noise was added to the original image for testing
+purposes. Top: original image with noise. Bottom: segmentation for various values of $\mu$.}
 \label{fig:cvmu}
@@
 \caption{The complement operator causes the contour to go through an infinite loop between two segmentations: the correct segmentation, and the segmentation
-of $A \cap B$.}
+of $A_1 \cap A_2$. Bottom: correct and incorrect segmentation respectively}
 \label{fig:sccomp}
@@
 Both the Chan-Vese model and the Sandberg-Chan model defines the initial contour $\phi_0$ as a signed distance function where $\phi_0$ is positive inside the
-contour $C$, zero on the boundary of $C$ (\partial C), and negative outside the contour $C$. Figure [\ref{fig:phinote}] illustrates those assumptions.
+contour $C$, zero on the boundary of $C$ ($\partial C$), and negative outside the contour $C$. Figure [\ref{fig:phinote}] illustrates those assumptions.
 

### 2010-12-13 cs870/project/Report/Paper.tex
need to fix errors in refs

@@
 \usepackage{cvpr}
-\usepackage{cite}
-\usepackage{url}
 \usepackage{graphicx}
-\usepackage{clrscode}
 \usepackage{tabularx}
 \usepackage{amsmath}
-\usepackage{array}
-\usepackage{color}
-\usepackage{balance}
+\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=true]{hyperref}
 
@@
 Image segmentation or boundary detection is a very important problem in the area of Image Processing, and has received a lot of attention in the past. The
-classical Active Contour model (or Snakes) proposed by Kass et al.~\cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
+classical Active Contour model (or Snakes) proposed by Kass et al. \cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
 a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it very highly adopted later on
@@
 and Vese~\cite{chan2001active} propose an active contour model that does not depend on the edges (i.e the gradient) for propagating the curve to detect the
-boundary of the object. Instead, they use a region-based approach based on the Mumford-Shah model~\cite{mumford1989optimal} to divide the image
+boundary of the object. Instead, they use a region-based approach based on the Mumford-Shah model \cite{mumford1989optimal} to divide the image
 into two regions: one inside the propagating curve, and one outside. The curve is at the boundary of the object if there is no difference in intensities inside
@@
 
-\begin{figure}[t]
+\begin{figure}[t!]
 \centering
-\includegraphics[width=7cm]{sc_unionsimple.png}
+\includegraphics[width=6cm]{sc_unionsimple.png}
 \caption{Simple union example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.32 s, 
@@
 
-\begin{figure}[t]
+\begin{figure}[t!]
 \centering
-\includegraphics[width=7cm]{sc_intersectionsimple.png}
+\includegraphics[width=6cm]{sc_intersectionsimple.png}
 \caption{Simple intersection example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.83 s, 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=12cm]{cv_eg12.png}
+\includegraphics[width=7cm]{cv_eg12.png}
 \caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
 \label{fig:cv_eg12}
-\end{figure*}
+\end{figure}
 
@@
 \end{figure*}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=7cm]{family.png}
+\caption{Combining iterations = 7. union 4.01 s intersection 4.12 s}
+\label{fig:donutlogic}
+\end{figure}
 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=12cm]{sc_noisy.png}
+\includegraphics[width=7cm]{sc_noisy.png}
 \caption{Same image as Figure~\ref{fig:threech}, but with 'salt \& pepper' noise with a variation of 0.1. Decreasing $\lambda$ from 255*255 to 125 allowed less
@@
 \label{fig:sc_noisy}
-\end{figure*}
+\end{figure}
 
@@
 z_i^{in}$ and $z_i^{out'} = 1 - z_i^{out}$. The results was actually successfully detected, but continuous flickering occurred, and the curve did not stop in
-one position. We discuss this results in more details in Section~\ref{sec:disc}.
+one position. We discuss this results in more details in Section~\ref{sec:difficulties}.
 
@@
 \label{sec:difficulties}
+We have faced many challenges throughout the implementation of this project. Some of the challenges caused major delays in the implementation cycle and hindered
+us from improving our code and adding more features to it.
+
+\subsection{Varying $\mu$}
+$\mu$ is a scaling factor that is used in both the Chan-Vese model and the Sandberg-Chan model to control the weight given to the length of the contour.
+Therefore, varying the value of $\mu$ will affect the ability to segment objects in noisy images, especially when the noise is similar to Matlab's \textit{salt
+and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure [\ref{fig:cvmu}] shows the
+segmentation of an object in a noisy for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting that
+this challenge kept us busy for over two days, after which we decided we should move on to other parts of the project.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{cvmu.png}
+\caption{$\mu$ does not have any effect on the length of the contour. \textit{Salt and pepper} noise was added to the original image for testing
+purposes. Top: original image with noise. Bottom: segmentation for various values of $\mu$.}
+\label{fig:cvmu}
+\end{figure*}
+
+\subsection{Computing the complement}
+The Sandberg-Chan model supports the complement logic operator, so one can segment the logic object ($A \cap \neg B$) (i.e. segment what is in channel A and not
+in channel B). The model defines the complement of the object in channel \textit{i} as follows:
+
+\begin{equation}
+\label{eqn:complement}
+\begin{align*}
+{z_{i}}^{in^{'}} = 1 - {z_{i}}^{in}, \\
+{z_{i}}^{out^{'}} = 1 - {z_{i}}^{out}.
+\end{align*}
+\end{equation}
+
+Although the definition is very simple, it does not work in practice. We implemented the complement logic operator as defined in Equation
+[\ref{eqn:complement}], and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
+flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure [\ref{fig:sccomp}] shows this scenrio. We tried to
+debug the reasons for this problem for three days.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{sccomp.png}
+\caption{The complement operator causes the contour to go through an infinite loop between two segmentations: the correct segmentation, and the segmentation
+of $A_1 \cap A_2$. Bottom: correct and incorrect segmentation respectively}
+\label{fig:sccomp}
+\end{figure*}
+
+\subsection{Defining the initial contour ($\phi_0$)}
+Both the Chan-Vese model and the Sandberg-Chan model defines the initial contour $\phi_0$ as a signed distance function where $\phi_0$ is positive inside the
+contour $C$, zero on the boundary of $C$ ($\partial C$), and negative outside the contour $C$. Figure [\ref{fig:phinote}] illustrates those assumptions.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=\textwidth]{phinote.png}
+\caption{Assumptions about the way $\phi_0$ is defined.}
+\label{fig:phinote}
+\end{figure}
+
+Since we model our initial contour as a circle, we have to modify the regular equation of the circle to match those assumption. Therefore, we initialize
+$\phi_0$ as follows:
+
+\begin{equation}
+\label{eqn:phinote}
+\phi_{0} (x, y) = radius - \sqrt{ (x - c_x)^2 + (y - c_y)^2 },
+\end{equation}
+where $c_x$ and $c_y$ are the x-coordinate and y-coordinate of the center of the circle respectively.
+
+Although this definition of $\phi_0$ is compatible with most of the images we used to test our implementation, it did fail to correctly segment the objects in
+some images. In fact, the results were the opposite of the used logic operator (i.e. the output is the intersection of the channels, while it should be the
+union). After spending quite sometime to resolve this problem and checking what might be the source of the error, we tried using the regular equation of a
+circle to initialize $\phi_0$ (Equation [\ref{eq:phinotenormal]}). We were very intrigued to notice that this modification resulted in the correct segmentation
+of the input channels. Figure [\ref{fig:phidef}] shows the final segmentation using both definitions for $\phi_0$.
+
+\begin{equation}
+\label{eqn:phinotenormal}
+\phi_{0} (x, y) = \sqrt{ (x - c_x)^2 + (y - c_y)^2 } - radius,
+\end{equation}
+where $c_x$ and $c_y$ are the x-coordinate and y-coordinate of the center of the circle respectively.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=\textwidth]{phidef.png}
+\caption{}
+\label{fig:phidef}
+\end{figure*}
+
+\subsection{Mistyped Equations}
+As previously mentiond in Section [\ref{sec:sandberg-chan}], there is a typo in the equations that define $z_in$ and $z_out$ in the Sandberg-Chan model
+(equation 3 in thier paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
+typos consumed over a week of our time until we were able to find out what are the correct equations in both cases. We could have spent this time overcoming
+other challenges that we faced while implementing both models.
 
@@
 \label{sec:concl}
-The conclusion goes here.
+In this project, we have provided an implementation for the Chan-Vese model for active contours without edges and the Sandberg-Chan logic framework for
+multi-channel images, both written in Matlab. Our experiments have shown that the Chan-Vese model is effective for various types of images. In specific, it is
+useful when an edge-based segmentation is not applicable (i.e. edges do not depend on a gradient). However, our implementation fails to respond to different
+values for the contour length scaling parameter, $\mu$. This affects the ability of our implementation to handle Matlab's \textit{salt \& pepper} noise. On the
+other hand, changing the values for the intensity weight parameters $\lambda_1$ and $\lambda_2$ affects the amount of details obtained inside and outside the
+segmented object respectively.
+
+We have also shown that our implementation for the Sandberg-Chan logic framework handles different logic combinations for multiple channels. We have also
+experimented using it as a segmentation tool for colored images by modeling an input colored images as a 3-channel image (red, green, blue). However, our
+implementation for the \textit{complement} logic operation seems to be missing a hidden trick in the Sandberg-Chan logic framework.
 
 \bibliographystyle{IEEEtran}
-
 \bibliography{references}

### 2010-12-13 trunk/cs870/project/CurveEvolution/evolveChannels.m
curve evolution

@@ -52,7 +52,7 @@ for n=1:iterations
     title('Segmentation');
     hold on;
 
-    % Plot the contour on the image
+	% Plot the contour on the image
     contour(phi, [0 0], 'r', 'LineWidth', 1);
     title(strcat('Iteration: ', num2str(n)));
     drawnow;

### 2010-12-13 cs870/project/CurveEvolution/updatePhiChannels.m
curve evolution

@@ -32,4 +32,3 @@ phi_t =  normOfGradient * ...
 
 % Update phi
 resultingPhi = phi + deltaT * phi_t;
-

### 2010-12-13 trunk/cs870/project/InitialPhi/cone.m
phi0

@@ -24,3 +24,4 @@ axes = cell(2, 1);
 % The equation is flipped because phi should be positive inside, and negative
 % outside.
 phi = radius - sqrt( (axes{1} - center(1)).^2 + (axes{2} - center(2)).^2 );
+% phi = sqrt( (axes{1} - center(1)).^2 + (axes{2} - center(2)).^2 ) - radius;

### 2010-12-13 cs870/project/Logic/scintersection.m
Logic

@@ -17,7 +17,7 @@ switch region
     case 'inside'
         for i = 1 : size(varargin, 2)
             if(size(find(complement == i) > 0))
-                output = (1 - sccomplement(zin(varargin{i}, phi))).*output;
+                output = (zin(varargin{i}, phi)).*output;
             else
                 output = (1 - zin(varargin{i}, phi)).*output;
             end
@@ -25,7 +25,7 @@ switch region
     case 'outside'
         for i = 1 : size(varargin, 2)
             if(size(find(complement == i) > 0))
-                output = (1 - sccomplement(zout(varargin{i}, phi))).*output;
+                output = (zout(varargin{i}, phi)).*output;
             else
                 output = (1 - zout(varargin{i}, phi)).*output;
             end

### 2010-12-13 cs870/project/Logic/scunion.m

@@ -17,7 +17,7 @@ switch(region)
     case 'inside'
         for i = 1 : size(varargin, 2)
             if(size(find(complement == i) > 0))
-                output = sccomplement(zin(varargin{i}, phi)).*output;
+                output = (1 - zin(varargin{i}, phi)).*output;
             else
                 output = zin(varargin{i}, phi).*output;
             end
@@ -25,7 +25,7 @@ switch(region)
     case 'outside'
         for i = 1 : size(varargin, 2)
             if(size(find(complement == i) > 0))
-                output = sccomplement(zout(varargin{i}, phi)).*output;
+                output = (1 - zout(varargin{i}, phi)).*output;
             else
                 output = zout(varargin{i}, phi).*output;
             end
@@ -33,5 +33,4 @@ switch(region)
 end
 
 output = real(output);
-
 output = output.^(1/size(varargin,2));

### 2010-12-13 trunk/cs870/project/Logic/fin.m
Logic

@@ -16,7 +16,7 @@ function output = fin(phi, logicop, complement, varargin)
 
 switch logicop
     case 'union'        
-        output = scunion(phi, 'inside', complement, varargin{:});               
+        output = scunion(phi, 'inside', complement, varargin{:});
     case 'intersection'        
         output = scintersection(phi, 'inside', complement, varargin{:});
 end

### 2010-12-13 cs870/project/Tests/test1.m
tests

@@ -6,7 +6,7 @@ startTime = cputime;
 
 iterations = 100;
 
-image = readGSImage('multiple.jpg', true);
+image = readGSImage('brain.jpg', false);
 imageSize = size(image, 1);
 
 phi0 = cone(floor(imageSize/4), [floor(imageSize/2) floor(imageSize/2)], size(image));

### 2010-12-13 cs870/project/Tests/test2.m

@@ -3,17 +3,13 @@ function test2()
 
 startTime = cputime;
 
-iterations = 10;
+iterations = 100;
 
-image = imread('flowers.jpg');
-% image = readGSImage('image.jpg');
+image = readGSImage('donut.jpg', false);
 imageSize = size(image, 1);
 
-[r g b] = readRGBImage('flowers.jpg');
-
-% image1 = readGSImage('image1.jpg');
-% image2 = readGSImage('image2.jpg');
-% image3 = readGSImage('image3.jpg');
+image1 = readGSImage('donut1.jpg', false);
+image2 = readGSImage('donut2.jpg', false);
 
 phi0 = cone(imageSize/4, [imageSize/2 imageSize/2], size(image));
 
@@ -21,14 +17,13 @@ mu = 0.01;
 
 lambda = 255*255;
 
-logicop = 'union';
+logicop = 'intersection';
 
 doReinit = false;
 
 complement = 0;
 
-% evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, complement, image1, image2, image3);
-evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, complement, r, g, b);
+evolveChannels(iterations, phi0, mu, lambda, logicop, image, doReinit, complement, image1, image2);
 
 timeTaken = cputime - startTime;
 

### 2010-12-13 cs870/project/Report/Paper.tex

@@
 \usepackage{cvpr}
+\usepackage{times}
 \usepackage{graphicx}
-\usepackage{tabularx}
-\usepackage{amsmath}
-\usepackage[pagebackref=true,breaklinks=true,letterpaper=true,colorlinks,bookmarks=true]{hyperref}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\usepackage{url}
+
+\cvprfinalcopy % *** Uncomment this line for the final submission
+
+\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
+\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
+
+% Pages are numbered in submission mode, and unnumbered in camera-ready
+\ifcvprfinal\pagestyle{empty}\fi
 
@@
 In this project, we provide an implementation for the Chan-Vese model for active contours without edges, and the Sandberg-Chan logic framework for active
-contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It differs
-from other active contour modesl in that it does not depend on the gradient of the contour ($\phi$), therefore it is more capable of detecting objects whose
+contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It
+differs
+from other active contour models in that it does not depend on the gradient of the contour ($\phi$), therefore it is more capable of detecting objects whose
 boundaries may not be defined by a gradient (e.g. blurry images, noisy images). The Sandberg-Chan model builds on top of the Chan-Vese model by allowing the
@@
 Image segmentation or boundary detection is a very important problem in the area of Image Processing, and has received a lot of attention in the past. The
-classical Active Contour model (or Snakes) proposed by Kass et al. \cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
+classical Active Contour model (or Snakes) proposed by Kass et al.~\cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
 a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it very highly adopted later on
@@
 usually had two steps. They would either first segment the object in each channel separately then combine the segmented objects according to the logic operation
-through bitwise operations (CITE) or they would apply logic operations to the different images then segment the resulting image (CITE). The first approach is
+through bitwise operations (E.g.~\cite{sapiro1997color, zhu2002region}) or they would apply logic operations to the different images then segment the resulting
+image. The first approach is
 very costly, and the second approach requires a lot of prior knowledge about the intensities of each image. To overcome these drawbacks, the Sandberg-Chan
 model is based on the idea of fitting a single contour to the object on all channels according to the logic operator, and based on regions.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=6cm]{explaining.png}
+\caption{Region Based Model}
+\label{fig:region}
+\end{figure}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[t!]
 \centering
-\includegraphics[width=6cm]{explaining.png}
-\caption{Region Based Model}
-\label{fig:region}
+\includegraphics[width=6cm]{fitting.png}
+\caption{Chan-Vese Curve Fitting}
+\label{fig:fitting}
 \end{figure}
-
-(REST OF EQUATION HERE)
 
@@
 outside the curve respectively. Since we want to give both forces equal weight, the authors set $\lambda_{1} = \lambda_{2} = 1$ in all their experiments.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=0.7\textwidth]{logicop.png}
+\caption{Region Based View Union}
+\label{fig:logic-op}
+\end{figure*}
 
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=6cm]{fitting.png}
-\caption{Chan-Vese Curve Fitting}
-\label{fig:fitting}
-\end{figure}
-
 More formally, the authors define two logic variables $z_{i}^{in}$ and $z_{i}^{out}$ to denote whether a point $(x,y)$ should be in the moving curve C or not
@@
 divide by the maximum intensity squared. The equations shown below have been corrected for that.
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=12cm]{logicop.png}
-\caption{Region Based View Union}
-\label{fig:logic-op}
-\end{figure*}
 
@@
 replaced by Equation~\ref{eqn:finters}.
-
-\begin{equation}
-\label{eqn:svpde}
-\frac{\partial{\phi}}{\partial{t}} = \delta(\phi)[\mu\bigtriangledown(\frac{\bigtriangledown \phi}{|\bigtriangledown \phi|})  - \lambda(f_{in}(z_{1}^{in}, ...,
-z_{n}^{in}) - f_{out}(z_{1}^{out}, ..., z_{n}^{out})
-\end{equation}
-
-with the boundary condition
-
-$\frac{\delta(\phi)}{|\bigtriangledown\phi|}\frac{\partial\phi}{\partial n^{\rightarrow}} = 0$
-
-\section{Implementation}
-\label{sec:implementation}
-
-During implementation, we tried to stick to the authors' formulas and guidelines as much as possible. In this section we explain the choices we made regardless
-implementation. For the chan-vese model, we used the PDE given in Equation 9 in the original paper~\cite{chan2001active} which is shown in
-Equation~\ref{eqn:cvpde}.  We, first, tried to implement the delta dirac function, $\delta_{\varepsilon}\phi$, as defined in the paper. However, we did not
-get non-zero values everywhere as indicated by the authors. Accordingly, we chose to use $|\bigtriangledown\phi|$ instead of the delta dirac function as this
-was indicated as a valid alternative by the authors. To calculate $c_{1}$ and $c_{2}$, we simply calculated the mean of the values inside $\phi$ (specifically
-where $\phi > 0$) and the mean of the values outside $\phi$ (specifically where $\phi < 0$) respectively. To actually solve the PDE, we chose to use an explicit
-time stepping scheme for the finite difference discretization. That is, in each time step $\phi_{t} = \phi + \triangle t*force$. This was simpler to implement,
-and although $\triangle t$ must be chosen carefully to satisfy the stability condition, we did not suffer from performance problems due to this restriction.
-
-For the sandberg-chan model, we used the PDE given in original paper as well, and shown in Equation~\ref{eqn:svpde}. Again, we used $|\bigtriangledown\phi|$
-instead of the dirac function. We calculated $z_{in}$ and $z_{out}$ according to Equations~\ref{eqn:zin, eqn:zout} shown above which have been corrected for
-the typo in the original paper. For all other equations and calculations necessary in the model, we closely followed the original paper in our implementation.
-We also used an explicit time stepping scheme in this model.
-
-\section{Experimental Results}
-\label{sec:results}
-
-In order to make sure we correctly implemented the models, we had several evaluation criteria. Table~\ref{tab:criteria} summarizes these criteria, and explains
-the goals of each. Some of these criteria apply for both models, while other apply to one or the other. For the rest of this section, we proceed by presenting
-our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
-cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
-the images in the Chan-Vese model were run on a BLA COMPUTER and all the images in the Sandberg-Chan model were run on a BLA computer. As shown in the results,
-we tried different image sizes to ensure that our solution converges.
 
@@
 
+\begin{equation}
+\label{eqn:svpde}
+\frac{\partial{\phi}}{\partial{t}} = \delta(\phi)[\mu\bigtriangledown(\frac{\bigtriangledown \phi}{|\bigtriangledown \phi|})  - \lambda(f_{in}(z_{1}^{in}, ...,
+z_{n}^{in}) - f_{out}(z_{1}^{out}, ..., z_{n}^{out})
+\end{equation}
+
+\section{Implementation}
+\label{sec:implementation}
+
+During implementation, we tried to stick to the authors' formulas and guidelines as much as possible. In this section we explain the choices we made regardless
+implementation. For the chan-vese model, we used the PDE given in Equation 9 in the original paper~\cite{chan2001active} which is shown in
+Equation~\ref{eqn:cvpde}.  We, first, tried to implement the delta dirac function, $\delta_{\varepsilon}\phi$, as defined in the paper. However, we did not
+get non-zero values everywhere as indicated by the authors. Accordingly, we chose to use $|\bigtriangledown\phi|$ instead of the delta dirac function as this
+was indicated as a valid alternative by the authors. To calculate $c_{1}$ and $c_{2}$, we simply calculated the mean of the values inside $\phi$ (specifically
+where $\phi > 0$) and the mean of the values outside $\phi$ (specifically where $\phi < 0$) respectively. To actually solve the PDE, we chose to use an explicit
+time stepping scheme for the finite difference discretization. That is, in each time step $\phi_{t} = \phi + \triangle t*force$. This was simpler to implement,
+and although $\triangle t$ must be chosen carefully to satisfy the stability condition, we did not suffer from performance problems due to this restriction.
+
+For the sandberg-chan model, we used the PDE given in original paper as well, and shown in Equation~\ref{eqn:svpde}. Again, we used $|\bigtriangledown\phi|$
+instead of the dirac function. We calculated $z_{in}$ and $z_{out}$ according to Equations~\ref{eqn:zin, eqn:zout} shown above which have been corrected for
+the typo in the original paper. For all other equations and calculations necessary in the model, we closely followed the original paper in our implementation.
+We also used an explicit time stepping scheme in this model.
+
+\section{Experimental Results}
+\label{sec:results}
+
+In order to make sure we correctly implemented the models, we had several evaluation criteria. Table~\ref{tab:criteria} summarizes these criteria, and explains
+the goals of each. Some of these criteria apply for both models, while other apply to one or the other. For the rest of this section, we proceed by presenting
+our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
+cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
+the images in the Chan-Vese model were run on a BLA COMPUTER and all the images in the Sandberg-Chan model were run on a BLA computer. As shown in the results,
+we tried different image sizes to ensure that our solution converges.
+
 \subsection{Chan-Vese Model Results}
@@
 $\lambda_{1} = \lambda_{2} = 1$.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.5\textwidth]{cv_eg1.png}
+\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
+0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
+\label{fig:cv_eg1}
+\end{figure}
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg2.png}
+\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2)} + 32.8$, $\mu =
+0.01$, no reinitialization, cpu = 1.95 s, took 7 iterations.}
+\label{fig:cv_eg2}
+\end{figure*}
 
@@
 \centering
-\includegraphics[width=11cm]{cv_eg1.png}
-\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
-0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
-\label{fig:cv_eg1}
-\end{figure*}
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=12cm]{cv_eg2.png}
-\caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2)} + 32.8$, $\mu =
-0.01$, no reinitialization, cpu = 1.95 s, took 7 iterations.}
-\label{fig:cv_eg2}
-\end{figure*}
-
-\begin{figure*}[t]
-\centering
 \includegraphics[width=12cm]{cv_eg8.png}
@@
 
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=12cm]{cv_eg3.png}
+\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2)} + 20$, $\mu =
+0.01$, no reinitialization, cpu = 2.26 s, took 9 iterations.}
+\label{fig:cv_eg3}
+\end{figure*}
+
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=12cm]{cv_eg4.png}
+\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
+$\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
+\label{fig:cv_eg4}
+\end{figure*}
+
 \subsubsection*{Independence of Initial Curve Position}
 
-To ensure that the position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the same brain image
+To ensure that the position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the lesame brain image
 used in Figure~\ref{fig:cv_eg2}. In Figure~\ref{fig:cv_eg2}, the initial contour was completely overlapping the object. Accordingly, we tried two other
@@
 \centering
-\includegraphics[width=12cm]{cv_eg3.png}
-\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2)} + 20$, $\mu =
-0.01$, no reinitialization, cpu = 2.26 s, took 9 iterations.}
-\label{fig:cv_eg3}
+\includegraphics[width=12cm]{cv_eg5.png}
+\caption{Successful detection of holes.  Top: the evolving curve (in red) over time where the first image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.5)^2 + (y - 65.5)^2)} + 32.8$,
+$\mu =0.01$, no reinitialization, cpu = 8.12 s, took 12 iterations.}
+\label{fig:cv_eg5}
 \end{figure*}
 
-\begin{figure*}[t]
+\begin{figure*}[t!h!]
 \centering
-\includegraphics[width=12cm]{cv_eg4.png}
-\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
-$\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
-\label{fig:cv_eg4}
+\includegraphics[width=12cm]{cv_eg6.png}
+\caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 15.07 s, took 23 iterations.}
+\label{fig:cv_eg6}
 \end{figure*}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.5\textwidth]{cv_eg7.png}
+\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
+image shows the initial
+contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
+\label{fig:cv_eg7}
+\end{figure}
 
@@
 
-\begin{figure*}[t]
-\centering
-\includegraphics[width=12cm]{cv_eg5.png}
-\caption{Successful detection of holes.  Top: the evolving curve (in red) over time where the first image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.5)^2 + (y - 65.5)^2)} + 32.8$,
-$\mu =0.01$, no reinitialization, cpu = 8.12 s, took 12 iterations.}
-\label{fig:cv_eg5}
-\end{figure*}
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=12cm]{cv_eg6.png}
-\caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
-$\mu =0.01$, no reinitialization, cpu = 15.07 s, took 23 iterations.}
-\label{fig:cv_eg6}
-\end{figure*}
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=10cm]{cv_eg7.png}
-\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
-image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
-$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
-\label{fig:cv_eg7}
-\end{figure*}
-
 \subsubsection*{Reasonable Performance with Blurred Images}
@@
 
-\begin{figure*}[t]
+\begin{figure*}[t!]
 \centering
@@
 \subsubsection*{Reasonable Performance with Noisy Images}
-
-\begin{figure}[t!]
-\centering
-\includegraphics[width=6cm]{sc_unionsimple.png}
-\caption{Simple union example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.32 s, 
-iterations = 2.}
-\label{fig:sc_unionsimple}
-\end{figure}
-
-\begin{figure}[t!]
-\centering
-\includegraphics[width=6cm]{sc_intersectionsimple.png}
-\caption{Simple intersection example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.83 s, 
-iterations = 2.}
-\label{fig:sc_intersectionsimple}
-\end{figure}
 
@@
 
-\begin{figure}[t]
+\begin{figure}[t!]
 \centering
-\includegraphics[width=7cm]{threech.png}
-\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 * 200. $\phi_{0}(x,y) = - \sqrt{(x -
-100)^2 + (y - 100)^2)} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
-5. For intersection then union: cpu = 1.24 and iterations = 4.}
-\label{fig:threech}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=7cm]{cv_eg12.png}
+\includegraphics[width=0.5\textwidth]{cv_eg12.png}
 \caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
@@
 
+\begin{figure}[t!]
+\centering
+\includegraphics[width=6cm]{sc_unionsimple.png}
+\caption{Simple union example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.32 s, 
+iterations = 2.}
+\label{fig:sc_unionsimple}
+\end{figure}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=6cm]{sc_intersectionsimple.png}
+\caption{Simple intersection example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.83 s, 
+iterations = 2.}
+\label{fig:sc_intersectionsimple}
+\end{figure}
+
 \subsubsection*{Logic Operators: union and intersection}
@@
 
-\begin{figure*}[t]
+\subsubsection*{Independence of Initial Curve Position}
+
+\begin{figure*}[t!h!]
 \centering
@@
 
-\begin{figure}[t]
+\begin{figure}[t!h!]
 \centering
-\includegraphics[width=7cm]{family.png}
-\caption{Combining iterations = 7. union 4.01 s intersection 4.12 s}
-\label{fig:donutlogic}
+\includegraphics[width=0.3\textwidth]{family.png}
+\caption{Combining missing information from two different channels. Size = 114 * 114. $\phi_{0}(x,y) = - \sqrt{(x - 57)^2 + (y - 57)^2)} +
+28.5$, no reinitialization. Union: cpu = 4.01 s and iterations = 7. Intersection: cpu = 4.12 s and iterations = 7.}
+\label{fig:family}
 \end{figure}
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=12cm]{flowers.png}
-\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
-Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
-size = cpu = 1.73 sec, iterations = 7.}
-\label{fig:flowers}
-\end{figure*}
-
-\subsubsection*{Independence of Initial Curve Position}
 
@@
 \centering
-\includegraphics[width=7cm]{sc_position.png}
+\includegraphics[width=0.3\textwidth]{sc_position.png}
 \caption{Same example shown in Figure~\ref{fig:sc_unionsimple}, but with different initial curve position. Initial curve does not overlap with either objects.
@@
 
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=12cm]{flowers.png}
+\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
+Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
+size = cpu = 1.73 sec, iterations = 7.}
+\label{fig:flowers}
+\end{figure*}
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.3\textwidth]{threech.png}
+\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 * 200. $\phi_{0}(x,y) = - \sqrt{(x -
+100)^2 + (y - 100)^2)} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
+5. For intersection then union: cpu = 1.24 and iterations = 4.}
+\label{fig:threech}
+\end{figure}
+
 \begin{figure}[t]
 \centering
-\includegraphics[width=7cm]{sc_noisy.png}
+\includegraphics[width=0.5\textwidth]{sc_noisy.png}
 \caption{Same image as Figure~\ref{fig:threech}, but with 'salt \& pepper' noise with a variation of 0.1. Decreasing $\lambda$ from 255*255 to 125 allowed less
@@
 
-\subsection{Varying $\mu$}
+\subsection{Varying mu}
 $\mu$ is a scaling factor that is used in both the Chan-Vese model and the Sandberg-Chan model to control the weight given to the length of the contour.
 Therefore, varying the value of $\mu$ will affect the ability to segment objects in noisy images, especially when the noise is similar to Matlab's \textit{salt
-and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure [\ref{fig:cvmu}] shows the
+and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure~\ref{fig:cvmu} shows the
 segmentation of an object in a noisy for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting that
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{cvmu.png}
+\includegraphics[width=0.5\textwidth]{cvmu.png}
 \caption{$\mu$ does not have any effect on the length of the contour. \textit{Salt and pepper} noise was added to the original image for testing
@@
 \label{fig:cvmu}
-\end{figure*}
+\end{figure}
 
@@
 \label{eqn:complement}
-\begin{align*}
+%\begin{align*}
 {z_{i}}^{in^{'}} = 1 - {z_{i}}^{in}, \\
 {z_{i}}^{out^{'}} = 1 - {z_{i}}^{out}.
-\end{align*}
+%\end{align*}
 \end{equation}
 
-Although the definition is very simple, it does not work in practice. We implemented the complement logic operator as defined in Equation
-[\ref{eqn:complement}], and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
-flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure [\ref{fig:sccomp}] shows this scenrio. We tried to
+Although the definition is very simple, it does not work in practice. We implemented the complement logic operator as defined in Equation~\ref{eqn:complement},
+and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
+flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure~\ref{fig:sccomp} shows this scenrio. We tried to
 debug the reasons for this problem for three days.
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{sccomp.png}
+\includegraphics[width=0.25\textwidth]{sccomp.png}
 \caption{The complement operator causes the contour to go through an infinite loop between two segmentations: the correct segmentation, and the segmentation
@@
 \label{fig:sccomp}
-\end{figure*}
+\end{figure}
 
-\subsection{Defining the initial contour ($\phi_0$)}
+\subsection{Defining the initial contour phi0}
 Both the Chan-Vese model and the Sandberg-Chan model defines the initial contour $\phi_0$ as a signed distance function where $\phi_0$ is positive inside the
-contour $C$, zero on the boundary of $C$ ($\partial C$), and negative outside the contour $C$. Figure [\ref{fig:phinote}] illustrates those assumptions.
+contour $C$, zero on the boundary of $C$ ($\partial C$), and negative outside the contour $C$. Figure~\ref{fig:phinote} illustrates those assumptions.
 
@@
 \centering
-\includegraphics[width=\textwidth]{phinote.png}
+\includegraphics[width=0.5\textwidth]{phinote.png}
 \caption{Assumptions about the way $\phi_0$ is defined.}
@@
 union). After spending quite sometime to resolve this problem and checking what might be the source of the error, we tried using the regular equation of a
-circle to initialize $\phi_0$ (Equation [\ref{eq:phinotenormal]}). We were very intrigued to notice that this modification resulted in the correct segmentation
-of the input channels. Figure [\ref{fig:phidef}] shows the final segmentation using both definitions for $\phi_0$.
+circle to initialize $\phi_0$ (Equation~\ref{eq:phinotenormal]}). We were very intrigued to notice that this modification resulted in the correct segmentation
+of the input channels. Figure~\ref{fig:phidef} shows the final segmentation using both definitions for $\phi_0$.
 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{phidef.png}
+\includegraphics[width=0.2\textwidth]{phidef.png}
 \caption{}
 \label{fig:phidef}
-\end{figure*}
+\end{figure}
 
 \subsection{Mistyped Equations}
-As previously mentiond in Section [\ref{sec:sandberg-chan}], there is a typo in the equations that define $z_in$ and $z_out$ in the Sandberg-Chan model
+As previously mentiond in Section~\ref{sec:sandberg-chan}, there is a typo in the equations that define $z_in$ and $z_out$ in the Sandberg-Chan model
 (equation 3 in thier paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
@@
 
-\bibliographystyle{IEEEtran}
 \bibliography{references}
+\bibliographystyle{ieee}
 

### 2010-12-13 trunk/cs870/project/Report/Paper.tex
final version

@@
 \usepackage{cvpr}
-\usepackage{cite}
+\usepackage{times}
+\usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage{algorithmic}
 \usepackage{url}
-\usepackage{graphicx}
-\usepackage{tabularx}
-\usepackage{amsmath}
-\usepackage{array}
-\usepackage{color}
-\usepackage{balance}
+\usepackage{appendix}
+
+\cvprfinalcopy % *** Uncomment this line for the final submission
+
+\def\cvprPaperID{****} % *** Enter the CVPR Paper ID here
+\def\httilde{\mbox{\tt\raisebox{-.5ex}{\symbol{126}}}}
+
+% Pages are numbered in submission mode, and unnumbered in camera-ready
+\ifcvprfinal\pagestyle{empty}\fi
 
@@
 
-\title{An Implementation of the Active Contours without Edges model and the Logic Framework for Active contours on multi-channel images}
+\title{An Implementation of the Active Contours without Edges Model and the Logic Framework for Active Contours on Multi-channel Images}
 
@@
 \{karim, snadi\}@cs.uwaterloo.ca \\
+CS870 - Fall 2010\\
 David R. Cheriton School of Computer Science\\
@@
 In this project, we provide an implementation for the Chan-Vese model for active contours without edges, and the Sandberg-Chan logic framework for active
-contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It differs
-from other active contour modesl in that it does not depend on the gradient of the contour ($\phi$), therefore it is more capable of detecting objects whose
+contours on multi-channel images. The Chan-Vese model is a special case of the more general Mumford-Shah functional for segmentation and level sets. It
+differs
+from other active contour models in that it is not edge dependent, therefore it is more capable of detecting objects whose
 boundaries may not be defined by a gradient (e.g. blurry images, noisy images). The Sandberg-Chan model builds on top of the Chan-Vese model by allowing the
@@
 classical Active Contour model (or Snakes) proposed by Kass et al.~\cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
-a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it very highly adopted later on
-(E.g.~\cite{caselles1997geodesic}. Most of these models used the level set formulation for propagating fronts to evolve the curve. However, these models highly
-depended on curvature motion (motion defined by the gradient of the curve) which led to poor performance in smoothed edges. To overcome these limitations, Chan
+a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it were highly adopted later on
+(E.g.~\cite{caselles1997geodesic}. Most of these models use the level set formulation for propagating fronts to evolve the curve. However, these models highly
+depend on motion defined by the gradient of the image which leads to poor performance in smoothed edges. To overcome these limitations, Chan
 and Vese~\cite{chan2001active} propose an active contour model that does not depend on the edges (i.e the gradient) for propagating the curve to detect the
-boundary of the object. Instead, they use a region-based approach based on the Mumford-Shah model~\cite{mumford1989optimal} to divide the image
+boundary of the object. Instead, they use a region-based approach based on the Mumford-Shah model \cite{mumford1989optimal} to divide the image
 into two regions: one inside the propagating curve, and one outside. The curve is at the boundary of the object if there is no difference in intensities inside
@@
 
-The Active Contours without Edges model proposed by Chan and Vese (referred to as Chan-Vese model throughout the rest of this paper) was very successful in
-detecting objects even in noisy or blurry images. It could also detect holes in objects which was usually a limitation in previous models. As an extension to
+The Active Contours without Edges model proposed by Chan and Vese (referred to as Chan-Vese model throughout the rest of this paper) is very successful in
+detecting objects even in noisy or blurry images. It can also detect holes in objects which is usually a limitation in previous models. As an extension to
 this work, Sandberg and Chan~\cite{sandberg2005logic} propose a logic framework (referred to as the Sandberg-Chan model throughout the rest of this
-paper) that  performs logical operations on multiple images according to the curve propagation proposed in the Chan-Vese model. To achieve that, previous models
+paper) that  performs logic operations on multiple images according to the curve propagation proposed in the Chan-Vese model. To achieve that, previous models
 usually had two steps. They would either first segment the object in each channel separately then combine the segmented objects according to the logic operation
-through bitwise operations (CITE) or they would apply logic operations to the different images then segment the resulting image (CITE). The first approach is
+through bitwise operations~\cite{sapiro1997color, zhu2002region} or they would apply logic operations to the different images then segment the resulting
+image. The first approach is
 very costly, and the second approach requires a lot of prior knowledge about the intensities of each image. To overcome these drawbacks, the Sandberg-Chan
@@
 
-In this paper, we report on our findings after implementing each of these models. We implemented both models in Matlab, and experimented with several images.
-In this paper, we explain the details of our implementation as well as our results. The rest of this paper is organized as follows: Section~\ref{sec:bg} first
-provides brief background about each of the two models implemented in this paper. Section~\ref{sec:implementation} then explains our implementation. In this
-section, we explain how we implemented the models and any variations from the original papers. Section~\ref{sec:results} explains our evaluation criteria, and
-presents the results we obtained. Section~\ref{sec:difficulties} discusses some of the difficulties we faced, and points out some BLA. Section~\ref{sec:concl}
+\begin{figure}[t]
+\centering
+\includegraphics[width=6cm]{explaining.png}
+\caption{Region Based Model}
+\label{fig:region}
+\end{figure}
+
+In this paper, we report on our findings after implementing each of these models. We implement both models in Matlab, and experiment with several images.
+We explain the details of our implementation as well as our results. The rest of this paper is organized as follows:
+Section~\ref{sec:bg} provides brief background about each of the two models implemented in this paper. Section~\ref{sec:implementation} then explains our
+implementation. In this section, we explain how we implemented the models and any variations from the original papers. Section~\ref{sec:results} explains our
+evaluation criteria, and
+presents the results we obtained. Section~\ref{sec:difficulties} discusses some of the difficulties we faced during implementation. Section~\ref{sec:concl}
 concludes this paper by summarizing our findings.
@@
 
-The Chan-Vese model is a region based model for detecting objects in an image. It is based on a restriction Mumford-Shah model which divides an image into
-regions and represents each region by a piecewise constant (the minimal partition problem). Figure~\ref{fig:region} shows what is meant by a region based
-model. The figure shows an image $u_{0}$ which has a gray triangular object in it. The red curve, C, is the initial contour used to detect this object. The
+The Chan-Vese model~\cite{chan2001active} is a region based model for detecting objects in an image. It is based on a restriction of the Mumford-Shah model
+which divides an image into
+regions and represents each region by a piecewise constant (also called the minimal partition problem). Figure~\ref{fig:region} shows this region based model.
+The figure shows an image $u_{0}$ which has a gray triangular object in it. The red curve, $C$, is the initial contour used to detect this object. The
 main idea behind the model is that the curve divides the image into two regions: that inside the curve and that outside. Each region is represented by a
-constant, c, which is the average intensity of the image values in each region. In order for the curve to fit the object, there must be no variation of the
+constant, $c$, which is the average intensity of the image values in each region. In order for the curve to fit the object, there must be minimal variation of
+the
 intensities inside the curve as well as outside. In other words, this turns into a minimization problem of the difference of intensities inside (fitting term
 one, $F_{1}$) plus those
-outside (fitting term two, $F_{2}$). For example, in the figure, the point $(x_{1},y_{1})$ will have to be outside the curve in order to minimize the difference
+outside (fitting term two, $F_{2}$). For example, in Figure~\ref{fig:region}, the point $(x_{1},y_{1})$ will have to be outside the curve in order to minimize
+the difference
 between the points inside
@@
 
-More formally, the Euler-Lagrange equation (as derived in the original paper) representing the time motion of the curve C is shown in Equation~\ref{eqn:cvpde}.
+More formally, the Euler-Lagrange equation (as derived in the original paper) representing the time motion of the curve $C$ is shown in
+Equation~\ref{eqn:cvpde}.
 
@@
 \end{equation}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=6cm]{explaining.png}
-\caption{Region Based Model}
-\label{fig:region}
-\end{figure}
-
-(REST OF EQUATION HERE)
-
-The parameter $\mu$ is a scaling parameter for the length of the curve represented in terms of curvature. The smalled $\mu$ is, the more the length of the curve
-can increase without penalizing the minimization. This allows the model to detect smalled objects and holes. The larger $\mu$ is, the less freedom there is for
-the curve to increase in length, and thus, it will only be able to detect larger objects. The parameter $\nu$ is also a scaling term for the area of the curve.
-However, the authors do not use the area term in the Euler-Lagrange derivation, and always set $\nu$ to 0. It seems that $\mu$ is sufficient to scale the curve
-according to the objects that need to be detected. Finally, $\lambda_{1}$ and $\lambda_{2}$ are weighting parameters for the forces inside the curve and
-outside the curve respectively. Since we want to give both forces equal weight, the authors set $\lambda_{1} = \lambda_{2} = 1$ in all their experiments.
-
-\subsection{Sandberg-Chan Model}
-\label{sec:sandberg-chan}
-
-The main goal of the Sandberg-Chan model~\cite{sandberg2005logic} is to perform logic operations on a combination of different images accurately and
-efficiently. For example, finding the union of two images where different parts of the object are occluded in each image so that a complete object can be
-obtained. In order to do that, they use the Chan-Vese model to detect objects, and simultaneously include the logic operations in the minimization problem.
-Since they are using the Chan-Vese model, the problem must be viewed in terms of regions as well. Accordingly, they define the two main logic operations
-(union and intersection) in terms of regions where the union of two images is the union of the insides of the curve with respect to the images plus the
-intersection of the outsides of the curve with respect to the images. For example, Figure~\ref{fig:logic-op} (taken from the original paper) shows that the
-union of $A_{1}$ and $A_{2}$ (in terms of its shape) can be obtained by taking the union of the insides of the object or the intersection of the outsides.
-Similarly, the intersection of two images would be the intersection of the insides plus the union of the outsides. Thus, in order to obtain accurate results
-irrespective of the object's intensity outside versus inside, we should consider \textit{both} the logical operation required for the region inside the object
-as well as that outside.
 
@@
 
-More formally, the authors define two logic variables $z_{i}^{in}$ and $z_{i}^{out}$ to denote whether a point $(x,y)$ should be in the moving curve C or not
-with respect to image i. Since we are trying to minimize the fitting of the curve, they use 0 to denote true and 1 to denote false (the reverse of the usual
-convention. Following the Chan-Vese model, they represent each of the regions inside and outside the curve by a constant c. In this paper, they represent
+The parameter $\mu$ is a scaling parameter for the length of the curve represented in terms of curvature. The smaller $\mu$ is, the more the length of the curve
+can increase without penalizing the minimization. This allows the model to detect smaller objects and holes. The larger $\mu$ is, the less freedom there is for
+the curve to increase in length, and thus, it will only be able to detect larger objects. The parameter $\nu$ is also a scaling term for the area of the curve.
+However, the authors do not use the area term in the Euler-Lagrange derivation, and always set $\nu$ to 0. It seems that $\mu$ is sufficient to scale the curve
+according to the objects that need to be detected. Finally, $\lambda_{1}$ and $\lambda_{2}$ are weighting parameters for the forces inside the curve and
+outside the curve respectively. Since we want to give both forces equal weight, the authors set $\lambda_{1} = \lambda_{2} = 1$ in all their experiments.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=0.7\textwidth]{logicop.png}
+\caption{Region Based View Union}
+\label{fig:logic-op}
+\end{figure*}
+
+\subsection{Sandberg-Chan Model}
+\label{sec:sandberg-chan}
+
+The main goal of the Sandberg-Chan model~\cite{sandberg2005logic} is to perform logic operations on a combination of different images accurately and
+efficiently. For example, finding the union of two images where different parts of the object are occluded in each image so that a complete object can be
+obtained. This example will be shown later in Figure~\ref{fig:family}. In order to do that, they use the Chan-Vese model to detect objects, and simultaneously
+include the logic operations in the minimization problem.
+Since they are using the Chan-Vese model, the problem must be viewed in terms of regions as well. Accordingly, they define the two main logic operations
+(union and intersection) in terms of regions where the union of two images is the union of the inside of the curve with respect to the images plus the
+intersection of the outside of the curve with respect to the images. For example, Figure~\ref{fig:logic-op}~\cite{sandberg2005logic} shows that the
+union of $A_{1}$ and $A_{2}$ (in terms of its shape) can be obtained by taking the union of the insides of the object or the intersection of the outsides.
+Similarly, the intersection of two images would be the intersection of the insides plus the union of the outsides. Thus, in order to obtain accurate results
+irrespective of the object's intensity outside versus inside, we should consider \textit{both} the logic operation required for the region inside the object
+as well as that outside.
+
+More formally, the authors define two logic variables $z_{i}^{in}$ and $z_{i}^{out}$ to denote whether a point $(x,y)$ should be in the moving curve $C$ or not
+with respect to image $i$. Since we are trying to minimize the fitting of the curve, they use $0$ to denote $true$ and $1$ to denote $false$ (the reverse of the
+usual
+convention). Following the Chan-Vese model, they represent each of the regions inside and outside the curve by a constant $c$. In this paper, they represent
 $c_{in}$ as $c_{+}$ and $c_{out}$ as $c_{-}$. Equations~\ref{eqn:zin} and~\ref{eqn:zout} show how to calculate $z_{in}$ and $z_{out}$ respectively in terms of
@@
 \label{eqn:finters}
-f_{\cap} = 1 = ((1 - z_{1}) .(1-  z_{2}))^{1/2}
+f_{\cap} = 1 - ((1 - z_{1}) .(1-  z_{2}))^{1/2}
 \end{equation}
@@
 Based on these functions, the Euler-Lagrange equation for any logic operation is shown in Equation~\ref{eqn:svpde}. According to the desired logic operation,
-$f_{in}$ and $f_{out}$ will be specified accordingly. For example, if we are doing the union of the images, then we will need to perform a union operation on
+$f_{in}$ and $f_{out}$ will be specified. For example, if we are doing the union of the images, then we will need to perform a union operation on
 the insides. Thus, $f_{in}$ will be replaced by Equation~\ref{eqn:funion}. Similarly, we will need to do the intersection of the outsides so $f_{out}$ will be
 replaced by Equation~\ref{eqn:finters}.
-
-\begin{equation}
-\label{eqn:svpde}
-\frac{\partial{\phi}}{\partial{t}} = \delta(\phi)[\mu\bigtriangledown(\frac{\bigtriangledown \phi}{|\bigtriangledown \phi|})  - \lambda(f_{in}(z_{1}^{in}, ...,
-z_{n}^{in}) - f_{out}(z_{1}^{out}, ..., z_{n}^{out})
-\end{equation}
-
-with the boundary condition
-
-$\frac{\delta(\phi)}{|\bigtriangledown\phi|}\frac{\partial\phi}{\partial n^{\rightarrow}} = 0$
-
-\section{Implementation}
-\label{sec:implementation}
-
-During implementation, we tried to stick to the authors' formulas and guidelines as much as possible. In this section we explain the choices we made regardless
-implementation. For the chan-vese model, we used the PDE given in Equation 9 in the original paper~\cite{chan2001active} which is shown in
-Equation~\ref{eqn:cvpde}.  We, first, tried to implement the delta dirac function, $\delta_{\varepsilon}\phi$, as defined in the paper. However, we did not
-get non-zero values everywhere as indicated by the authors. Accordingly, we chose to use $|\bigtriangledown\phi|$ instead of the delta dirac function as this
-was indicated as a valid alternative by the authors. To calculate $c_{1}$ and $c_{2}$, we simply calculated the mean of the values inside $\phi$ (specifically
-where $\phi > 0$) and the mean of the values outside $\phi$ (specifically where $\phi < 0$) respectively. To actually solve the PDE, we chose to use an explicit
-time stepping scheme for the finite difference discretization. That is, in each time step $\phi_{t} = \phi + \triangle t*force$. This was simpler to implement,
-and although $\triangle t$ must be chosen carefully to satisfy the stability condition, we did not suffer from performance problems due to this restriction.
-
-For the sandberg-chan model, we used the PDE given in original paper as well, and shown in Equation~\ref{eqn:svpde}. Again, we used $|\bigtriangledown\phi|$
-instead of the dirac function. We calculated $z_{in}$ and $z_{out}$ according to Equations~\ref{eqn:zin, eqn:zout} shown above which have been corrected for
-the typo in the original paper. For all other equations and calculations necessary in the model, we closely followed the original paper in our implementation.
-We also used an explicit time stepping scheme in this model.
-
-\section{Experimental Results}
-\label{sec:results}
-
-In order to make sure we correctly implemented the models, we had several evaluation criteria. Table~\ref{tab:criteria} summarizes these criteria, and explains
-the goals of each. Some of these criteria apply for both models, while other apply to one or the other. For the rest of this section, we proceed by presenting
-our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
-cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
-the images in the Chan-Vese model were run on a BLA COMPUTER and all the images in the Sandberg-Chan model were run on a BLA computer. As shown in the results,
-we tried different image sizes to ensure that our solution converges.
 
@@
 
+\begin{equation}
+\label{eqn:svpde}
+\frac{\partial{\phi}}{\partial{t}} = \delta(\phi)[\mu\bigtriangledown(\frac{\bigtriangledown \phi}{|\bigtriangledown \phi|})  - \lambda(f_{in}(z_{1}^{in}, ...,
+z_{n}^{in}) - f_{out}(z_{1}^{out}, ..., z_{n}^{out})
+\end{equation}
+
+\section{Implementation}
+\label{sec:implementation}
+
+During implementation, we tried to stick to the authors' formulas and guidelines as much as possible. In this section we explain what methods we use to
+numerically solve the PDEs. For the chan-vese model, we use the PDE given in Equation 9 in the original paper~\cite{chan2001active} which is shown in
+Equation~\ref{eqn:cvpde}.  We, first, tried to implement the dirac delta function, $\delta_{\varepsilon}\phi$, as defined in the paper. However, it resulted in
+worse segmentations. Accordingly, we chose to use $|\bigtriangledown\phi|$ instead of the dirac delta function as this
+was indicated as a valid alternative by the authors. To calculate $c_{1}$ and $c_{2}$, we simply calculated the mean of the values inside $\phi$ (specifically
+where $\phi > 0$) and the mean of the values outside $\phi$ (specifically where $\phi < 0$) respectively. To actually solve the PDE, we chose to use an explicit
+time stepping scheme for the finite difference discretization. That is, in each time step $\phi^{n+1} = \phi^n + \triangle t*phi_t$. This was simpler to
+implement, and although $\triangle t$ must be chosen carefully to satisfy the stability condition ($\triangle t <= (1/imageSize)^2$), we did not suffer from
+performance problems due to this restriction.
+
+For the Sandberg-Chan model, we used the PDE given in original paper as well, and shown in Equation~\ref{eqn:svpde}. Again, we used $|\bigtriangledown\phi|$
+instead of the dirac delta function. We calculated $z_{in}$ and $z_{out}$ according to Equations~\ref{eqn:zin} and~\ref{eqn:zout} shown above which have been
+corrected for
+the typo in the original paper. For all other equations and calculations necessary in the model, we closely followed the original paper in our implementation.
+We also used an explicit time stepping scheme in this model.
+
+We use the same stopping condition for both model. That is, to check if we have reached the stationary solution, we check if more points have been added inside
+the contour or not. More formally, we check if $(\phi^{n+1} > 0) == (\phi_{n} > 0)$. If the contour did not move, then new points will be added to its inside,
+and this condition will be satisfied.
+
+\section{Experimental Results}
+\label{sec:results}
+
+In order to make sure we correctly implemented the models, we had several evaluation criteria. Table~\ref{tab:criteria} summarizes these criteria, and explains
+the goals of each. Some of these criteria apply for both models, while other apply to one or the other. For the rest of this section, we proceed by presenting
+our experimental results for both models. For each model, we proceed in the order of these criteria starting with the simplest
+cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
+the images in the Chan-Vese model were run on an Intel Core 2 Duo processor with 2.00 GHz and a 2GB RAM and all the images in the Sandberg-Chan
+model were run on a Intel Centrino Duo 2.0 processor and a 1GB RAM computer. As shown in the results,
+we tried different image sizes to ensure that our solution converges.
+
 \subsection{Chan-Vese Model Results}
@@
 
-\subsubsection*{Successful Detection of Boundaries}
-
-Figure~\ref{fig:cv_eg1} shows that the implemented model can successfully detect object boundaries in a simple image. In this case, the evolving curve was
-successfully able to detect the boundaries of both objects. Figure~\ref{fig:cv_eg2} shows a slightly more complicated example where the evolving contour
-successfully detects the contour of the brain. Note that there is a tiny area to the bottom left of the brain that was also successfully detected. To ensure
-that open boundaries (e.g. lines) can also be detected, we used an image with several shapes shown in Figure~\ref{fig:cv_eg8};
-
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=11cm]{cv_eg1.png}
+\includegraphics[width=0.5\textwidth]{cv_eg1.png}
 \caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
-0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
+Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$, $\mu =
+0.01$, no reinitialization, cpu = 2.9s, iterations = 7.}
 \label{fig:cv_eg1}
-\end{figure*}
+\end{figure}
 
@@
 \caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2)} + 32.8$, $\mu =
-0.01$, no reinitialization, cpu = 1.95 s, took 7 iterations.}
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2} + 32.8$, $\mu =
+0.01$, no reinitialization, cpu = 1.95 s, iterations = 7.}
 \label{fig:cv_eg2}
 \end{figure*}
+
+\subsubsection*{Successful Detection of Boundaries}
+
+Figure~\ref{fig:cv_eg1} shows that the implemented model can successfully detect object boundaries in a simple image. In this case, the evolving curve was
+successfully able to detect the boundaries of both objects. Figure~\ref{fig:cv_eg2} shows a slightly more complicated example where the evolving contour
+successfully detects the contour of the brain. Note that there is a tiny area near the center to the left of the brain that was also successfully detected. To
+ensure
+that open boundaries (e.g. lines) can also be detected, we used an image with several shapes shown in Figure~\ref{fig:cv_eg8}.
 
@@
 \caption{Successful detection of object with open boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
-0.01$, no reinitialization, cpu = 5.19 s, took 11 iterations.}
+Bottom: evolving segmentation over time until all objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$, $\mu =
+0.01$, no reinitialization, cpu = 5.19 s, iterations = 11.}
 \label{fig:cv_eg8}
+\end{figure*}
+
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=12cm]{cv_eg3.png}
+\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2} + 20$, $\mu =
+0.01$, no reinitialization, cpu = 2.26 s, iterations = 9.}
+\label{fig:cv_eg3}
+\end{figure*}
+
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=12cm]{cv_eg4.png}
+\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2} + 10$,
+$\mu =0.01$, no reinitialization, cpu = 2.01 s, iterations = 9.}
+\label{fig:cv_eg4}
 \end{figure*}
@@
 
-To ensure that the position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the same brain image
+To ensure that the initial position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the same
+brain image
 used in Figure~\ref{fig:cv_eg2}. In Figure~\ref{fig:cv_eg2}, the initial contour was completely overlapping the object. Accordingly, we tried two other
@@
 \centering
-\includegraphics[width=12cm]{cv_eg3.png}
-\caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2)} + 20$, $\mu =
-0.01$, no reinitialization, cpu = 2.26 s, took 9 iterations.}
-\label{fig:cv_eg3}
+\includegraphics[width=12cm]{cv_eg5.png}
+\caption{Successful detection of holes.  Top: the evolving curve (in red) over time where the first image shows the initial
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.5)^2 + (y - 65.5)^2} + 32.8$,
+$\mu =0.01$, no reinitialization, cpu = 8.12 s, iterations = 12.}
+\label{fig:cv_eg5}
 \end{figure*}
 
-\begin{figure*}[t]
+\begin{figure*}[t!h!]
 \centering
-\includegraphics[width=12cm]{cv_eg4.png}
-\caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
-$\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
-\label{fig:cv_eg4}
+\includegraphics[width=12cm]{cv_eg6.png}
+\caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 15.07 s, iterations =  23.}
+\label{fig:cv_eg6}
 \end{figure*}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.5\textwidth]{cv_eg7.png}
+\caption{Inability to detect low intensities that do not have much variation from their background.  Top: the evolving curve (in red) over time where the
+first
+image shows the initial
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 1.35 s, iterations = 6.}
+\label{fig:cv_eg7}
+\end{figure}
 
@@
 image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 255 x 255, $\phi_{0}(x,y) = - \sqrt{(x - 127.5)^2 + (y - 127.5)^2)} +
-63.75$, $\mu =0.01$, no reinitialization, cpu = 5.41 s, took 15 iterations.}
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 255 x 255, $\phi_{0}(x,y) = - \sqrt{(x - 127.5)^2 + (y - 127.5)^2} +
+63.75$, $\mu =0.01$, no reinitialization, cpu = 5.41 s, iterations = 15.}
 \label{fig:cv_eg9}
@@
 
+Figure~\ref{fig:cv_eg10} shows how two objects in a noisy image were successfully detected. In this image, we added a Gaussian noise with mean = 0 and
+variation = 0.01
+variance using Matlab's built-in noise function. The image shows that although some of the noise was detected in intermediate iterations, the contour continued
+to evolve until it was only surrounding the desired objects. Unfortunately, the effect of noise was not completely ignored in all cases. For example,
+Figure~\ref{fig:cv_eg11} shows that for the same image, but with \textit{Salt \& pepper} noise, the noise was detected as objects. This should not have been the
+case
+since we used a large value $\mu = 5$ for the length scaling parameter. However, varying $\mu$ does not have the intended effect as explained
+in the next paragraph.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[width=12cm]{cv_eg10.png}
+\caption{Ability to detect contours in an image with Gaussian noise.  Top: the evolving curve (in red) over time where the first
+image shows the initial
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} +
+75$, $\mu =0.01$, no reinitialization, cpu = 4.4 s, iterations = 7.}
+\label{fig:cv_eg10}
+\end{figure*}
+
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=12cm]{cv_eg11.png}
+\caption{Figure~\ref{fig:cv_eg10}, but with \textit{Salt \& pepper} noise. Noise was still detected despite increasing $\mu$.  Top: the evolving
+curve
+(in red) over time where the first
+image shows the initial
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} +
+75$, $\mu = 5$, no reinitialization, cpu = 12.27 s, iterations = 7.}
+\label{fig:cv_eg11}
+\end{figure*}
+
+\subsubsection*{Successful Response to Parameter Settings}
+
+There are mainly three parameters in this model that can be varied: $\mu, \lambda_{1}, \lambda_{2}$. We would usually want both $\lambda_{1}$ and $\lambda_{2}$
+to be equal to 1 to indicate that we care both about the difference of intensities inside and outside. A smaller weighting for one of them would mean that we
+would ignore some of the variation in intensities in this area. A larger weighting means that we are magnifying the difference in intensities in that area
+(i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we vary
+$\lambda_1$ and $\lambda_2$. When we set $\lambda_1 = 0.2$ and $\lambda_2 = 1$, less details within the interior of the brain is detected. When
+we set $\lambda_1 = 2$ and $\lambda_2 = 1$, we can see that more details within the brain are detected. When we set $\lambda_1 = 5$, and $\lambda_2 = 0.01$ (a
+very extreme case), we can see that parts of outside boundaries are not well detected, while many details are detected within the brain.
+
+The second parameter, $\mu$ controls how much we allow the length of the curve to increase. If $\mu$ is small, it means that the curve length can increase to
+detect multiple smaller objects without penalizing our minimization problem. On the other hand if $\mu$ is large, it means that any change in the curve length
+will be scaled up, and thus will limit the curve expansion to keep the force to a minimum. Unfortunately, we were not able to see this effect in our
+experiments. We tried varying $\mu$ to be able to detect groupings of objects instead of the individual objects, but the segmentation was invariant
+to $\mu$. We give more details about how we tried to handle this, and why we believe it is not working in Section~\ref{sec:difficulties}.
+
+\subsection{Sandberg-Chan Model Results}
+
+For the Sandberg-Chan model, the successful detection of boundaries criterion is implicitly included in the logic operations functionality. We therefore, show
+the other criteria. Unless otherwise specified, $\mu = 0.1$ and $\lambda = 255*255$ in the experiments below.
+
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=0.5\textwidth]{cv_eg12.png}
+\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
+\label{fig:cv_eg12}
+\end{figure*}
+
 \begin{figure}[t!]
@@
 \includegraphics[width=6cm]{sc_unionsimple.png}
-\caption{Simple union example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.32 s, 
+\caption{Simple union example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$,  no reinitialization, cpu = 1.32 s, 
 iterations = 2.}
@@
 \includegraphics[width=6cm]{sc_intersectionsimple.png}
-\caption{Simple intersection example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,  no reinitialization, cpu = 1.83 s, 
+\caption{Simple intersection example. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$,  no reinitialization, cpu = 1.83 s, 
 iterations = 2.}
@@
 \end{figure}
-
-Figure~\ref{fig:cv_eg10} shows how two objects in a noisy image were successfully detected. In this image, we added Gaussian noise with mean zero and 0.01
-variance using Matlab's built-in noise function. The image shows that although some of the noise was detected in intermediate iterations, the contour continued
-to evolve until it was only surrounding the desired objects. Unfortunately, the effect of noise was not completely ignored in all cases. For example,
-Figure~\ref{fig:cv_eg11} shows that for the same image, but with 'salt \& pepper' noise, the noise was detected as objects. This should not have been the case
-since we used a large value $\mu = 5$ for the length scaling parameter. However, for some reason, varying $\mu$ did not have the intended effect as explained
-in the next section.
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=12cm]{cv_eg10.png}
-\caption{Ability to detect contours in an image with Gaussian noise.  Top: the evolving curve (in red) over time where the first
-image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
-75$, $\mu =0.01$, no reinitialization, cpu = 4.4 s, took 7 iterations.}
-\label{fig:cv_eg10}
-\end{figure*}
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=12cm]{cv_eg11.png}
-\caption{Same image as Figure~\ref{fig:cv_eg10}, but with 'salt \& pepper' noise. Noise was still detected despite increasing $\mu$.  Top: the evolving curve
-(in red) over time where the first
-image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
-75$, $\mu = 5$, no reinitialization, cpu = 12.27 s, took 7 iterations.}
-\label{fig:cv_eg11}
-\end{figure*}
-
-\subsubsection*{Successful Response to Parameter Settings}
-
-There are mainly three parameters in this model that can be varied: $\mu, \lambda_{1}, \lambda_{2}$. We would usually want both $\lambda_{1}$ and $\lambda_{2}$
-to be equal to 1 to indicate that we care both about the difference of intensities inside and outside. A smaller weighting for one of them would mean that we
-would ignore some of the variances in intensities in this area. A larger weighting means that we are magnifying the difference in intensities in that area
-(i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we vary
-$\lambda_1$ and $\lambda_2$. When we set $\lambda_1 = 0.2$ and $\lambda_2 = 1$, less details within the interior of the brain is detected. When
-we set $\lambda_1 = 2$ and $\lambda_2 = 1$, we can see that more details within the brain are detected. When we set $\lambda_1 = 5$, and $\lambda_2 = 0.01$ (a
-very extreme case), we can see that parts of outside boundaries are not well detected, while many details are detected within the brain.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=7cm]{threech.png}
-\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 * 200. $\phi_{0}(x,y) = - \sqrt{(x -
-100)^2 + (y - 100)^2)} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
-5. For intersection then union: cpu = 1.24 and iterations = 4.}
-\label{fig:threech}
-\end{figure}
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=7cm]{cv_eg12.png}
-\caption{Effect of varying $\lambda_1$ which controls how much details are detected inside the contour. The figure shows the final segmentation in each case.}
-\label{fig:cv_eg12}
-\end{figure}
-
-The second parameter, $\mu$ controls how much we allow the length of the curve to increase. If $\mu$ is small, it means that the curve length can increase to
-detect multiple smaller objects without penalizing our minimization problem. On the other hand if $\mu$ is large, it means that any change in the curve length
-will be scaled up, and thus will limit the curve expansion to keep the force to a minimum. Unfortunately, we were not able to see this effect in our
-experiments. We tried varying increasing $\mu$ to be able to detect groupings of objects instead of the individual objects, but the segmentation was invariant
-to $\mu$. We give more details about how we tried to handle this, and why we believe it is not working in Section~\ref{sec:difficulties}.
-
-\subsection{Sandberg-Chan Model Results}
-
-For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly included in the logic operations functionality. We therefore, show
-the other criteria. Unless otherwise specified, $\mu = 0.1$ and $\lambda = 255*255$ in the experiments below.
 
@@
 
-Figure~\ref{fig:donutlogic} shows a more complicated example for both the intersection and union with the donut image shown in the previous section. We used
-two versions of the donuts with each one having a different part occluded. The results of the union and intersection operators are shown in the figure.
-Additionally, this example shows that the Sandberg-Chan model can still successfully detect holes despite the addition of the logic operations.
-
 A practical usage of this logic operations framework is to recover missing parts from different images, and combine them to produce a more complete image. We
@@
 
-\begin{figure*}[t]
+Figure~\ref{fig:donutlogic} shows a more complicated example for both the intersection and union with the donut image shown in the previous section. We used
+two versions of the donut with each one having a different part occluded. The results of the union and intersection operators are shown in the figure.
+Additionally, this example shows that the Sandberg-Chan model can still successfully detect holes despite the addition of the logic operations.
+
+We also wanted to make sure that the way the intensities in each channel are defined does not
+matter. Figure~\ref{fig:diffint} shows the union and intersection performed on two channels of opposite intensities inside and outside the objects. The
+segmentation was indeed contrast invariant as claimed by the authors. Finally, we show that our implementation can be extended to more than two channels.
+Figure~\ref{fig:threech} shows the union, intersection, and the intersection of the union of three different channels.
+
+\subsubsection*{Independence of Initial Curve Position}
+
+\begin{figure}[h!]
+\centering
+\includegraphics[width=0.3\textwidth]{family.png}
+\caption{Combining missing information from two different channels. Size = 114 x 114. $\phi_{0}(x,y) = \sqrt{(x - 57)^2 + (y - 57)^2} -
+28.5$, no reinitialization. Union: cpu = 4.01 s and iterations = 7. Intersection: cpu = 4.12 s and iterations = 7.}
+\label{fig:family}
+\end{figure}
+
+\begin{figure*}[t!]
 \centering
 \includegraphics[width=12cm]{donutlogic.png}
-\caption{Performing union and intersection on images with holes. Size = 300 * 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
+\caption{Performing union and intersection on images with holes. Size = 300 x 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
 75$, no reinitialization. For union: cpu = 8.63 sec and iterations = 10. For intersection: cpu = 7.68 sec and iterations = 10.}
@@
 
-\begin{figure}[t]
+\begin{figure}[t!h!]
 \centering
-\includegraphics[width=7cm]{family.png}
-\caption{Combining iterations = 7. union 4.01 s intersection 4.12 s}
-\label{fig:donutlogic}
+\includegraphics[width=0.3\textwidth]{diffint.png}
+\caption{Segmentation for union and intersection of channels having reverse contrast.. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
+75$,  no reinitialization. For union: cpu = 1.32 sec and iterations = 2. For intersection: cpu = 1.83 sec and iterations = 2}
+\label{fig:diffint}
 \end{figure}
 
-\begin{figure*}[t]
+Similar to the previous model, we need to ensure that the position of the initial curve does not alter our results. Figure~\ref{fig:sc_position} shows
+the segmentation results obtained for the same image used in Figure~\ref{fig:sc_unionsimple}, but with a different position for the initial curve. Both
+segmentation are the same.
+
+\subsubsection*{Reasonable Performance with Blurred Images}
+
+Figure~\ref{fig:sc_blurry} shows the union performed on two blurry triangles. We blurred the triangles used in Figure~\ref{fig:threech}, and performed the
+union operation on them. The full triangle was segmented (the union) despite the blurred edges.
+
+\subsubsection*{Reasonable Performance with Noisy Images}
+
+To be able to perform well in noisy images, $\lambda$ must be decreased in order to ignore the noise. We added \textit{Salt \& pepper} noise with a variation of
+0.1 to
+the images shown in Figure~\ref{fig:sc_noisy}, and tried to perform the union. In the first case shown, we set $\lambda = 255*255$ (i.e $ max(u)^2$) which is
+the default setting
+used for clean images, and in the second case, we set $\lambda = 125$. We show the final segmentation for both cases. In the first case, the noise was detected,
+while in the second case, the noise was ignored since $\lambda$ was small. This shows that we also satisfy the response to parameter settings criteria since
+$\lambda$ was the only variation mentioned by the authors in the original paper~\cite{sandberg2005logic}. We note that the contour evolved much more slowly
+with a smaller $\lambda$, and thus took much more CPU time, and a larger number of iterations.
+
+\begin{figure}[t!]
 \centering
-\includegraphics[width=12cm]{flowers.png}
-\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
-Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
-size = cpu = 1.73 sec, iterations = 7.}
-\label{fig:flowers}
-\end{figure*}
-
-\subsubsection*{Independence of Initial Curve Position}
-
-Similar to the previous model, we needed to ensure that the position of the initial curve does not alter our results. Figure~\ref{fig:sc_position} shows
-the segmentation results obtained for the same image used in Figure~\ref{fig:sc_unionsimple}, but with a different position for the initial curve.
+\includegraphics[width=0.3\textwidth]{threech.png}
+\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 x 200. $\phi_{0}(x,y) = - \sqrt{(x -
+100)^2 + (y - 100)^2} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
+5. For intersection then union: cpu = 1.24 and iterations = 4.}
+\label{fig:threech}
+\end{figure}
 
@@
 
-\subsubsection*{Reasonable Performance with Blurred Images}
-
-Figure~\ref{fig:sc_blurry} shows the union performed on two blurry triangles. We blurred the triangles used in Figure~\ref{fig:threech}, and performed the
-union operation on them. The full triangle was segmented (the union) despite the blurred edges.
-
-\begin{figure*}[t]
-\centering
-\includegraphics[width=12cm]{cv_eg11.png}
-\caption{Successful union with blurred images. Top: the evolving curve for $A_1 \cup A_2$
-(in red) over time where the first
-image shows the initial
-contour. Bottom: evolving segmentation over time until the union is detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
-50$, no reinitialization, cpu = 6 s, took 11 iterations.}
-\label{fig:sc_blurry}
-\end{figure*}
-
-\subsubsection*{Reasonable Performance with Noisy Images}
-
-To be able to perform well in noisy images, $\lambda$ must be decreased in order to ignore the noise. We added 'Salt \& pepper' noise with a variation of 0.1 to
-the images shown in Figure~\ref{fig:sc_noisy}, and tried to perform the union. In the first case shown, we set $\lambda = 255*255$ which is the normal setting
-used for clean images, and in the second case, we set $\lambda = 25$. We show the final segmentation for both cases. In the first case, the noise was detected,
-while in the second case, the noise was ignored since $\lambda$ was small. This shows that we also satisfy the response to parameter settings criteria since
-$\lambda$ was the only variation mentioned by the authors in the original paper~\cite{sandberg2005logic}. We note that the contour evolved much more slowly
-with a smalled $\lambda$, and thus took much more CPU time, and a larger number of iterations.
-
-\begin{figure}[t]
-\centering
-\includegraphics[width=7cm]{sc_noisy.png}
-\caption{Same image as Figure~\ref{fig:threech}, but with 'salt \& pepper' noise with a variation of 0.1. Decreasing $\lambda$ from 255*255 to 125 allowed less
-noise to be detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
-50$, no reinitialization. Cpu = 3.17 sec and iterations = 5 when $\lambda = 255*255$. Cpu = 146.46 sec and iterations = 125 when $\lambda =
-125$.}
-\label{fig:sc_noisy}
-\end{figure}
-
 \subsubsection*{Complement}
@@
 Unfortunately, the complement did not work properly in our implementation. We closely followed the definitions in the original paper where $z_i^{in'} = 1 -
-z_i^{in}$ and $z_i^{out'} = 1 - z_i^{out}$. The results was actually successfully detected, but continuous flickering occurred, and the curve did not stop in
-one position. We discuss this results in more details in Section~\ref{sec:difficulties}.
+z_i^{in}$ and $z_i^{out'} = 1 - z_i^{out}$. The result was actually successfully detected, but continuous flickering occurred, and the curve did not stop
+evolving We discuss this results in more details in Section~\ref{sec:difficulties}.
 
@@
 the image in Figure~\ref{fig:cv_eg7} which the Chan-Vese model was not able to completely segment. We extracted three channels from the colored image: Red,
-Green, Black, and performed the union on the three channels. The results are shown in Figure~\ref{fig:flowers}. The complete contour of the flowers was
+Green, Blue, and performed the union on the three channels. The results are shown in Figure~\ref{fig:flowers}. The complete contour of the flowers was
 successfully detected unlike that obtained in Figure~\ref{fig:cv_eg7}.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.4\textwidth]{sc_blurry.png}
+\caption{Successful union with blurred images. Top: the evolving curve for $A_1 \cup A_2$
+(in red) over time where the first
+image shows the initial
+contour. Bottom: evolving segmentation over time until the union is detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
+50$, no reinitialization, cpu = 6 s, iterations = 11.}
+\label{fig:sc_blurry}
+\end{figure}
 
@@
 Therefore, varying the value of $\mu$ will affect the ability to segment objects in noisy images, especially when the noise is similar to Matlab's \textit{salt
-and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure [\ref{fig:cvmu}] shows the
-segmentation of an object in a noisy for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting that
+and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure~\ref{fig:cvmu} shows the
+segmentation of an object in a noisy image for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting
+that
 this challenge kept us busy for over two days, after which we decided we should move on to other parts of the project.
 
-\begin{figure*}[t]
+\begin{figure*}[t!]
 \centering
-\includegraphics[width=\textwidth]{cvmu.png}
+\includegraphics[width=12cm]{flowers.png}
+\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
+Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
+size = cpu = 1.73 sec, iterations = 7.}
+\label{fig:flowers}
+\end{figure*}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.4\textwidth]{sc_noisy.png}
+\caption{Union of noisy channels. Same image as Figure~\ref{fig:threech}, but with \textit{Salt \& pepper'} noise with a variation of 0.1. Decreasing $\lambda$
+from 255*255 to 125 allowed less
+noise to be detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
+50$, no reinitialization. Cpu = 3.17 sec and iterations = 5 when $\lambda = 255*255$. Cpu = 146.46 sec and iterations = 125 when $\lambda =
+125$.}
+\label{fig:sc_noisy}
+\end{figure}
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.5\textwidth]{cvmu.png}
 \caption{$\mu$ does not have any effect on the length of the contour. \textit{Salt and pepper} noise was added to the original image for testing
@@
 \label{fig:cvmu}
-\end{figure*}
+\end{figure}
 
-\subsection{Computing the complement}
+\subsection{Computing the Complement}
 The Sandberg-Chan model supports the complement logic operator, so one can segment the logic object ($A \cap \neg B$) (i.e. segment what is in channel A and not
@@
 \label{eqn:complement}
-\begin{align*}
+%\begin{align*}
 {z_{i}}^{in^{'}} = 1 - {z_{i}}^{in}, \\
 {z_{i}}^{out^{'}} = 1 - {z_{i}}^{out}.
-\end{align*}
+%\end{align*}
 \end{equation}
 
-Although the definition is very simple, it does not work in practice. We implemented the complement logic operator as defined in Equation
-[\ref{eqn:complement}], and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
-flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure [\ref{fig:sccomp}] shows this scenrio. We tried to
+Although the definition is very simple, it did not work in practice. We implemented the complement logic operator as defined in Equation~\ref{eqn:complement},
+and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
+flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure~\ref{fig:sccomp} shows this scenario. We tried to
 debug the reasons for this problem for three days.
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{sccomp.png}
+\includegraphics[width=0.25\textwidth]{sccomp.png}
 \caption{The complement operator causes the contour to go through an infinite loop between two segmentations: the correct segmentation, and the segmentation
@@
 \label{fig:sccomp}
-\end{figure*}
+\end{figure}
 
-\subsection{Defining the initial contour ($\phi_0$)}
-Both the Chan-Vese model and the Sandberg-Chan model defines the initial contour $\phi_0$ as a signed distance function where $\phi_0$ is positive inside the
-contour $C$, zero on the boundary of $C$ ($\partial C$), and negative outside the contour $C$. Figure [\ref{fig:phinote}] illustrates those assumptions.
+\subsection{Defining the initial contour $phi_0$}
+Both the Chan-Vese model and the Sandberg-Chan model define the initial contour ($\phi_0$) as a signed distance function where $\phi_0$ is positive inside the
+contour $C$, zero on the boundary of $C$ ($\partial C$), and negative outside the contour $C$. Figure~\ref{fig:phinote} illustrates those assumptions.
 
@@
 \centering
-\includegraphics[width=\textwidth]{phinote.png}
+\includegraphics[width=0.3\textwidth]{phinote.png}
 \caption{Assumptions about the way $\phi_0$ is defined.}
@@
 union). After spending quite sometime to resolve this problem and checking what might be the source of the error, we tried using the regular equation of a
-circle to initialize $\phi_0$ (Equation [\ref{eq:phinotenormal]}). We were very intrigued to notice that this modification resulted in the correct segmentation
-of the input channels. Figure [\ref{fig:phidef}] shows the final segmentation using both definitions for $\phi_0$.
+circle to initialize $\phi_0$ (Equation~\ref{eqn:phinotenormal}). We were very intrigued to notice that this modification resulted in the correct segmentation
+of the input channels. Figure~\ref{fig:phidef} shows the final segmentation using both definitions for $\phi_0$.
 
@@
 
-\begin{figure*}[t]
+\begin{figure}[t]
 \centering
-\includegraphics[width=\textwidth]{phidef.png}
-\caption{}
+\includegraphics[width=0.4\textwidth]{phidef.png}
+\caption{Union and intersection depending on the way $\phi$ is defined}
 \label{fig:phidef}
-\end{figure*}
+\end{figure}
 
 \subsection{Mistyped Equations}
-As previously mentiond in Section [\ref{sec:sandberg-chan}], there is a typo in the equations that define $z_in$ and $z_out$ in the Sandberg-Chan model
-(equation 3 in thier paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
+As previously mentioned in Section~\ref{sec:sandberg-chan}, there is a typo in the equations that define $z_{in}$ and $z_{out}$ in the Sandberg-Chan model
+(equation 3 in their paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
 typos consumed over a week of our time until we were able to find out what are the correct equations in both cases. We could have spent this time overcoming
@@
 \label{sec:concl}
+In this project, we have provided an implementation for the Chan-Vese model for active contours without edges and the Sandberg-Chan logic framework for
+multi-channel images, both written in Matlab. Our experiments have shown that the Chan-Vese model is effective for various types of images. In specific, it is
+useful when an edge-based segmentation is not applicable (i.e. edges do not depend on a gradient). However, our implementation fails to respond to different
+values for the contour length scaling parameter, $\mu$. This affects the ability of our implementation to handle Matlab's \textit{Salt \& pepper} noise. On the
+other hand, changing the values for the intensity weight parameters $\lambda_1$ and $\lambda_2$ affects the amount of details obtained inside and outside the
+segmented object respectively.
 
-\bibliographystyle{IEEEtran}
+We have also shown that our implementation for the Sandberg-Chan logic framework handles different logic combinations for multiple channels. We have also
+experimented using it as a segmentation tool for colored images by modeling an input colored images as a 3-channels (red, green, blue). However, our
+implementation for the \textit{complement} logic operation seems to be missing a hidden trick in the Sandberg-Chan logic framework.
+
 \bibliography{references}
+\bibliographystyle{ieee}
+
+\appendix
+\appendixpage
+
+\section{Division of Labor}
+
+We both implemented the Chan-Vese model together since this was the starting point of the project. Sarah did most of the initial implementation for the
+Sandberg-Chan model by defining all the necessary functionalities for the logic operations for two images. Karim then extended it to multiple images. Karim
+also tested the Sandberg-Chan model, and took most of the snapshots in that part for the paper. Sarah tested the Chan-Vese model, and took all its snapshots
+for the paper. We both worked on finding an appropriate stopping condition for the contour evolution, and invested a lot of time to debug the problems we
+faced that were described in Section~\ref{sec:difficulties}.
 

### 2010-12-13 cs870/project/Report/Paper.tex
final draft khalas isA :)

@@
 \usepackage{url}
+\usepackage{appendix}
 
@@
 differs
-from other active contour models in that it does not depend on the gradient of the contour ($\phi$), therefore it is more capable of detecting objects whose
+from other active contour models in that it is not edge dependent, therefore it is more capable of detecting objects whose
 boundaries may not be defined by a gradient (e.g. blurry images, noisy images). The Sandberg-Chan model builds on top of the Chan-Vese model by allowing the
@@
 classical Active Contour model (or Snakes) proposed by Kass et al.~\cite{kass1988snakes} was the first model to use the idea of energy minimization to attract
-a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it very highly adopted later on
-(E.g.~\cite{caselles1997geodesic}. Most of these models used the level set formulation for propagating fronts to evolve the curve. However, these models highly
-depended on curvature motion (motion defined by the gradient of the curve) which led to poor performance in smoothed edges. To overcome these limitations, Chan
+a contour to the edges of the objects in an image. The Snakes model was very successful and variations of it were highly adopted later on
+(E.g.~\cite{caselles1997geodesic}. Most of these models use the level set formulation for propagating fronts to evolve the curve. However, these models highly
+depend on motion defined by the gradient of the image which leads to poor performance in smoothed edges. To overcome these limitations, Chan
 and Vese~\cite{chan2001active} propose an active contour model that does not depend on the edges (i.e the gradient) for propagating the curve to detect the
@@
 
-The Active Contours without Edges model proposed by Chan and Vese (referred to as Chan-Vese model throughout the rest of this paper) was very successful in
-detecting objects even in noisy or blurry images. It could also detect holes in objects which was usually a limitation in previous models. As an extension to
+The Active Contours without Edges model proposed by Chan and Vese (referred to as Chan-Vese model throughout the rest of this paper) is very successful in
+detecting objects even in noisy or blurry images. It can also detect holes in objects which is usually a limitation in previous models. As an extension to
 this work, Sandberg and Chan~\cite{sandberg2005logic} propose a logic framework (referred to as the Sandberg-Chan model throughout the rest of this
-paper) that  performs logical operations on multiple images according to the curve propagation proposed in the Chan-Vese model. To achieve that, previous models
+paper) that  performs logic operations on multiple images according to the curve propagation proposed in the Chan-Vese model. To achieve that, previous models
 usually had two steps. They would either first segment the object in each channel separately then combine the segmented objects according to the logic operation
-through bitwise operations (E.g.~\cite{sapiro1997color, zhu2002region}) or they would apply logic operations to the different images then segment the resulting
+through bitwise operations~\cite{sapiro1997color, zhu2002region} or they would apply logic operations to the different images then segment the resulting
 image. The first approach is
@@
 
-In this paper, we report on our findings after implementing each of these models. We implemented both models in Matlab, and experimented with several images.
-In this paper, we explain the details of our implementation as well as our results. The rest of this paper is organized as follows: Section~\ref{sec:bg} first
-provides brief background about each of the two models implemented in this paper. Section~\ref{sec:implementation} then explains our implementation. In this
-section, we explain how we implemented the models and any variations from the original papers. Section~\ref{sec:results} explains our evaluation criteria, and
-presents the results we obtained. Section~\ref{sec:difficulties} discusses some of the difficulties we faced, and points out some BLA. Section~\ref{sec:concl}
+In this paper, we report on our findings after implementing each of these models. We implement both models in Matlab, and experiment with several images.
+We explain the details of our implementation as well as our results. The rest of this paper is organized as follows:
+Section~\ref{sec:bg} provides brief background about each of the two models implemented in this paper. Section~\ref{sec:implementation} then explains our
+implementation. In this section, we explain how we implemented the models and any variations from the original papers. Section~\ref{sec:results} explains our
+evaluation criteria, and
+presents the results we obtained. Section~\ref{sec:difficulties} discusses some of the difficulties we faced during implementation. Section~\ref{sec:concl}
 concludes this paper by summarizing our findings.
@@
 
-The parameter $\mu$ is a scaling parameter for the length of the curve represented in terms of curvature. The smalled $\mu$ is, the more the length of the curve
-can increase without penalizing the minimization. This allows the model to detect smalled objects and holes. The larger $\mu$ is, the less freedom there is for
+The parameter $\mu$ is a scaling parameter for the length of the curve represented in terms of curvature. The smaller $\mu$ is, the more the length of the curve
+can increase without penalizing the minimization. This allows the model to detect smaller objects and holes. The larger $\mu$ is, the less freedom there is for
 the curve to increase in length, and thus, it will only be able to detect larger objects. The parameter $\nu$ is also a scaling term for the area of the curve.
@@
 efficiently. For example, finding the union of two images where different parts of the object are occluded in each image so that a complete object can be
-obtained. In order to do that, they use the Chan-Vese model to detect objects, and simultaneously include the logic operations in the minimization problem.
+obtained. This example will be shown later in Figure~\ref{fig:family}. In order to do that, they use the Chan-Vese model to detect objects, and simultaneously
+include the logic operations in the minimization problem.
 Since they are using the Chan-Vese model, the problem must be viewed in terms of regions as well. Accordingly, they define the two main logic operations
-(union and intersection) in terms of regions where the union of two images is the union of the insides of the curve with respect to the images plus the
-intersection of the outsides of the curve with respect to the images. For example, Figure~\ref{fig:logic-op} (taken from the original paper) shows that the
+(union and intersection) in terms of regions where the union of two images is the union of the inside of the curve with respect to the images plus the
+intersection of the outside of the curve with respect to the images. For example, Figure~\ref{fig:logic-op}~\cite{sandberg2005logic} shows that the
 union of $A_{1}$ and $A_{2}$ (in terms of its shape) can be obtained by taking the union of the insides of the object or the intersection of the outsides.
 Similarly, the intersection of two images would be the intersection of the insides plus the union of the outsides. Thus, in order to obtain accurate results
-irrespective of the object's intensity outside versus inside, we should consider \textit{both} the logical operation required for the region inside the object
+irrespective of the object's intensity outside versus inside, we should consider \textit{both} the logic operation required for the region inside the object
 as well as that outside.
 
-More formally, the authors define two logic variables $z_{i}^{in}$ and $z_{i}^{out}$ to denote whether a point $(x,y)$ should be in the moving curve C or not
-with respect to image i. Since we are trying to minimize the fitting of the curve, they use 0 to denote true and 1 to denote false (the reverse of the usual
-convention. Following the Chan-Vese model, they represent each of the regions inside and outside the curve by a constant c. In this paper, they represent
+More formally, the authors define two logic variables $z_{i}^{in}$ and $z_{i}^{out}$ to denote whether a point $(x,y)$ should be in the moving curve $C$ or not
+with respect to image $i$. Since we are trying to minimize the fitting of the curve, they use $0$ to denote $true$ and $1$ to denote $false$ (the reverse of the
+usual
+convention). Following the Chan-Vese model, they represent each of the regions inside and outside the curve by a constant $c$. In this paper, they represent
 $c_{in}$ as $c_{+}$ and $c_{out}$ as $c_{-}$. Equations~\ref{eqn:zin} and~\ref{eqn:zout} show how to calculate $z_{in}$ and $z_{out}$ respectively in terms of
@@
 Based on these functions, the Euler-Lagrange equation for any logic operation is shown in Equation~\ref{eqn:svpde}. According to the desired logic operation,
-$f_{in}$ and $f_{out}$ will be specified accordingly. For example, if we are doing the union of the images, then we will need to perform a union operation on
+$f_{in}$ and $f_{out}$ will be specified. For example, if we are doing the union of the images, then we will need to perform a union operation on
 the insides. Thus, $f_{in}$ will be replaced by Equation~\ref{eqn:funion}. Similarly, we will need to do the intersection of the outsides so $f_{out}$ will be
@@
 
-During implementation, we tried to stick to the authors' formulas and guidelines as much as possible. In this section we explain the choices we made regardless
-implementation. For the chan-vese model, we used the PDE given in Equation 9 in the original paper~\cite{chan2001active} which is shown in
-Equation~\ref{eqn:cvpde}.  We, first, tried to implement the delta dirac function, $\delta_{\varepsilon}\phi$, as defined in the paper. However, we did not
-get non-zero values everywhere as indicated by the authors. Accordingly, we chose to use $|\bigtriangledown\phi|$ instead of the delta dirac function as this
+During implementation, we tried to stick to the authors' formulas and guidelines as much as possible. In this section we explain what methods we use to
+numerically solve the PDEs. For the chan-vese model, we use the PDE given in Equation 9 in the original paper~\cite{chan2001active} which is shown in
+Equation~\ref{eqn:cvpde}.  We, first, tried to implement the dirac delta function, $\delta_{\varepsilon}\phi$, as defined in the paper. However, it resulted in
+worse segmentations. Accordingly, we chose to use $|\bigtriangledown\phi|$ instead of the dirac delta function as this
 was indicated as a valid alternative by the authors. To calculate $c_{1}$ and $c_{2}$, we simply calculated the mean of the values inside $\phi$ (specifically
 where $\phi > 0$) and the mean of the values outside $\phi$ (specifically where $\phi < 0$) respectively. To actually solve the PDE, we chose to use an explicit
-time stepping scheme for the finite difference discretization. That is, in each time step $\phi_{t} = \phi + \triangle t*force$. This was simpler to implement,
-and although $\triangle t$ must be chosen carefully to satisfy the stability condition, we did not suffer from performance problems due to this restriction.
+time stepping scheme for the finite difference discretization. That is, in each time step $\phi^{n+1} = \phi^n + \triangle t*phi_t$. This was simpler to
+implement, and although $\triangle t$ must be chosen carefully to satisfy the stability condition ($\triangle t <= (1/imageSize)^2$), we did not suffer from
+performance problems due to this restriction.
 
-For the sandberg-chan model, we used the PDE given in original paper as well, and shown in Equation~\ref{eqn:svpde}. Again, we used $|\bigtriangledown\phi|$
-instead of the dirac function. We calculated $z_{in}$ and $z_{out}$ according to Equations~\ref{eqn:zin, eqn:zout} shown above which have been corrected for
+For the Sandberg-Chan model, we used the PDE given in original paper as well, and shown in Equation~\ref{eqn:svpde}. Again, we used $|\bigtriangledown\phi|$
+instead of the dirac delta function. We calculated $z_{in}$ and $z_{out}$ according to Equations~\ref{eqn:zin} and~\ref{eqn:zout} shown above which have been
+corrected for
 the typo in the original paper. For all other equations and calculations necessary in the model, we closely followed the original paper in our implementation.
 We also used an explicit time stepping scheme in this model.
+
+We use the same stopping condition for both model. That is, to check if we have reached the stationary solution, we check if more points have been added inside
+the contour or not. More formally, we check if $(\phi^{n+1} > 0) == (\phi_{n} > 0)$. If the contour did not move, then new points will be added to its inside,
+and this condition will be satisfied.
 
@@
 cases, and incrementally challenging the model. Unless otherwise stated, all the segmentation was performed on a gray scale version of the original image. All
-the images in the Chan-Vese model were run on a BLA COMPUTER and all the images in the Sandberg-Chan model were run on a BLA computer. As shown in the results,
+the images in the Chan-Vese model were run on an Intel Core 2 Duo processor with 2.00 GHz and a 2GB RAM and all the images in the Sandberg-Chan
+model were run on a Intel Centrino Duo 2.0 processor and a 1GB RAM computer. As shown in the results,
 we tried different image sizes to ensure that our solution converges.
@@
 \caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
-0.01$, no reinitialization, cpu = 2.9s, 7 iterations.}
+Bottom: evolving segmentation over time until the objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$, $\mu =
+0.01$, no reinitialization, cpu = 2.9s, iterations = 7.}
 \label{fig:cv_eg1}
@@
 \caption{Successful detection of object boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2)} + 32.8$, $\mu =
-0.01$, no reinitialization, cpu = 1.95 s, took 7 iterations.}
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.6)^2 + (y - 65.5)^2} + 32.8$, $\mu =
+0.01$, no reinitialization, cpu = 1.95 s, iterations = 7.}
 \label{fig:cv_eg2}
@@
 successfully able to detect the boundaries of both objects. Figure~\ref{fig:cv_eg2} shows a slightly more complicated example where the evolving contour
-successfully detects the contour of the brain. Note that there is a tiny area to the bottom left of the brain that was also successfully detected. To ensure
-that open boundaries (e.g. lines) can also be detected, we used an image with several shapes shown in Figure~\ref{fig:cv_eg8};
+successfully detects the contour of the brain. Note that there is a tiny area near the center to the left of the brain that was also successfully detected. To
+ensure
+that open boundaries (e.g. lines) can also be detected, we used an image with several shapes shown in Figure~\ref{fig:cv_eg8}.
 
@@
 \caption{Successful detection of object with open boundaries. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, $\mu =
-0.01$, no reinitialization, cpu = 5.19 s, took 11 iterations.}
+Bottom: evolving segmentation over time until all objects are detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$, $\mu =
+0.01$, no reinitialization, cpu = 5.19 s, iterations = 11.}
 \label{fig:cv_eg8}
@@
 \caption{Initial curve position partially overlapping object. Top: the evolving curve (in red) over time where the first image shows the initial contour.
-Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2)} + 20$, $\mu =
-0.01$, no reinitialization, cpu = 2.26 s, took 9 iterations.}
+Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 30)^2 + (y - 30)^2} + 20$, $\mu =
+0.01$, no reinitialization, cpu = 2.26 s, iterations = 9.}
 \label{fig:cv_eg3}
@@
 \caption{Initial curve position not overlapping any area of the object. Top: the evolving curve (in red) over time where the first image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2)} + 10$,
-$\mu =0.01$, no reinitialization, cpu = 2.01 s, took 9 iterations.}
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 120)^2 + (y - 120)^2} + 10$,
+$\mu =0.01$, no reinitialization, cpu = 2.01 s, iterations = 9.}
 \label{fig:cv_eg4}
@@
 
-To ensure that the position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the lesame brain image
+To ensure that the initial position of the curve does not affect the final segmentation, we tested two other positions for the initial curve for the same
+brain image
 used in Figure~\ref{fig:cv_eg2}. In Figure~\ref{fig:cv_eg2}, the initial contour was completely overlapping the object. Accordingly, we tried two other
@@
 \caption{Successful detection of holes.  Top: the evolving curve (in red) over time where the first image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.5)^2 + (y - 65.5)^2)} + 32.8$,
-$\mu =0.01$, no reinitialization, cpu = 8.12 s, took 12 iterations.}
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 131 x 131, $\phi_{0}(x,y) = - \sqrt{(x - 65.5)^2 + (y - 65.5)^2} + 32.8$,
+$\mu =0.01$, no reinitialization, cpu = 8.12 s, iterations = 12.}
 \label{fig:cv_eg5}
@@
 \caption{Successful detection of multiple holes.  Top: the evolving curve (in red) over time where the first image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
-$\mu =0.01$, no reinitialization, cpu = 15.07 s, took 23 iterations.}
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 15.07 s, iterations =  23.}
 \label{fig:cv_eg6}
@@
 \includegraphics[width=0.5\textwidth]{cv_eg7.png}
-\caption{Inability to detect low intensities that have not much variation from their background.  Top: the evolving curve (in red) over time where the first
+\caption{Inability to detect low intensities that do not have much variation from their background.  Top: the evolving curve (in red) over time where the
+first
 image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$,
-$\mu =0.01$, no reinitialization, cpu = 1.35 s, took 6 iterations.}
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} + 75$,
+$\mu =0.01$, no reinitialization, cpu = 1.35 s, iterations = 6.}
 \label{fig:cv_eg7}
@@
 that of the background This is a known drawback of the Chan-Vese model since it only divides the image into two regions, and requires a big variation in
-intensities for the curve to change. Another example showing this limitation is shown in Figure~\ref{fig:cv_eg7} where the flower parts with very low
+intensities for the curve to change. Figure~\ref{fig:cv_eg6} shows that multiple holes in an image can also be successfully detected. Another example showing
+the inability to detect objects with intensities close to that of the background is shown in Figure~\ref{fig:cv_eg7} where the flower parts with very low
 intensities are not detected. We will show that we can overcome this limitation with the Sandberg-Chan model for this particular example.
-Figure~\ref{fig:cv_eg6} shows that multiple holes in an image can also be successfully detected.
 
@@
 
-Figure~\ref{fig:cv_eg10} shows how two objects in a noisy image were successfully detected. In this image, we added Gaussian noise with mean zero and 0.01
+Figure~\ref{fig:cv_eg10} shows how two objects in a noisy image were successfully detected. In this image, we added a Gaussian noise with mean = 0 and
+variation = 0.01
 variance using Matlab's built-in noise function. The image shows that although some of the noise was detected in intermediate iterations, the contour continued
 to evolve until it was only surrounding the desired objects. Unfortunately, the effect of noise was not completely ignored in all cases. For example,
-Figure~\ref{fig:cv_eg11} shows that for the same image, but with 'salt \& pepper' noise, the noise was detected as objects. This should not have been the case
-since we used a large value $\mu = 5$ for the length scaling parameter. However, for some reason, varying $\mu$ did not have the intended effect as explained
-in the next section.
+Figure~\ref{fig:cv_eg11} shows that for the same image, but with \textit{Salt \& pepper} noise, the noise was detected as objects. This should not have been the
+case
+since we used a large value $\mu = 5$ for the length scaling parameter. However, varying $\mu$ does not have the intended effect as explained
+in the next paragraph.
 
@@
 image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
-75$, $\mu =0.01$, no reinitialization, cpu = 4.4 s, took 7 iterations.}
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} +
+75$, $\mu =0.01$, no reinitialization, cpu = 4.4 s, iterations = 7.}
 \label{fig:cv_eg10}
@@
 \includegraphics[width=12cm]{cv_eg11.png}
-\caption{Same image as Figure~\ref{fig:cv_eg10}, but with 'salt \& pepper' noise. Noise was still detected despite increasing $\mu$.  Top: the evolving curve
+\caption{Figure~\ref{fig:cv_eg10}, but with \textit{Salt \& pepper} noise. Noise was still detected despite increasing $\mu$.  Top: the evolving
+curve
 (in red) over time where the first
 image shows the initial
-contour.Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
-75$, $\mu = 5$, no reinitialization, cpu = 12.27 s, took 7 iterations.}
+contour. Bottom: evolving segmentation over time until the object is detected. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2} +
+75$, $\mu = 5$, no reinitialization, cpu = 12.27 s, iterations = 7.}
 \label{fig:cv_eg11}
@@
 to be equal to 1 to indicate that we care both about the difference of intensities inside and outside. A smaller weighting for one of them would mean that we
-would ignore some of the variances in intensities in this area. A larger weighting means that we are magnifying the difference in intensities in that area
+would ignore some of the variation in intensities in this area. A larger weighting means that we are magnifying the difference in intensities in that area
 (i.e. we want to detect any minor changes). We show an example of these variations in Figure~\ref{fig:cv_eg12} where we vary
@@
 
-\begin{figure}[t!]
+The second parameter, $\mu$ controls how much we allow the length of the curve to increase. If $\mu$ is small, it means that the curve length can increase to
+detect multiple smaller objects without penalizing our minimization problem. On the other hand if $\mu$ is large, it means that any change in the curve length
+will be scaled up, and thus will limit the curve expansion to keep the force to a minimum. Unfortunately, we were not able to see this effect in our
+experiments. We tried varying $\mu$ to be able to detect groupings of objects instead of the individual objects, but the segmentation was invariant
+to $\mu$. We give more details about how we tried to handle this, and why we believe it is not working in Section~\ref{sec:difficulties}.
+
+\subsection{Sandberg-Chan Model Results}
+
+For the Sandberg-Chan model, the successful detection of boundaries criterion is implicitly included in the logic operations functionality. We therefore, show
+the other criteria. Unless otherwise specified, $\mu = 0.1$ and $\lambda = 255*255$ in the experiments below.
+
+\begin{figure*}[t!]
 \centering
@@
 \label{fig:cv_eg12}
-\end{figure}
-
-The second parameter, $\mu$ controls how much we allow the length of the curve to increase. If $\mu$ is small, it means that the curve length can increase to
-detect multiple smaller objects without penalizing our minimization problem. On the other hand if $\mu$ is large, it means that any change in the curve length
-will be scaled up, and thus will limit the curve expansion to keep the force to a minimum. Unfortunately, we were not able to see this effect in our
-experiments. We tried varying increasing $\mu$ to be able to detect groupings of objects instead of the individual objects, but the segmentation was invariant
-to $\mu$. We give more details about how we tried to handle this, and why we believe it is not working in Section~\ref{sec:difficulties}.
-
-\subsection{Sandberg-Chan Model Results}
-
-For the Sandberg-Chan model, the successful detection of boundaries criteria is implicitly included in the logic operations functionality. We therefore, show
-the other criteria. Unless otherwise specified, $\mu = 0.1$ and $\lambda = 255*255$ in the experiments below.
+\end{figure*}
 
@@
 
+Figure~\ref{fig:donutlogic} shows a more complicated example for both the intersection and union with the donut image shown in the previous section. We used
+two versions of the donut with each one having a different part occluded. The results of the union and intersection operators are shown in the figure.
+Additionally, this example shows that the Sandberg-Chan model can still successfully detect holes despite the addition of the logic operations.
+
+We also wanted to make sure that the way the intensities in each channel are defined does not
+matter. Figure~\ref{fig:diffint} shows the union and intersection performed on two channels of opposite intensities inside and outside the objects. The
+segmentation was indeed contrast invariant as claimed by the authors. Finally, we show that our implementation can be extended to more than two channels.
+Figure~\ref{fig:threech} shows the union, intersection, and the intersection of the union of three different channels.
+
 \subsubsection*{Independence of Initial Curve Position}
 
-\begin{figure*}[t!h!]
+\begin{figure*}[t!]
 \centering
-\includegraphics[width=12cm]{donutlogic.png}
-\caption{Performing union and intersection on images with holes. Size = 300 * 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
+\includegraphics[width=0.3\textwidth]{family.png}
+\caption{Combining missing information from two different channels. Size = 114 x 114. $\phi_{0}(x,y) = \sqrt{(x - 57)^2 + (y - 57)^2} -
+28.5$, no reinitialization. Union: cpu = 4.01 s and iterations = 7. Intersection: cpu = 4.12 s and iterations = 7.}
+\label{fig:family}
+\end{figure*}
+
+\begin{figure*}[t!]
+\centering
+\includegraphics[width=0.5\textwidth]{donutlogic.png}
+\caption{Performing union and intersection on images with holes. Size = 300 x 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
 75$, no reinitialization. For union: cpu = 8.63 sec and iterations = 10. For intersection: cpu = 7.68 sec and iterations = 10.}
@@
 
-\begin{figure}[t!h!]
+\begin{figure}[h!]
 \centering
-\includegraphics[width=0.3\textwidth]{family.png}
-\caption{Combining missing information from two different channels. Size = 114 * 114. $\phi_{0}(x,y) = - \sqrt{(x - 57)^2 + (y - 57)^2)} +
-28.5$, no reinitialization. Union: cpu = 4.01 s and iterations = 7. Intersection: cpu = 4.12 s and iterations = 7.}
-\label{fig:family}
+\includegraphics[width=0.3\textwidth]{diffint.png}
+\caption{Segmentation for union and intersection of channels having reverse contrast.. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
+75$,  no reinitialization. For union: cpu = 1.32 sec and iterations = 2. For intersection: cpu = 1.83 sec and iterations = 2}
+\label{fig:diffint}
 \end{figure}
 
-Similar to the previous model, we needed to ensure that the position of the initial curve does not alter our results. Figure~\ref{fig:sc_position} shows
-the segmentation results obtained for the same image used in Figure~\ref{fig:sc_unionsimple}, but with a different position for the initial curve.
+Similar to the previous model, we need to ensure that the position of the initial curve does not alter our results. Figure~\ref{fig:sc_position} shows
+the segmentation results obtained for the same image used in Figure~\ref{fig:sc_unionsimple}, but with a different position for the initial curve. Both
+segmentation are the same.
+
+\subsubsection*{Reasonable Performance with Blurred Images}
+
+Figure~\ref{fig:sc_blurry} shows the union performed on two blurry triangles. We blurred the triangles used in Figure~\ref{fig:threech}, and performed the
+union operation on them. The full triangle was segmented (the union) despite the blurred edges.
+
+\subsubsection*{Reasonable Performance with Noisy Images}
+
+To be able to perform well in noisy images, $\lambda$ must be decreased in order to ignore the noise. We added \textit{Salt \& pepper} noise with a variation of
+0.1 to
+the images shown in Figure~\ref{fig:sc_noisy}, and tried to perform the union. In the first case shown, we set $\lambda = 255*255$ (i.e $ max(u)^2$) which is
+the default setting
+used for clean images, and in the second case, we set $\lambda = 125$. We show the final segmentation for both cases. In the first case, the noise was detected,
+while in the second case, the noise was ignored since $\lambda$ was small. This shows that we also satisfy the response to parameter settings criteria since
+$\lambda$ was the only variation mentioned by the authors in the original paper~\cite{sandberg2005logic}. We note that the contour evolved much more slowly
+with a smaller $\lambda$, and thus took much more CPU time, and a larger number of iterations.
+
+\begin{figure}[t!]
+\centering
+\includegraphics[width=0.3\textwidth]{threech.png}
+\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 x 200. $\phi_{0}(x,y) = - \sqrt{(x -
+100)^2 + (y - 100)^2} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
+5. For intersection then union: cpu = 1.24 and iterations = 4.}
+\label{fig:threech}
+\end{figure}
 
@@
 
-\subsubsection*{Reasonable Performance with Blurred Images}
+\subsubsection*{Complement}
 
-Figure~\ref{fig:sc_blurry} shows the union performed on two blurry triangles. We blurred the triangles used in Figure~\ref{fig:threech}, and performed the
-union operation on them. The full triangle was segmented (the union) despite the blurred edges.
+Unfortunately, the complement did not work properly in our implementation. We closely followed the definitions in the original paper where $z_i^{in'} = 1 -
+z_i^{in}$ and $z_i^{out'} = 1 - z_i^{out}$. The result was actually successfully detected, but continuous flickering occurred, and the curve did not stop
+evolving We discuss this results in more details in Section~\ref{sec:difficulties}.
+
+\subsubsection*{Other Tests}
+
+Since the Sandberg-Chan model can successfully detect objects in one channel which are not in another when performing the union, we were curious to know
+whether this would allow us to perform successful segmentation on colored images by combining the information in the three channels R,G,B. Accordingly, we used
+the image in Figure~\ref{fig:cv_eg7} which the Chan-Vese model was not able to completely segment. We extracted three channels from the colored image: Red,
+Green, Blue, and performed the union on the three channels. The results are shown in Figure~\ref{fig:flowers}. The complete contour of the flowers was
+successfully detected unlike that obtained in Figure~\ref{fig:cv_eg7}.
 
@@
 \centering
-\includegraphics[width=12cm]{cv_eg11.png}
+\includegraphics[width=0.3\textwidth]{sc_blurry.png}
 \caption{Successful union with blurred images. Top: the evolving curve for $A_1 \cup A_2$
@@
 contour. Bottom: evolving segmentation over time until the union is detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
-50$, no reinitialization, cpu = 6 s, took 11 iterations.}
+50$, no reinitialization, cpu = 6 s, iterations = 11.}
 \label{fig:sc_blurry}
@@
 
-\subsubsection*{Reasonable Performance with Noisy Images}
-
-To be able to perform well in noisy images, $\lambda$ must be decreased in order to ignore the noise. We added 'Salt \& pepper' noise with a variation of 0.1 to
-the images shown in Figure~\ref{fig:sc_noisy}, and tried to perform the union. In the first case shown, we set $\lambda = 255*255$ which is the normal setting
-used for clean images, and in the second case, we set $\lambda = 25$. We show the final segmentation for both cases. In the first case, the noise was detected,
-while in the second case, the noise was ignored since $\lambda$ was small. This shows that we also satisfy the response to parameter settings criteria since
-$\lambda$ was the only variation mentioned by the authors in the original paper~\cite{sandberg2005logic}. We note that the contour evolved much more slowly
-with a smalled $\lambda$, and thus took much more CPU time, and a larger number of iterations.
-
-\begin{figure*}[t!]
+\begin{figure*}[t!h!]
 \centering
-\includegraphics[width=12cm]{flowers.png}
+\includegraphics[width=0.5\textwidth]{flowers.png}
 \caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
@@
 
-\begin{figure}[t!]
-\centering
-\includegraphics[width=0.3\textwidth]{threech.png}
-\caption{Logic operations performed on three channels. The figure shows the final segmentation in each case. Size = 200 * 200. $\phi_{0}(x,y) = - \sqrt{(x -
-100)^2 + (y - 100)^2)} + 50$, $\mu = 0.1$, no reinitialization. For union: cpu = 3.92 sec and iterations = 6. For intersection: cpu = 3.11 sec and iterations =
-5. For intersection then union: cpu = 1.24 and iterations = 4.}
-\label{fig:threech}
-\end{figure}
+\section{Difficulties and Discussion}
+\label{sec:difficulties}
+We have faced many challenges throughout the implementation of this project. Some of the challenges caused major delays in the implementation cycle and hindered
+us from improving our code and adding more features to it.
+
+\subsection{Varying $\mu$}
+$\mu$ is a scaling factor that is used in both the Chan-Vese model and the Sandberg-Chan model to control the weight given to the length of the contour.
+Therefore, varying the value of $\mu$ will affect the ability to segment objects in noisy images, especially when the noise is similar to Matlab's \textit{salt
+and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure~\ref{fig:cvmu} shows the
+segmentation of an object in a noisy image for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting
+that
+this challenge kept us busy for over two days, after which we decided we should move on to other parts of the project.
 
@@
 \centering
-\includegraphics[width=0.5\textwidth]{sc_noisy.png}
-\caption{Same image as Figure~\ref{fig:threech}, but with 'salt \& pepper' noise with a variation of 0.1. Decreasing $\lambda$ from 255*255 to 125 allowed less
+\includegraphics[width=0.4\textwidth]{sc_noisy.png}
+\caption{Union of noisy channels. Same image as Figure~\ref{fig:threech}, but with \textit{Salt \& pepper'} noise with a variation of 0.1. Decreasing $\lambda$
+from 255*255 to 125 allowed less
 noise to be detected. Size = 200 x 200, $\phi_{0}(x,y) = - \sqrt{(x - 100)^2 + (y - 100)^2)} +
@@
 \end{figure}
-
-\subsubsection*{Complement}
-
-Unfortunately, the complement did not work properly in our implementation. We closely followed the definitions in the original paper where $z_i^{in'} = 1 -
-z_i^{in}$ and $z_i^{out'} = 1 - z_i^{out}$. The results was actually successfully detected, but continuous flickering occurred, and the curve did not stop in
-one position. We discuss this results in more details in Section~\ref{sec:difficulties}.
-
-\subsubsection*{Other Tests}
-
-Since the Sandberg-Chan model can successfully detect objects in one channel which are not in another when performing the union, we were curious to know
-whether this would allow us to perform successful segmentation on colored images by combining the information in the three channels R,G,B. Accordingly, we used
-the image in Figure~\ref{fig:cv_eg7} which the Chan-Vese model was not able to completely segment. We extracted three channels from the colored image: Red,
-Green, Black, and performed the union on the three channels. The results are shown in Figure~\ref{fig:flowers}. The complete contour of the flowers was
-successfully detected unlike that obtained in Figure~\ref{fig:cv_eg7}.
-
-\section{Difficulties and Discussion}
-\label{sec:difficulties}
-We have faced many challenges throughout the implementation of this project. Some of the challenges caused major delays in the implementation cycle and hindered
-us from improving our code and adding more features to it.
-
-\subsection{Varying mu}
-$\mu$ is a scaling factor that is used in both the Chan-Vese model and the Sandberg-Chan model to control the weight given to the length of the contour.
-Therefore, varying the value of $\mu$ will affect the ability to segment objects in noisy images, especially when the noise is similar to Matlab's \textit{salt
-and pepper} noise. However, varying $\mu$ in our implementation does not really have any impact on the length of the contour. Figure~\ref{fig:cvmu} shows the
-segmentation of an object in a noisy for different values for $\mu$. The resulting segmentation is the same for all values of $\mu$. It is worth noting that
-this challenge kept us busy for over two days, after which we decided we should move on to other parts of the project.
 
@@
 
-\subsection{Computing the complement}
+\subsection{Computing the Complement}
 The Sandberg-Chan model supports the complement logic operator, so one can segment the logic object ($A \cap \neg B$) (i.e. segment what is in channel A and not
@@
 
-Although the definition is very simple, it does not work in practice. We implemented the complement logic operator as defined in Equation~\ref{eqn:complement},
+Although the definition is very simple, it did not work in practice. We implemented the complement logic operator as defined in Equation~\ref{eqn:complement},
 and the segmentation does not stop at the required result. It rather keeps evolving after reaching the required solution, then keeps
-flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure~\ref{fig:sccomp} shows this scenrio. We tried to
+flipping between two segmentations: the correct segmentation, and the segmentation of $A \cap B$. Figure~\ref{fig:sccomp} shows this scenario. We tried to
 debug the reasons for this problem for three days.
@@
 
-\subsection{Defining the initial contour phi0}
-Both the Chan-Vese model and the Sandberg-Chan model defines the initial contour $\phi_0$ as a signed distance function where $\phi_0$ is positive inside the
+\subsection{Defining the initial contour $phi_0$}
+Both the Chan-Vese model and the Sandberg-Chan model define the initial contour ($\phi_0$) as a signed distance function where $\phi_0$ is positive inside the
 contour $C$, zero on the boundary of $C$ ($\partial C$), and negative outside the contour $C$. Figure~\ref{fig:phinote} illustrates those assumptions.
@@
 \centering
-\includegraphics[width=0.5\textwidth]{phinote.png}
+\includegraphics[width=0.3\textwidth]{phinote.png}
 \caption{Assumptions about the way $\phi_0$ is defined.}
@@
 union). After spending quite sometime to resolve this problem and checking what might be the source of the error, we tried using the regular equation of a
-circle to initialize $\phi_0$ (Equation~\ref{eq:phinotenormal]}). We were very intrigued to notice that this modification resulted in the correct segmentation
+circle to initialize $\phi_0$ (Equation~\ref{eqn:phinotenormal}). We were very intrigued to notice that this modification resulted in the correct segmentation
 of the input channels. Figure~\ref{fig:phidef} shows the final segmentation using both definitions for $\phi_0$.
@@
 \centering
-\includegraphics[width=0.2\textwidth]{phidef.png}
-\caption{}
+\includegraphics[width=0.4\textwidth]{phidef.png}
+\caption{Union and intersection depending on the way $\phi$ is defined}
 \label{fig:phidef}
@@
 \subsection{Mistyped Equations}
-As previously mentiond in Section~\ref{sec:sandberg-chan}, there is a typo in the equations that define $z_in$ and $z_out$ in the Sandberg-Chan model
-(equation 3 in thier paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
+As previously mentioned in Section~\ref{sec:sandberg-chan}, there is a typo in the equations that define $z_{in}$ and $z_{out}$ in the Sandberg-Chan model
+(equation 3 in their paper). In addition, they have another typo in the expansion of the the Eular-Lagrange equation for the PDE that models the problem. Both
 typos consumed over a week of our time until we were able to find out what are the correct equations in both cases. We could have spent this time overcoming
@@
 useful when an edge-based segmentation is not applicable (i.e. edges do not depend on a gradient). However, our implementation fails to respond to different
-values for the contour length scaling parameter, $\mu$. This affects the ability of our implementation to handle Matlab's \textit{salt \& pepper} noise. On the
+values for the contour length scaling parameter, $\mu$. This affects the ability of our implementation to handle Matlab's \textit{Salt \& pepper} noise. On the
 other hand, changing the values for the intensity weight parameters $\lambda_1$ and $\lambda_2$ affects the amount of details obtained inside and outside the
@@
 We have also shown that our implementation for the Sandberg-Chan logic framework handles different logic combinations for multiple channels. We have also
-experimented using it as a segmentation tool for colored images by modeling an input colored images as a 3-channel image (red, green, blue). However, our
+experimented using it as a segmentation tool for colored images by modeling an input colored images as a 3-channels (red, green, blue). However, our
 implementation for the \textit{complement} logic operation seems to be missing a hidden trick in the Sandberg-Chan logic framework.
@@
 \bibliographystyle{ieee}
+
+\appendix
+\appendixpage
+
+\section{Division of Labor}
+
+We both implemented the Chan-Vese model together since this was the starting point of the project. Sarah did most of the initial implementation for the
+Sandberg-Chan model by defining all the necessary functionalities for the logic operations for two images. Karim then extended it to multiple images. Karim
+also tested the Sandberg-Chan model, and took most of the snapshots in that part for the paper. Sarah tested the Chan-Vese model, and took all its snapshots
+for the paper. We both worked on finding an appropriate stopping condition for the contour evolution, and invested a lot of time to debug the problems we
+faced that were described in Section~\ref{sec:difficulties}.
 

### 2010-12-13 trunk/cs870/project/Report/Paper.tex
Final commit isA isA isA

@@
 where $\phi > 0$) and the mean of the values outside $\phi$ (specifically where $\phi < 0$) respectively. To actually solve the PDE, we chose to use an explicit
-time stepping scheme for the finite difference discretization. That is, in each time step $\phi^{n+1} = \phi^n + \triangle t*phi_t$. This was simpler to
+time stepping scheme for the finite difference discretization. That is, in each time step $\phi^{n+1} = \phi^n + \triangle t*\phi_t$. This was simpler to
 implement, and although $\triangle t$ must be chosen carefully to satisfy the stability condition ($\triangle t <= (1/imageSize)^2$), we did not suffer from
@@
 
-\begin{figure}[h!]
+\begin{figure*}[t!]
 \centering
@@
 \label{fig:family}
-\end{figure}
+\end{figure*}
 
@@
 \centering
-\includegraphics[width=12cm]{donutlogic.png}
+\includegraphics[width=0.5\textwidth]{donutlogic.png}
 \caption{Performing union and intersection on images with holes. Size = 300 x 300. $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} +
@@
 
-\begin{figure}[t!h!]
+\begin{figure}[h!]
 \centering
@@
 
-\begin{figure}[t]
+\begin{figure*}[t]
 \centering
-\includegraphics[width=0.4\textwidth]{sc_blurry.png}
+\includegraphics[width=0.3\textwidth]{sc_blurry.png}
 \caption{Successful union with blurred images. Top: the evolving curve for $A_1 \cup A_2$
@@
 \label{fig:sc_blurry}
-\end{figure}
+\end{figure*}
+
+\begin{figure*}[t!h!]
+\centering
+\includegraphics[width=0.5\textwidth]{flowers.png}
+\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
+Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
+size = cpu = 1.73 sec, iterations = 7.}
+\label{fig:flowers}
+\end{figure*}
 
@@
 this challenge kept us busy for over two days, after which we decided we should move on to other parts of the project.
-
-\begin{figure*}[t!]
-\centering
-\includegraphics[width=12cm]{flowers.png}
-\caption{Ability to detect the full contour on a colored image with low variation between foreground and background by taking the union of the RGB channels.
-Full contour detected as opposed to Figure~\ref{fig:cv_eg7}. Size = 300 x 300, $\phi_{0}(x,y) = - \sqrt{(x - 150)^2 + (y - 150)^2)} + 75$, no reinitialization,
-size = cpu = 1.73 sec, iterations = 7.}
-\label{fig:flowers}
-\end{figure*}
 
@@
 
-\subsection{Defining the initial contour $phi_0$}
+\subsection{Defining the initial contour $\phi_0$}
 Both the Chan-Vese model and the Sandberg-Chan model define the initial contour ($\phi_0$) as a signed distance function where $\phi_0$ is positive inside the
